### 2022-05-10 paper/main.tex
Initial Overleaf Import

% Copyright 2022 the authors

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry} % because Overleaf is weird.
\PassOptionsToPackage{hyphens}{url}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
%TODO: change N to N^3 in the appropriate locations
% define epsilon, they depend on the dimension

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.40in}
\setlength{\topmargin}{-0.50in}
\setlength{\footskip}{0.25in}
\pagestyle{myheadings}
\markboth{foo}{Hogg \& Villar / Emulating n-body simulations with equivariant CNNs}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

\title{\bfseries%
Emulating cosmological n-body simulations with equivariant convolutional neural networks}
\author{Hogg, Villar, others}
\date{2022 April}

\begin{document}

\maketitle\thispagestyle{empty}

\section{Introduction}

Cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
For this reason, accurate, fast emulators would be of enormous economic and scientific value.
This point has not gone unnoticed!
There are now quite a few successful projects implementing such emulators (CITE THINGS).

However, there are dangers involved in emulation of this kind.
HOGG: Epistemological and ontological problems!

The epistemological problems can be somewhat ameliorated (though not entirely solved) if the emulators are required to obey all of the exact symmetries that the n-body simulations themselves are expected to obey.
That is what motivates the present project.

HOGG: Summarize here how we are going to proceed, in human language.

\section{Problem setup}

For our purposes, a cosmological simulation is a set of $N^3$ particles $n$ with $1\leq n\leq N^3$, each of which has a (scalar) mass $m_n$, a (vector) position $q_n$, and a (vector) veloctiy $v_n$, and each of which is a function of time (or computed or stored at a set of times $t$).
The positions $q_n$ live in a cubical 3-torus of side length $L$.
In the dark-matter-only case, the simulation dynamics depend only on the masses, positions, velocities $\{m_n, q_n, v_n\}_{n=1}^N$, time $t$, Newton's constant $G$, and (hopefully weakly) the number $N^3$ and length $L$.

Because the simulations are designed to describe the evolution of the Universe, the particles begin in a nearly-uniform density state, usually very close to a uniform, torus-filling, cubical lattice.
The particles begin at a very early time $t$ with tiny (vector) offsets $\Delta q_n$ away from lattice sites $q^{(0)}_n$ such that $q_n = q^{(0)}_n + \Delta q_n$, and tiny (vector) velocities $v_n$.
In detail these positional offsets and velocities are closely related by something like a Zeldovich approximation.

The uniform, torus-filling, cubical lattice of $N^3$ points $q^{(0)}_n$ can be thought of as a 3-dimensional ``image''.
Each particle can be thought of as a ``pixel'' in that image.
At each snapshot of the simulation at time $t$, each pixel has an associated (vector) positional offset $\Delta q_n$, (vector) velocity $v_n$, and a few scalars ($m_n, t, G$ at least).
In addition to these native scalars, we can produce at each pixel $n$ three additional scalars $q_n^\top q_n, q_n^\top v_n, v_n^\top v_n$, which are the inner products of the vectors with themselves and each other.
At late times, the simulation shows almost no memory of the initial uniform, torus-filling, cubical lattice.
However, each particle has an identity $n$ that can be associated with its original lattice location $q^{(0)}_n$, so its properties can be thought of as being encoded in that original lattice-location pixel.
That is, this is a problem of scalars and vectors on a uniform, cubical lattice in a cubical 3-torus.

There are some exact and approximate symmetries to consider in this problem.
The approximate symmetries are the fundamental symmetries of classical physics, which include translation, rotation, and parity.
These symmetries are only approximate in the n-body simulation context, because there is a cubical 3-torus with an orientation and a size, and there is a cubical lattice inside that 3-torus.
Generators of the exact symmetries are the three torus-axis-aligned spatial rotations by $\pi/2$ and one reflection across one of the three axes of the torus.

\section{Linear geometric convolution operators}

We consider the orthogonal group $O(d)$ and $g\in O(d)$. For our purposes $d$ will be 2 or 3, but everything generalizes.  
We consider an image $A$ in a cubical $d$-torus with $N$ equally spaced pixels in each dimension ($N^d$ pixels total). Each pixel can contain tensors and pseudotensors (of which scalars, pseudoscalars, vectors, and pseudovectors are particular cases) that we define below.
Similarly, each linear geometric convolution operator $C$ will be a pixelized cubical image patch with $M$ pixels (with $M$ odd) in each dimension ($M^d$ pixels total) with a tensor or pseudotensor in each pixel, which acts as a linear pixel weight.

For some of the expressions below we will use Ricci and Levi-Civita summation notation (or Einstein summation notation).
In this notation, tensor products are written in component form, and repeated indices are summed over.
In this notation, the product of two 2-tensors (represented as two $d\times d$ matrices $A$ and $B$) is written as
\begin{equation}
    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} ~,
\end{equation}
where $[A]_{i,k}$ (for example) is the $i,k$ element of matrix $A$, the repeated index $k$ is implicitly summed over the integers from 1 to $d$.
This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice).

\begin{definition}[vectors, $k$-tensors, and the world]
Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
It also acts on lists of vectors $(\mathbb R^d)^n$ by acting on each of the $n$ vectors by the same action.
It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = g\cdot v_1\otimes \ldots \otimes g\cdot v_k$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
Note that $O(d)$ can be seen as acting on the world.
\end{definition}

One consequence of the above defnitions is that 
$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor iff (in summation notation)
$[g\cdot T]_{i_1,\ldots, i_k} = [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}$ for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
For example, a 2-tensor is defined by its transformation property
$[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
which, in normal matrix notation, is written as
$g\cdot T = M(g)\,T\,M(g)^\top$.

\begin{definition}[pseudovector and $k$-pseudotensor]
SOLE CAN WE SAY HERE that these are ``like a vector and like a tensor'' but there is an additonal $\det(g)$ coming in?
\end{definition}

\begin{definition}[scalar and pseudoscalar]
We say that $a\in \mathbb R$ is a $0$-tensor or scalar iff $a$ is an invariant of the action.
Namely, $a(W) = a(g\cdot W)$ for all $g\in O(d)$.
We say that
$a\in \mathbb R$ is a $0$-pseudotensor or pseudoscalar iff $a(g \cdot W) = \det(g)\,a(W)$ for all $g\in O(d)$,
where $\det(g)$ is the determinant of the matrix representation $M(g)$ of $g$.
\end{definition}

\begin{definition}[outer products of tensors]
If $a$ is a $k$-tensor and $b$ is a $k'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$
Similarly, if $a$ is a $k$-pseudotensor and $b$ is a $k'$-pseudotensor, then $a\otimes b$ is a $(k+k')$-tensor.
If $a$ is a $k$-tensor and $b$ is a $k'$-pseudotensor, or if $a$ is a $k$-pseudotensor and $b$ is a $k'$-tensor, then $a\otimes b$ is a $(k+k')$-pseudotensor.
\end{definition}

\begin{definition}[linear geometric convolution]
Given an image $A$ of geometric objects with $N^d$ pixels, we define its linear convolution with a geometric filter $C$ as follows.
Each pixel $p$ (numbered $1\leq p\leq N^d$) of the image contains a geometric object $a_p$.
All of these $N^d$ geometric objects $a_p$ are of the same geometric type ($k$-tensor or $k$-pseudotensor).
The filter $C$ has side length $M\leq N$, where $M$ is an odd integer, such that it has $M^d$ total pixels.
Each pixel $m$ (numbered $1\leq m\leq M^d$) of the filter contains a geometric object $w_m$.
All of these $M^d$ geometric weights $w_m$ are of the same geometric type as well ($k'$-tensor or $k'$-pseudotensor).
For image pixel $p$ (with $1\leq p\leq N^d$) we take a $M^d$ patch of the image, centered at $p$ with contents $a(p)$ (where we assume that the image is in the $d$-torus, so every pixel centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the patch $P_p$ (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
\begin{equation}
    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m.
\end{equation}
\end{definition}

If the image pixels $a_m$ are $k$-tensors and the filter weights $w_m$ are $k'$-tensors, or if the image pixels are $k$-pseudotensors and the weights are $k'$-pseudotensors, then the output of the linear filter is an image of $(k+k')$-tensors.
If the image pixels are $k$-pseudotensors and the weights are $k'$-tensors, or if the image pixels are $k$-tensors and the weights are $k'$-pseudotensors, then the output is an image of $(k+k')$-pseudotensors.
This linear geometric convolution operation can be followed by contractions with Kroenecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.

Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-tensor.
Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired. SOLE DO WE NEED TO SAY MORE HERE?
Contraction of a $k$-pseudotensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-tensor.

For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $[a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
$[a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.

With these definitions we can make geometric filters, take outer products, and contract to make a combinatorically large set of nonlinear geometric convolution operators on images made up of geometric objects.
But before we go nonlinear, we will go equivariant.

\section{Equivariant linear geometric convolution operators}

We will now create linear convolution operators that are group-equivariant, for groups of reflections, rotations, and translations (on the cubical torus).
We will construct these equivariant convolution operators by an averaging over group operators, followed by a matrix factorization (singular-value decomposition).

\begin{definition}[Equivariant convolution operator]
Foo and bar.
\end{definition}

Given a $k\times k \times k$ patch ($k$ odd), centered on a particular pixel, we can consider the convolution operator with a fixed $k\times k \times k$ filter. The filter in each pixel can contain a tensor or pseudotensor

HOGG: What is a linear convolution operator?

HOGG: What is a linear geometric convolution operator?

HOGG: What would it mean for a linear geometric convolution operator to be equivariant to all relevant groups?
It's naturally translation-equivariant!
But then how do we do reflections and rotations?
Be sure to do scalar, pseudoscalar, vector, pseudovector, 2-tensor, and 2-pseudotensor cases.

HOGG: How does the group act, on the image, and on the linear geometric convolution filter, and both?

\section{Equivariant nonlinear geometric convolutions}

\section{Construction of all equivariant convolution operators}

Group averaging and SVD to get all equivariant linear convolution operators.

Outer products and contractions to get all equivariant nonlinaer convolution operators.

Lots of pretty pictures.

\end{document}

### 2022-05-19 paper/main.tex

@@
 It also acts on lists of vectors $(\mathbb R^d)^n$ by acting on each of the $n$ vectors by the same action.
-It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = g\cdot v_1\otimes \ldots \otimes g\cdot v_k$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
+It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
 Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
@@
 
-\begin{definition}[linear geometric convolution]
-Given an image $A$ of geometric objects with $N^d$ pixels, we define its linear convolution with a geometric filter $C$ as follows.
-Each pixel $p$ (numbered $1\leq p\leq N^d$) of the image contains a geometric object $a_p$.
-All of these $N^d$ geometric objects $a_p$ are of the same geometric type ($k$-tensor or $k$-pseudotensor).
-The filter $C$ has side length $M\leq N$, where $M$ is an odd integer, such that it has $M^d$ total pixels.
-Each pixel $m$ (numbered $1\leq m\leq M^d$) of the filter contains a geometric object $w_m$.
-All of these $M^d$ geometric weights $w_m$ are of the same geometric type as well ($k'$-tensor or $k'$-pseudotensor).
-For image pixel $p$ (with $1\leq p\leq N^d$) we take a $M^d$ patch of the image, centered at $p$ with contents $a(p)$ (where we assume that the image is in the $d$-torus, so every pixel centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
-Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the patch $P_p$ (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
+HOGG: Define a geometric image here.
+
+HOGG: Define a geometric filter here. Then shorten the defition below.
+
+\begin{definition}[geometric convolution, geometric convolution filter]
+Let $A$ be a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k$-tensor or $k$-pseudotensor as defined above.
+The image $A$ lives in a cubical $d$-torus such that the image has no boundary.
+Let $C$ be a $d$-dimensional image with $M^d$ pixels (with $M$ odd and $M<N$), in each of which there is a $k'$-tensor or $k'$-pseudotensor.
+We call $C$ a geometric convolution filter.
+The geometric convolution $A\star C$ is a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k+k'$-tensor or $k+k'$-pseudotensor, defined as follows:
+For image pixel $p$ in $A\star C$ (with $1\leq p\leq N^d$) we take the cubical $M^d$ patch of the image centered on pixel $p$ with contents $a(p)$ (the image is on the $d$-torus, so every pixel $p$ centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
+Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the cubical patch (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
 \begin{equation}
-    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m.
+    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m~.
 \end{equation}
+That is, geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor and pseudotensor inputs and outputs.
 \end{definition}
@@
 
-Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
+THIS PARAGRAPH SUX: Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
 The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
@@
 
-For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $[a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
+For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
 For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
-$[a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.
+$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.
 
@@
 
-\section{Equivariant linear geometric convolution operators}
+\section{Equivariant geometric convolution operators}
 
@@
 
+\begin{definition}[Group $B_d$ of symmetries of a $d$- hypercube]
+We denote by $B_d$ the group of rotation and reflection symmetries of the $d$-dimensional hypercube.
+\end{definition}
+
+Notes about the group $B_d$: In $d=2$ there are $8$ group elements, in $d=3$ there are $48$ group elements (and in $d=4$ there are 384).
+Because the groups $B_d$ are subgroups of $O(d)$, all determinants of the matrix representations of the group elements are either $+1$ or $-1$, and the matrix representation $M(g^{-1})$ of the inverse $g^{-1}$ of group element $g$ is the transpose of the matrix representation $M(g)$ of group element $g$.
+
+\begin{definition}[Action of $B_d$ on $k$-tensors and $k$-pseudotensors]
+\begin{align}
+\mbox{$k$-tensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= (g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) \\
+\mbox{$k$-pseudotensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= \det(g)\,(g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) ~,
+\end{align}
+And extend linearly
+\end{definition}
+
+This may seem familiar: The $k$-tensor and $k$-pseudotensor were defined by how they transform under the orthogonal group $O(d)$, and our group $B_d$ is a subgroup of the orthogonal group.
+
+\begin{definition}[Action of $B_d$ on geometric images]
+Given $A$ a cubic $N^d$ image with a $k$-tensor in each pixel, and $g\in B_d$ we define $g\cdot A$ a $N^d$-dimensional image with $k$-tensors in each pixel such that
+\begin{equation}
+    (g\cdot A)_p = g\cdot A_{g^{-1}\cdot p}
+\end{equation}
+\end{definition}
+
+\begin{definition}[The total group $G_d$ in play]
+Outer product of the above group and discrete translations on the $d$-cubic lattice.
+\end{definition}
+
 \begin{definition}[Equivariant convolution operator]
-Foo and bar.
+A geometric convolution filter $C$ is equivariant if it is an equivariant function with respect to the action above. 
 \end{definition}
@@
 The filter in each pixel can contain a tensor or pseudotensor
-
-HOGG: What is a linear convolution operator?
-
-HOGG: What is a linear geometric convolution operator?
 
@@
 
-\section{Equivariant nonlinear geometric convolutions}
+\section{Nonlinear equivariant local geometric operators}
 
-\section{Construction of all equivariant convolution operators}
+\section{Construction of all equivariant operators}
 

### 2022-07-18 paper/main.tex
Updates from Overleaf

@@
 % Copyright 2022 the authors
+
+% To-Do list:
+% -----------
+% - What are our GOALS!!?
+%   -- one option: this formulation can express all equivariant differential operators discretized to a cubical d-torus. 
+%   -- another option: can we produce a characterization of all equivariant functions (maybe based on invariants or these filters) that don't require the use of non-linearities.
+%   -- another option: emulate cosmological simulations.
+% - change N to N^3 in the appropriate locations; already done??
+% - Understand what functions can and cannot be expressed on this model 
+% - Extend it to non-linear maps (eg, by multiplications and contractions)
+% - Should we be representing images in fourier space? Ask Leslie!
+% - Understand the literature
+%   -- Discretized vector calculus and differential geometry.
+%   -- Group-equivariant CNNs; what's been done?
+%   -- Understand how this formulation compares with steerable CNNs
+%   -- What is known about what functions CNNs can represent? Ask Joan!
+%  -- Define "local" functions
+
+% -- if the size of the filter is large enough, are we universal???? or can we express all polynomial functions? is there a version of Stone-Weierstrass for this?
+% --- add the identity and levi civita in the possible output tensors
+% --- define reordering of the indices
+% --- we don't need to be in the torus and we don't need to be square (or cubical)
+% --- generalize convolution to rectangular non-toroidal images
 
@@
 \usepackage{amsthm}
-\theoremstyle{definition}
+
+% theorem, definition and so on
+\theoremstyle{plain}
 \newtheorem{definition}{Definition}
-%TODO: change N to N^3 in the appropriate locations
-% define epsilon, they depend on the dimension
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{theorem}{Theorem}
+
+% text macros
+\newcommand{\sectionname}{Section}
+\newcommand{\secref}[1]{\sectionname~\ref{#1}}
 
@@
 \title{\bfseries%
-Emulating cosmological n-body simulations with equivariant convolutional neural networks}
-\author{Hogg, Villar, others}
-\date{2022 April}
+Equivariant convolution-based functions
+of images or lattices
+containing scalars, vectors, and tensors}
+\author{Hogg, Villar, Blum-Smith, others}
+\date{2022 June}
 
@@
 
+\paragraph{Abstract:} HERE IS OUR ABSTRACT.
+
 \section{Introduction}
 
-Cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
-For this reason, accurate, fast emulators would be of enormous economic and scientific value.
-This point has not gone unnoticed!
+Contemporary physics teems with data sets that are images or lattices or grids of geometric objects.
+Rectangular images or lattices of density (a scalar field), temperature (also a scalar field), velocity (a vector field), magnetic field (a pseudo-vector field), and polarization (a 2-tensor field) are regularly taken with detectors or made in simulations.
+That is, physics has lots of data that consists of geometric objects on regular grids.
+These objects (scalars, vectors, pseudo-vectors, and tensors, for example) are \emph{geometric} in the sense that they are defined by their transformation properties under rotation and reflection.
+For example, scalars are defined to be quantities in $\mathbb R$ that are invariant with respect to all rotations and reflections; vectors are defined (in $d$-dimensional space) to be quantities in $\mathbb R^d$ that transform under rotation or reflection by multiplication by the orthogonal $d\times d$ matrix representing the rotation or reflection; and so on.
+
+For one concrete example, cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
+For this reason, accurate, fast emulators might be of enormous economic and scientific value.
 There are now quite a few successful projects implementing such emulators (CITE THINGS).
+Importantly, the initial conditions for these simulations are tiny displacements (vectors) and velocities (vectors) relative to a cubical lattice on a 3-torus.
+That is, the initial conditions can be described as a 3-dimensional image of vectors on a regular grid.
+For another concrete example, astronomical imaging data taken of the interstellar medium can be represented as rectangular lattices of measurements of temperature (scalar), optical depth (scalar), and induced polarization of starlight (a 2-tensor).
+Here the data can be described as a 2-dimensional image of scalars and tensors on a regular grid.
 
-However, there are dangers involved in emulation of this kind.
-HOGG: Epistemological and ontological problems!
+The application of machine-learning techniques to these problems can be dangerous:
+Machine-learning models have enormous freedom; the laws of physics are constrained to obey a large set of extremely strict symmetries, including rotation, reflection, translation, permutation, and boost equivariances, plus constraints such as units and dimensions symmetry, unitarity, and myriad conservation laws.
+If we want to operate safely on these problems, we want these symmetries built in from the start, and exactly.
 
-The epistemological problems can be somewhat ameliorated (though not entirely solved) if the emulators are required to obey all of the exact symmetries that the n-body simulations themselves are expected to obey.
-That is what motivates the present project.
+HOGG: Differences between continuous and discrete symmetries.
 
@@
 
-\section{Problem setup}
+\section{Problem setup}\label{sec:setup}
 
@@
 
-\section{Linear geometric convolution operators}
+\section{Geometric convolutions}\label{sec:convolution}
 
@@
 
-One consequence of the above defnitions is that 
+One consequence of the above definitions is that 
 $T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor iff (in summation notation)
@@
 
-HOGG: Define a geometric image here.
-
-HOGG: Define a geometric filter here. Then shorten the defition below.
-
-\begin{definition}[geometric convolution, geometric convolution filter]
-Let $A$ be a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k$-tensor or $k$-pseudotensor as defined above.
-The image $A$ lives in a cubical $d$-torus such that the image has no boundary.
-Let $C$ be a $d$-dimensional image with $M^d$ pixels (with $M$ odd and $M<N$), in each of which there is a $k'$-tensor or $k'$-pseudotensor.
-We call $C$ a geometric convolution filter.
-The geometric convolution $A\star C$ is a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k+k'$-tensor or $k+k'$-pseudotensor, defined as follows:
-For image pixel $p$ in $A\star C$ (with $1\leq p\leq N^d$) we take the cubical $M^d$ patch of the image centered on pixel $p$ with contents $a(p)$ (the image is on the $d$-torus, so every pixel $p$ centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
-Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the cubical patch (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
+\begin{definition}[contractions of tensors]
+Let $a$ be a $k$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
+We define the $a^{(\alpha,\beta)}$ contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$.
+In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$'s are placed in the positions $\alpha$ and $\beta$:
 \begin{equation}
-    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m~.
+[a^{(\alpha,\beta)}]_{i_1,\ldots i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, x, \ldots , x, \ldots, i_k}
 \end{equation}
-That is, geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor and pseudotensor inputs and outputs.
 \end{definition}
 
-If the image pixels $a_m$ are $k$-tensors and the filter weights $w_m$ are $k'$-tensors, or if the image pixels are $k$-pseudotensors and the weights are $k'$-pseudotensors, then the output of the linear filter is an image of $(k+k')$-tensors.
-If the image pixels are $k$-pseudotensors and the weights are $k'$-tensors, or if the image pixels are $k$-tensors and the weights are $k'$-pseudotensors, then the output is an image of $(k+k')$-pseudotensors.
-This linear geometric convolution operation can be followed by contractions with Kroenecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
+\begin{definition}[permutations of tensor indices]
+MSV WRITE THIS.
+\end{definition}
 
-THIS PARAGRAPH SUX: Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
+\begin{definition}[$k$-tensor image]
+We say that $A$ is a $k$-tensor image if $A$ is a function $A: [N]^d \to (\mathbb R^d)^{\otimes k}$ such that $A(\bar \i)$ is a $k$-tensor for all $\bar \i \in [N]^d$, and $[N]=\{0,\ldots, N-1\}$.
+By a similar definition there are also $k$-pseudotensor images.
+We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
+\end{definition}
+
+\begin{definition}[geometric convolution]
+Let $A$ be a $k$-tensor image on the $d$-torus with side length $N$.
+Let $C$ be a $k'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
+The geometric convolution $A\ast C$ is a $k+k'$-tensor image such that
+\begin{equation}
+    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
+\end{equation}
+where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
+If $A$ is a $k$-pseudotensor image and $C$ is a $k'$-pseudotensor filter, then $A\ast C$ is also a $k+k'$-tensor.
+If $A$ is $k$-tensor and $C$ is $k'$-pseudotensor, or if $A$ is $k$-pseudotensor and $C$ is $k'$-tensor, then $A\ast C$ is a $k+k'$-pseudotensor.
+\end{definition}
+
+Geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor and pseudotensor inputs and outputs.
+
+The linear geometric convolution operation can be followed by contractions with Kroenecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
+The Kroenecker delta is the object with 2 indices such that if the two indices are identical the value is $1$ and the value is $0$ otherwise.
 The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
@@
 
-With these definitions we can make geometric filters, take outer products, and contract to make a combinatorically large set of nonlinear geometric convolution operators on images made up of geometric objects.
+With these definitions in hand, we can make geometric filters, take outer products, and contract to make a combinatorically large set of nonlinear geometric convolution operators on images made up of geometric objects.
 But before we go nonlinear, we will go equivariant.
 
-\section{Equivariant geometric convolution operators}
+\section{Equivariant geometric convolutions}\label{sec:equivariant}
 
@@
 
-\begin{definition}[Action of $B_d$ on geometric images]
-Given $A$ a cubic $N^d$ image with a $k$-tensor in each pixel, and $g\in B_d$ we define $g\cdot A$ a $N^d$-dimensional image with $k$-tensors in each pixel such that
+\begin{definition}[Action of $B_d$ on $k$-tensor images]
+Given a $k$-tensor image $A$ on the $d$-torus, and a group element $g\in B_d$, the action $g\cdot A$ produces a $k$-tensor image on the $d$-torus such that
 \begin{equation}
-    (g\cdot A)_p = g\cdot A_{g^{-1}\cdot p}
+    (g\cdot A)(\bar\i) = g\cdot A({g^{-1}\cdot \bar\i}) ~.
 \end{equation}
+SOmeThiNg about how $g^{-1}\cdot\bar\i$ works here, possibly preceded or followed by proofs.
+Because of what follows, this has to be defined BOTH for the $N^d$ boxels in the $d$-torus and on the $(2\,m+1)^d$ boxels of the filter (image patch).
 \end{definition}
 
-\begin{definition}[The total group $G_d$ in play]
-Outer product of the above group and discrete translations on the $d$-cubic lattice.
+It might be a bit surprising that the group element $g^{-1}$ appears in the definition of the action of the group on images.
+One way to think about it is that the pixels in the output (rotated, say) image are ``looked up'' or ``read out'' from the pixels in the original (unrotated) image.
+The pixel locations in the original image are found by going back, or inverting the transformation.
+
+\begin{definition}[The group $G_d$, and its action on $k$-tensor images]
+THIS NEEDS TO BE WRITTEN: $G_d$ is the group generated by the elements of $B_d$ and the discrete translations on the $N^d$-pixel lattice on the $d$-torus.
 \end{definition}
 
-\begin{definition}[Equivariant convolution operator]
-A geometric convolution filter $C$ is equivariant if it is an equivariant function with respect to the action above. 
+HOGG: Comment on the fact that $G_d$ is a subgroup of the Euclidean group, but that we care about the Euclidean group because we are imagining that the discrete images are images of a continuous world!!
+
+\begin{definition}[Equivariant convolution]
+We say that a convolution filter $C$ produces convolutions that are equivariant with respect to the group $G_d$ if for any $k$-tensor image $A$ on the $d$-torus, and any group element $g\in G_d$, we have
+\begin{equation}
+    g\cdot (A\ast C) = (g\cdot A)\ast C ~,
+\end{equation}
+where the group actions are defined above.
 \end{definition}
 
-Given a $k\times k \times k$ patch ($k$ odd), centered on a particular pixel, we can consider the convolution operator with a fixed $k\times k \times k$ filter.
-The filter in each pixel can contain a tensor or pseudotensor
+The intuition here is that the filter $C$ represents some physical law, and the image $A$ represents the initial conditions or state of the physical system on which the law is acting.
+The physical law (represented by $C$) is invariant to translations, rotations, and parity, while the \emph{output} of the physical law (represented by $A\ast C$) is \emph{equivariant} to translations, rotations, and parity acting on the state $A$.
 
-HOGG: What would it mean for a linear geometric convolution operator to be equivariant to all relevant groups?
-It's naturally translation-equivariant!
-But then how do we do reflections and rotations?
-Be sure to do scalar, pseudoscalar, vector, pseudovector, 2-tensor, and 2-pseudotensor cases.
+The following theorem generalizes the Cohen and Welling paper:
 
-HOGG: How does the group act, on the image, and on the linear geometric convolution filter, and both?
+\begin{theorem}
+A $k'$-tensor (or $k'$-pseudotensor) convolution filter $C$ produces convolutions that are equivariant with respect to the big group $G_d$ if $C$ is invariant under the small group $B_d$.
+\end{theorem}
 
-\section{Nonlinear equivariant local geometric operators}
+\begin{proof}
+By Definition \ref{} we have
+\begin{eqnarray}
+(g\cdot (A* C) )(\bar \i) &=& g\cdot( (A*C)(g^{-1} \cdot \bar \i)) \\
+&=& g \cdot \left (\sum_{a\in [-m,m]^d} A(g^{-1}\cdot \bar \i - a) \otimes C(a)\right)\\
+&=& \sum_{a\in [-m,m]^d} g \cdot A(g^{-1} \cdot \bar \i -a ) \otimes g \cdot C(a).
+\end{eqnarray}
+By Definition \ref{} we have
+\begin{eqnarray}
+((g\cdot A)* C) (\bar \i) &=& \sum_{a\in [-m,m]^d} g\cdot A(g^{-1} \cdot (\bar \i - a) ) C(a) \\
+&=& \sum_{a'\in [-m,m]^d} g \cdot A(g^{-1}\cdot \bar \i -a') \otimes C(g\cdot a'),
+\end{eqnarray}
+where $a'=g\cdot a$. Therefore the convolution is equivariant if $g\cdot C(g^{-1} \cdot a') = C(a')$ for all $g\in B_d$ and all $a' \in [-m,m]^d$. 
+\end{proof}
 
-\section{Construction of all equivariant operators}
+\section{Nonlinear equivariant local geometric operators}\label{sec:nonlinear}
 
-Group averaging and SVD to get all equivariant linear convolution operators.
+The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
+They are local in the sense that, at every pixel $\bar\i$ of the output image $(A\ast C)$, the convolution only makes use of the pixels of $A$ within a small neighborhood of of $\bar\i$.
+Because tensors can be multiplied together using the outer product, and contracted according to the tensor contractions defined in \secref{sec:convolution}, it is possible, by multiplication, to construct a family of \emph{nonlinear} local geometric operators that take as input geometric images and produce as output geometric images.
+This family of nonlinear local operators will be equivariant to the group $G_d$ when the linear local operators of which they are composed are equivariant.
 
-Outer products and contractions to get all equivariant nonlinaer convolution operators.
+\begin{definition}[Products of geometric images]
+Given an $N^d$-pixel $k_A$-tensor image $A$ and an $N^d$-pixel $k_B$-tensor image $B$, we can define a product $A\otimes B$, which is an $N^d$-pixel $(k_A+k_B)$-tensor image such that $(A\otimes B)(\bar\i)=A(\bar\i)\otimes B(\bar\i)$.
+These products can be contracted by the contraction rules above.
+\end{definition}
 
-Lots of pretty pictures.
+\begin{definition}[Polynomial local geometric operator of degree $\ell$]
+Given a $k$-tensor image $A$ and a set of $\ell$ $k'_j$-tensor geometric filters $C_j$, a geometric image can be formed by the degree-$\ell$ image product
+\begin{equation}
+    (A\ast C_1) \otimes (A\ast C_2) \otimes \cdots \otimes (A\ast C_\ell) ~. \label{monomial}
+\end{equation}
+This product can be subsequently contracted $r$ times to deliver a $K$-tensor geometric image with
+\begin{equation}
+    K = \ell\,k - 2\,r + \sum_{j=1}^\ell k'_j ~.
+\end{equation}
+A polynomial local geometric operator is a function that takes a $k$-tensor image $A$ and outputs the sum of $K$-tensor images constructed as contractions to order $K$ of tensor images from \eqref{monomial} with filters chosen by the user.
+\end{definition}
+
+\section{Construction of equivariant geometric convolution filters}
+
+In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
+These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
+Here we show how to find all such filters by a group-averaging technique.
+
+Briefly, the idea is the following:
+We start with a set of independent $k'$-tensor (or $k'$-pseudotensor) convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M$ odd) and dimension $d$.
+There are typically $M^d\,d^{k'}$ independent filters in this set.
+We average the filters over all group operators $g\in B_d$, where $B_d$ is the group of 90-degree rotations and reflections.
+We perform a singular value decomposition of these group-averaged filters to find the $p$ independent orthogonal filters that span the space of equivariant $k'$-tensor (or $k'$-pseudotensor) convolution filters at side length $M$ in $d$ dimensions.
+
+In detail, group averaging happens HOW? That is, what is the exact mathematical operation?
+
+Show all the filters for $d=2$, $M=3$.
+
+Show all the filters for $d=2$, $M=5$.
+
+Show all the filters for $d=3$, $M=3$.
+
+Give a table saying how many filters there are for each $d$ and each $M$ at each $k$ and each parity.
+
+Now show some fake-data geometric images and show the action of the $d=2$, $M=3$ filters on those images.
+
+Now show contractions of products of the above on those images.
+
+\section{Implementation notes}
+\begin{itemize}
+    \item find the invariant geometric filters
+    \item geometric convolution operator
+    \item pixel-wise outer product
+    \item contraction operator, enumeration of all contractions
+    \item output index permutation, enumeration of all permutations
+    \item enumeration of all monomials up to a fix order and degree 
+    \item remove repeated objects (SVD or just lstsq)?
+    \item linear combinations 
+    \item loss function
+    \item optimization
+\end{itemize}
+
+\section{Discussion and comments}
+
+Note that we never take eigenvalues. But these are also real scalars in some cases.
+
+Note that pseudo-$k$-tensor filters only exist for certain triples of $(d, m, k)$.
 

### 2022-07-27 ipynb/geometric.py
start moving functions to py files

@@ -1,0 +1,564 @@
+import numpy as np
+
+# ------------------------------------------------------------------------------
+# PART 1: Make and test a complete group
+
+def make_all_generators(D):
+
+    # Make the flip operator
+    foo = np.ones(D).astype(int)
+    foo[0] = -1
+    gg = np.diag(foo).astype(int)
+    generators = [gg, ]
+
+    # Make the 90-degree rotation operators
+    for i in range(D):
+        for j in range(i + 1, D):
+            gg = np.eye(D).astype(int)
+            gg[i, i] = 0
+            gg[j, j] = 0
+            gg[i, j] = -1
+            gg[j, i] = 1
+            generators.append(gg)
+
+    return np.array(generators)
+
+def make_all_operators(D):
+    generators = make_all_generators(D)
+    operators = np.array([np.eye(D).astype(int), ])
+    foo = 0
+    while len(operators) != foo:
+        foo = len(operators)
+        operators = make_new_operators(operators, generators)
+    return(operators)
+
+def make_new_operators(operators, generators):
+    for op in operators:
+        for gg in generators:
+            op2 = (gg @ op).astype(int)
+            operators = np.unique(np.append(operators, op2[None, :, :], axis=0), axis=0)
+    return operators
+
+def test_group(operators):
+    D = len(operators[0])
+    # Check that the list of group operators is closed
+    for gg in operators:
+        for gg2 in operators:
+            if ((gg @ gg2).astype(int) not in operators):
+                return False
+    print("group is closed under multiplication")
+    # Check that gg.T is gg.inv for all gg in group?
+    for gg in operators:
+        if not np.allclose(gg @ gg.T, np.eye(D)):
+            return False
+    print("group operators are the transposes of their inverses")
+    return True
+
+# ------------------------------------------------------------------------------
+# PART 2: Define a k-tensor.
+
+class ktensor:
+
+    def __init__(self, data, parity, D):
+        self.D = D
+        assert self.D > 1, \
+        "ktensor: geometry makes no sense if D<2."
+        self.parity = parity
+        assert np.abs(self.parity) == 1, \
+        "ktensor: parity must be 1 or -1."
+        if len(np.atleast_1d(data)) == 1:
[…]

### 2022-07-27 ipynb/geometric.py
move code to py file; start making figures for the paper

@@ -1,4 +1,10 @@
 import numpy as np
+import pylab as plt
+import matplotlib.cm as cm
+from matplotlib.colors import ListedColormap
+import itertools as it
+
+TINY = 1.e-5
 
 # ------------------------------------------------------------------------------
 # PART 1: Make and test a complete group
@@ -151,7 +157,7 @@ class geometric_filter:
 
     def make_pixels_and_keys(self):
         foo = range(-self.m, self.m + 1)
-        self._pixels = np.array([pp for pp in it.product(foo, repeat=D)]).astype(int)
+        self._pixels = np.array([pp for pp in it.product(foo, repeat=self.D)]).astype(int)
         self._keys = [self.hash(pp) for pp in self._pixels]
         return
 
@@ -265,7 +271,7 @@ def setup_plot():
     fig = plt.figure(figsize=FIGSIZE)
     return fig
 
-def finish_plot(title, pixels):
+def finish_plot(title, pixels, D):
     plt.title(title)
     if D == 2:
         plt.xlim(np.min(pixels)-0.5, np.max(pixels)+0.5)
@@ -295,6 +301,8 @@ def fill_boxes(xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
 def plot_scalars(xs, ys, ws, boxes=True, fill=True, symbols=True,
                  vmin=0., vmax=5., cmap=cmap, colorbar=False,
                  norm_fill=False):
+    M = np.sqrt(len(xs)).astype(int)
+    assert len(xs) == M * M
     if boxes:
         plot_boxes(xs, ys)
     if fill:
@@ -325,7 +333,7 @@ def plot_scalar_filter(filter, title):
         elif filter.D == 3:
             xs[i], ys[i] = pp[0] + XOFF * pp[2], pp[1] + XOFF * pp[2]
     plot_scalars(xs, ys, ws, norm_fill=True)
-    finish_plot(title, filter.pixels())
+    finish_plot(title, filter.pixels(), filter.D)
 
 def plot_vectors(xs, ys, ws, boxes=True, fill=True,
                  vmin=0., vmax=10., cmap=cmap, scaling=0.33):
@@ -355,12 +363,12 @@ def plot_vector_filter(filter, title):
         elif filter.D == 3:
             xs[i], ys[i] = pp[0] + XOFF * pp[2], pp[1] + YOFF * pp[2]
     plot_vectors(xs, ys, ws)
[…]

### 2022-07-27 paper/cosmology.tex

--- 2022-07-18 paper/main.tex
@@
 %   -- Understand how this formulation compares with steerable CNNs
-%   -- What is known about what functions CNNs can represent? Ask Joan!
-%  -- Define "local" functions
+%   -- What is known about what functions CNNs can represent? all linear translation equivariant functions can be written as convvolutions if the filters are as large as the image (simple observation)
+% what about non-linear? is there a polynomial basis here? 
+%  -- Define "local" functions, is there a characterization of functions that can be written as convolutions with small filters?
 
@@
 \pagestyle{myheadings}
-\markboth{foo}{Hogg \& Villar / Emulating n-body simulations with equivariant CNNs}
+\markboth{foo}{Hogg \& Villar /Geometric filters for geometric images}
 \newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
@@
 Equivariant convolution-based functions
-of images or lattices
-containing scalars, vectors, and tensors}
+on tensor fields}
 \author{Hogg, Villar, Blum-Smith, others}
@@
 
-\paragraph{Abstract:} HERE IS OUR ABSTRACT.
+\paragraph{Abstract:} We provide a construction for general equivariant convolution-based functions of images or lattices
+containing scalars, vectors, and tensors.
+The formulation, inspired by Einstein summation notation, does not use point-wise activation functions, but it can universally approximate all analytic equivariant functions with respect to discrete rotations of the image.
+CHECK WITH BEN
 
@@
 
-\begin{definition}[Group $B_d$ of symmetries of a $d$- hypercube]
+\begin{definition}[Group $B_d$ of symmetries of a $d$-hypercube]
 We denote by $B_d$ the group of rotation and reflection symmetries of the $d$-dimensional hypercube.

### 2022-07-27 paper/main.tex
content

@@
 
-\section{Problem setup}\label{sec:setup}
-
-For our purposes, a cosmological simulation is a set of $N^3$ particles $n$ with $1\leq n\leq N^3$, each of which has a (scalar) mass $m_n$, a (vector) position $q_n$, and a (vector) veloctiy $v_n$, and each of which is a function of time (or computed or stored at a set of times $t$).
-The positions $q_n$ live in a cubical 3-torus of side length $L$.
-In the dark-matter-only case, the simulation dynamics depend only on the masses, positions, velocities $\{m_n, q_n, v_n\}_{n=1}^N$, time $t$, Newton's constant $G$, and (hopefully weakly) the number $N^3$ and length $L$.
-
-Because the simulations are designed to describe the evolution of the Universe, the particles begin in a nearly-uniform density state, usually very close to a uniform, torus-filling, cubical lattice.
-The particles begin at a very early time $t$ with tiny (vector) offsets $\Delta q_n$ away from lattice sites $q^{(0)}_n$ such that $q_n = q^{(0)}_n + \Delta q_n$, and tiny (vector) velocities $v_n$.
-In detail these positional offsets and velocities are closely related by something like a Zeldovich approximation.
-
-The uniform, torus-filling, cubical lattice of $N^3$ points $q^{(0)}_n$ can be thought of as a 3-dimensional ``image''.
-Each particle can be thought of as a ``pixel'' in that image.
-At each snapshot of the simulation at time $t$, each pixel has an associated (vector) positional offset $\Delta q_n$, (vector) velocity $v_n$, and a few scalars ($m_n, t, G$ at least).
-In addition to these native scalars, we can produce at each pixel $n$ three additional scalars $q_n^\top q_n, q_n^\top v_n, v_n^\top v_n$, which are the inner products of the vectors with themselves and each other.
-At late times, the simulation shows almost no memory of the initial uniform, torus-filling, cubical lattice.
-However, each particle has an identity $n$ that can be associated with its original lattice location $q^{(0)}_n$, so its properties can be thought of as being encoded in that original lattice-location pixel.
-That is, this is a problem of scalars and vectors on a uniform, cubical lattice in a cubical 3-torus.
-
-There are some exact and approximate symmetries to consider in this problem.
-The approximate symmetries are the fundamental symmetries of classical physics, which include translation, rotation, and parity.
-These symmetries are only approximate in the n-body simulation context, because there is a cubical 3-torus with an orientation and a size, and there is a cubical lattice inside that 3-torus.
-Generators of the exact symmetries are the three torus-axis-aligned spatial rotations by $\pi/2$ and one reflection across one of the three axes of the torus.
-
 \section{Geometric convolutions}\label{sec:convolution}
@@
 We consider the orthogonal group $O(d)$ and $g\in O(d)$. For our purposes $d$ will be 2 or 3, but everything generalizes.  
-We consider an image $A$ in a cubical $d$-torus with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
+We consider an image $A$ in a $d$-torus with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
 Each pixel can contain tensors and pseudotensors (of which scalars, pseudoscalars, vectors, and pseudovectors are particular cases) that we define below.
@@
 
-\begin{definition}[vectors, $k$-tensors, and the world]
+\begin{definition}[$O(d)$ action on vectors and $k$-tensors] \label{def.tensors}
 Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
-It also acts on lists of vectors $(\mathbb R^d)^n$ by acting on each of the $n$ vectors by the same action.
 It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
-Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
-Note that $O(d)$ can be seen as acting on the world.
+
+It also acts on lists of $k$-tensors $((\mathbb R^d)^{\otimes k})^n$ by acting on each of the $n$ $k$-tensors by the same action.
+%Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
+%Note that $O(d)$ can be seen as acting on the world.
 \end{definition}
@@
 
-\begin{definition}[pseudovector and $k$-pseudotensor]
-SOLE CAN WE SAY HERE that these are ``like a vector and like a tensor'' but there is an additonal $\det(g)$ coming in?
-\end{definition}
-
-\begin{definition}[scalar and pseudoscalar]
+\begin{definition}[scalars and pseudoscalars]
 We say that $a\in \mathbb R$ is a $0$-tensor or scalar iff $a$ is an invariant of the action.
@@
 where $\det(g)$ is the determinant of the matrix representation $M(g)$ of $g$.
+\end{definition}
+
+\begin{definition}[pseudovectors and $k$-pseudotensors]
+We say that $T\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot T = \det(g)\, g \cdot T$ where $g\cdot T$ is the cannonical action on tensors as in Definition \ref{def.tensors}.
 \end{definition}

### 2022-07-28 ipynb/geometric.py
plotting multiple filters per plot

@@ -1,3 +1,17 @@
+"""
+# Core code for GeometricConvolutions
+
+## License:
+TBD
+
+## Authors:
+- David W. Hogg (NYU)
+
+## To-do items:
+- Need to implement index permutation operation.
+- Need to implement Levi-Civita contraction.
+"""
+
 import numpy as np
 import pylab as plt
 import matplotlib.cm as cm
@@ -271,58 +285,58 @@ def setup_plot():
     fig = plt.figure(figsize=FIGSIZE)
     return fig
 
-def finish_plot(title, pixels, D):
-    plt.title(title)
+def finish_plot(ax, title, pixels, D):
+    ax.set_title(title)
     if D == 2:
-        plt.xlim(np.min(pixels)-0.5, np.max(pixels)+0.5)
-        plt.ylim(np.min(pixels)-0.5, np.max(pixels)+0.5)
+        ax.set_xlim(np.min(pixels)-0.5, np.max(pixels)+0.5)
+        ax.set_ylim(np.min(pixels)-0.5, np.max(pixels)+0.5)
     if D == 3:
-        plt.xlim(np.min(pixels)-0.75, np.max(pixels)+0.75)
-        plt.ylim(np.min(pixels)-0.75, np.max(pixels)+0.75)
-    plt.gca().set_aspect("equal")
-    plt.gca().set_xticks([])
-    plt.gca().set_yticks([])
+        ax.set_xlim(np.min(pixels)-0.75, np.max(pixels)+0.75)
+        ax.set_ylim(np.min(pixels)-0.75, np.max(pixels)+0.75)
+    ax.set_aspect("equal")
+    ax.set_xticks([])
+    ax.set_yticks([])
 
-def plot_boxes(xs, ys):
+def plot_boxes(ax, xs, ys):
     for x, y in zip(xs, ys):
-        plt.plot([x-0.5, x-0.5, x+0.5, x+0.5, x-0.5],
+        ax.plot([x-0.5, x-0.5, x+0.5, x+0.5, x-0.5],
                  [y-0.5, y+0.5, y+0.5, y-0.5, y-0.5], "k-", lw=0.5)
 
-def fill_boxes(xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
+def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
     cmx = cm.ScalarMappable(cmap=cmap)
[…]

### 2022-07-28 paper/main.tex

@@
 \pagestyle{myheadings}
-\markboth{foo}{Hogg \& Villar /Geometric filters for geometric images}
+\markboth{foo}{Hogg, Blum-Smith, \& Villar /Geometric filters for geometric images}
 \newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
@@
 on tensor fields}
-\author{Hogg, Villar, Blum-Smith, others}
-\date{2022 June}
+\author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
+\date{}
 
@@
 
-\begin{definition}[$O(d)$ action on vectors and $k$-tensors] \label{def.tensors}
-Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
-It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
-
-It also acts on lists of $k$-tensors $((\mathbb R^d)^{\otimes k})^n$ by acting on each of the $n$ $k$-tensors by the same action.
-%Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
-%Note that $O(d)$ can be seen as acting on the world.
+\begin{definition}[vectors, scalars, and $k$-tensors] \label{def.tensors}
+We say that $v\in \mathbb R^d$ is a vector (or 1-tensor) if $O(d)$ acts on $v$ via the canonical action.
+Namely, for any $g\in O(d)$ we have $g\cdot v = M(g)\, v$ where $M(g)\in \mathbb R^{d\times d}$ is the the matrix representation of $g$.
+We say that $a\in \mathbb R$ is a scalar (or 0-tensor) if $O(d)$ acts trivially on $a$.
+Namely $g\cdot a = a$.
+We say that $T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor if $O(d)$ acts on $T$ as follows:
+$O(d)$ acts on a rank-1 $k$-tensor by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and it acts on higher rank tensors by linearly extending the actions to linear combinations of such rank-1 $k$-tensors.
 \end{definition}
@@
 
-\begin{definition}[scalars and pseudoscalars]
-We say that $a\in \mathbb R$ is a $0$-tensor or scalar iff $a$ is an invariant of the action.
-Namely, $a(W) = a(g\cdot W)$ for all $g\in O(d)$.
-We say that
-$a\in \mathbb R$ is a $0$-pseudotensor or pseudoscalar iff $a(g \cdot W) = \det(g)\,a(W)$ for all $g\in O(d)$,
-where $\det(g)$ is the determinant of the matrix representation $M(g)$ of $g$.
-\end{definition}
-
-\begin{definition}[pseudovectors and $k$-pseudotensors]
-We say that $T\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot T = \det(g)\, g \cdot T$ where $g\cdot T$ is the cannonical action on tensors as in Definition \ref{def.tensors}.
+\begin{definition}[pseudovectors, pseudoscalars, and $k$-pseudotensors]
+We say that $B\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot B := \det(g)\, g \cdot B$ where $g\cdot B$ is the canonical action on $k$-tensors as in Definition \ref{def.tensors}.
+The pseudovectors and pseudoscalars are the 1-pseudotensors and 0-pseudotensors respectively.
 \end{definition}
@@
 \begin{definition}[permutations of tensor indices]
-MSV WRITE THIS.
+Given a $k$-tensor or $k$-pseudotensor $T\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $T^\sigma$ the result of permuting the indices of $T$ according to $\sigma$:
+\begin{equation}
+    [T^\sigma]_{i_1, \ldots, i_k} := [T]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
+\end{equation}
 \end{definition}
@@
 
+\section{Expressive power of this architecture}
+We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).
+
+We can cite that convolutions with small patches may be seen as low pass filters in some way. Reference?
+
 \section{Construction of equivariant geometric convolution filters}
-
 In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.

### 2022-07-28 ipynb/geometric.py
arrows, details

@@ -312,7 +312,7 @@ def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
         ax.fill_between([x - 0.5, x + 0.5], [y - 0.5, y - 0.5], [y + 0.5, y + 0.5],
                              color=c, zorder=zorder)
 
-def plot_scalars(ax, M, D, xs, ys, ws, boxes=True, fill=True, symbols=True,
+def plot_scalars(ax, M, xs, ys, ws, boxes=True, fill=True, symbols=True,
                  vmin=0., vmax=5., cmap=cmap, colorbar=False,
                  norm_fill=False):
     if boxes:
@@ -346,7 +346,7 @@ def plot_scalar_filter(filter, title, ax=None):
             xs[i], ys[i] = pp
         elif filter.D == 3:
             xs[i], ys[i] = pp[0] + XOFF * pp[2], pp[1] + XOFF * pp[2]
-    plot_scalars(ax, filter.M, filter.D, xs, ys, ws, norm_fill=True)
+    plot_scalars(ax, filter.M, xs, ys, ws, norm_fill=True)
     finish_plot(ax, title, filter.pixels(), filter.D)
 
 def plot_vectors(ax, xs, ys, ws, boxes=True, fill=True,
@@ -356,10 +356,14 @@ def plot_vectors(ax, xs, ys, ws, boxes=True, fill=True,
     if fill:
         fill_boxes(ax, xs, ys, np.sum(np.abs(ws), axis=-1), vmin, vmax, cmap)
     for x, y, w in zip(xs, ys, ws):
-        if np.sum(w * w) > TINY:
+        normw = np.linalg.norm(w)
+        if normw > TINY:
             ax.arrow(x - scaling * w[0], y - scaling * w[1],
                      2 * scaling * w[0], 2 * scaling * w[1],
-                     length_includes_head=True, head_width=0.1, color="k")
+                     length_includes_head=True,
+                     head_width= 0.24 * scaling * normw,
+                     head_length=0.72 * scaling * normw,
+                     color="k")
 
 def plot_vector_filter(filter, title, ax=None):
     assert filter.k == 1
@@ -396,7 +400,8 @@ def plot_filters(filters, name):
             plot_scalar_filter(ff, "{} {}".format(name, i), ax=ax)
         if ff.k == 1:
             plot_vector_filter(ff, "{} {}".format(name, i), ax=ax)
-    return
+    return fig
+
[…]

### 2022-07-28 notebooks/geometric.py
notes to self

@@ -8,8 +8,10 @@ TBD
 - David W. Hogg (NYU)
 
 ## To-do items:
+- Fix sizing of multi-filter plots.
 - Need to implement index permutation operation.
 - Need to implement Levi-Civita contraction.
+- Make all monomial images.
 """
 
 import numpy as np

### 2022-07-29 notebooks/geometric.py
renaming files; plotting improvements

@@ -81,6 +81,18 @@ def test_group(operators):
 
 class ktensor:
 
+    def name(k, parity):
+        nn = "tensor"
+        if k == 0:
+            nn = "scalar"
+        if k == 1:
+            nn = "vector"
+        if parity < 0:
+            nn = "pseudo" + nn
+        if k > 1:
+            nn = "${}$-".format(k) + nn
+        return nn
+
     def __init__(self, data, parity, D):
         self.D = D
         assert self.D > 1, \
@@ -287,22 +299,29 @@ def setup_plot():
     fig = plt.figure(figsize=FIGSIZE)
     return fig
 
+def nobox(ax):
+    # ax.set_xticks([])
+    # ax.set_yticks([])
+    ax.axis("off")
+    return
+
 def finish_plot(ax, title, pixels, D):
     ax.set_title(title)
     if D == 2:
-        ax.set_xlim(np.min(pixels)-0.5, np.max(pixels)+0.5)
-        ax.set_ylim(np.min(pixels)-0.5, np.max(pixels)+0.5)
+        ax.set_xlim(np.min(pixels)-0.55, np.max(pixels)+0.55)
+        ax.set_ylim(np.min(pixels)-0.55, np.max(pixels)+0.55)
     if D == 3:
         ax.set_xlim(np.min(pixels)-0.75, np.max(pixels)+0.75)
         ax.set_ylim(np.min(pixels)-0.75, np.max(pixels)+0.75)
     ax.set_aspect("equal")
-    ax.set_xticks([])
-    ax.set_yticks([])
+    nobox(ax)
+    return
 
 def plot_boxes(ax, xs, ys):
     for x, y in zip(xs, ys):
         ax.plot([x-0.5, x-0.5, x+0.5, x+0.5, x-0.5],
-                 [y-0.5, y+0.5, y+0.5, y-0.5, y-0.5], "k-", lw=0.5)
+                 [y-0.5, y+0.5, y+0.5, y-0.5, y-0.5], "k-", lw=0.5, zorder=10)
+    return
 
 def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
     cmx = cm.ScalarMappable(cmap=cmap)
@@ -313,6 +332,7 @@ def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
     for x, y, c in zip(xs, ys, cs):
[…]

### 2022-07-29 notebooks/geometric.py
tiny tweaks to figure placement; still sux

@@ -423,13 +423,13 @@ def plot_no_filter(ax):
 def plot_filters(filters, names, n):
     assert len(filters) <= n
     bar = 10. # figure width in inches?
-    fig, axes = plt.subplots(1, n, figsize = (bar, 1.1 * bar / n))
+    fig, axes = plt.subplots(1, n, figsize = (bar, 0.1 + bar / n)) # magic
     if n == 1:
         axes = [axes, ]
     bar = 0.2
     foo = bar / n # fractional?
     plt.subplots_adjust(left=foo/2, right=1-foo/2, wspace=foo,
-                        bottom=bar/2, top=1-bar/2, hspace=bar)
+                        bottom=foo)
     for i, (ff, name) in enumerate(zip(filters, names)):
         if ff.k == 0:
             plot_scalar_filter(ff, name, ax=axes[i])

### 2022-07-30 notebooks/geometric.py
tweaks in preparation for making image figures

@@ -617,7 +617,7 @@ def setup_image_plot():
     ff = plt.figure(figsize=(8, 6))
     return ff
 
-def plot_scalar_image(image, vmin=None, vmax=None, ax=None):
+def plot_scalar_image(image, vmin=None, vmax=None, ax=None, colorbar=False):
     assert image.D == 2
     assert image.k == 0
     if ax is None:
@@ -631,7 +631,7 @@ def plot_scalar_image(image, vmin=None, vmax=None, ax=None):
     if vmax is None:
         vmax = np.percentile(plotdata[:, 2], 97.5)
     plot_scalars(ax, image.N, plotdata[:, 0], plotdata[:, 1], plotdata[:, 2],
-                 symbols=False, vmin=vmin, vmax=vmax, colorbar=True)
+                 symbols=False, vmin=vmin, vmax=vmax, colorbar=colorbar)
     image_axis(ax, plotdata)
     return ax
 

### 2022-07-30 notebooks/geometric.py
make all first- and second-order monomials

@@ -378,11 +378,15 @@ def plot_scalar_filter(filter, title, ax=None):
     return ax
 
 def plot_vectors(ax, xs, ys, ws, boxes=True, fill=True,
-                 vmin=0., vmax=10., cmap=cmap, scaling=0.33):
+                 vmin=0., vmax=10., cmap=cmap, scaling=None):
     if boxes:
         plot_boxes(ax, xs, ys)
     if fill:
-        fill_boxes(ax, xs, ys, np.sum(np.abs(ws), axis=-1), vmin, vmax, cmap)
+        fill_boxes(ax, xs, ys, np.linalg.norm(ws, axis=-1), vmin, vmax, cmap)
+    if scaling is None:
+        nws = np.linalg.norm(ws, axis=-1)
+        I = nws > TINY
+        scaling = 0.33 / np.median(nws[I])
     for x, y, w in zip(xs, ys, ws):
         normw = np.linalg.norm(w)
         if normw > TINY:
@@ -655,3 +659,9 @@ def plot_vector_image(image, ax=None):
     image_axis(ax, plotdata)
     return ax
 
+def plot_image(image, **kwargs):
+    assert image.D == 2
+    if image.k == 0:
+        plot_scalar_image(image, **kwargs)
+    if image.k == 1:
+        plot_vector_image(image, **kwargs)

### 2022-07-30 notebooks/geometric.py
vmax

@@ -378,7 +378,7 @@ def plot_scalar_filter(filter, title, ax=None):
     return ax
 
 def plot_vectors(ax, xs, ys, ws, boxes=True, fill=True,
-                 vmin=0., vmax=10., cmap=cmap, scaling=None):
+                 vmin=0., vmax=5., cmap=cmap, scaling=None):
     if boxes:
         plot_boxes(ax, xs, ys)
     if fill:

### 2022-07-30 notebooks/geometric.py
m

@@ -229,8 +229,8 @@ class geometric_filter:
         return newfilter
 
     def __str__(self):
-        return "<geometric filter object in D={} with M={}, k={}, and parity={}>".format(
-            self.D, self.M, self.k, self.parity)
+        return "<geometric filter object in D={} with M={} (m={}), k={}, and parity={}>".format(
+            self.D, self.M, self.m, self.k, self.parity)
 
     def times_group_element(self, gg):
         newfilter = self.copy()

### 2022-08-01 notebooks/geometric.py
colorful figures

@@ -18,6 +18,7 @@ import numpy as np
 import pylab as plt
 import matplotlib.cm as cm
 from matplotlib.colors import ListedColormap
+import cmastro as cma
 import itertools as it
 
 TINY = 1.e-5
@@ -288,11 +289,15 @@ class geometric_filter:
 FIGSIZE = (4, 3)
 XOFF, YOFF = 0.15, -0.1
 
-def make_colormap():
-    foo = np.linspace(0., 0.5, 256)
-    cmap = cm.get_cmap("gray_r")
-    colors = [cmap(f) for f in foo]
-    return ListedColormap(colors)
+def make_colormap(unph=True):
+    if unph:
+        return cm.get_cmap("cma:unph")
+    else:
+        foo = np.linspace(0., 0.5, 256)
+        cmap = cm.get_cmap("gray_r")
+        colors = [cmap(f) for f in foo]
+        return ListedColormap(colors)
+    return
 cmap = make_colormap()
 
 def setup_plot():
@@ -335,16 +340,11 @@ def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
     return
 
 def plot_scalars(ax, M, xs, ys, ws, boxes=True, fill=True, symbols=True,
-                 vmin=0., vmax=5., cmap=cmap, colorbar=False,
-                 norm_fill=False):
+                 vmin=-2., vmax=2., cmap=cmap, colorbar=False):
     if boxes:
         plot_boxes(ax, xs, ys)
     if fill:
-        if norm_fill:
-            nws = np.abs(ws)
-        else:
-            nws = ws
-        fill_boxes(ax, xs, ys, nws, vmin, vmax, cmap, colorbar=colorbar)
+        fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, colorbar=colorbar)
     if symbols:
         height = ax.get_window_extent().height
         ss = (5 * height / M) * np.abs(ws)
@@ -373,20 +373,16 @@ def plot_scalar_filter(filter, title, ax=None):
             xs[i], ys[i] = pp
         elif filter.D == 3:
             xs[i], ys[i] = pp[0] + XOFF * pp[2], pp[1] + XOFF * pp[2]
-    plot_scalars(ax, filter.M, xs, ys, ws, norm_fill=True)
+    plot_scalars(ax, filter.M, xs, ys, ws)
     finish_plot(ax, title, filter.pixels(), filter.D)
     return ax
 
[…]

### 2022-08-01 paper/main.tex
Updates from Overleaf
incorporate new figures

@@
 % --- generalize convolution to rectangular non-toroidal images
+
+%-- is it the same to do convolutions + products + convolutions + products
+
+%-- a counting argument may tell us that all the linear functions that are equivariant  (Molien's formula)
+
+%-- BEN: compute Molien's formula for this (translations semidirect D_4) action
+% 1/|G|\sum_{g\in G} det(I-\phi(g)t)^{-1}
 
@@
 
-Show all the filters for $d=2$, $M=3$.
+\begin{figure}[tp]
+\includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter+1_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
+\caption{All the filters for $d=2$, $m=1$. HOGG: notes!}
+\end{figure}
 
-Show all the filters for $d=2$, $M=5$.
-
-Show all the filters for $d=3$, $M=3$.
+\begin{figure}[tp]
+\includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-0_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter+1_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
+\caption{All the filters for $d=2$, $m=2$. HOGG: notes!}
+\end{figure}
 
@@
 
+\begin{figure}
+    \includegraphics[width=\textwidth]{notebooks/monomials_1.png}
+    \caption{A scalar image $s$ and convolutions with four chosen filters.}
+\end{figure}
+
 Now show contractions of products of the above on those images.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{notebooks/monomials_2.png}
+    \caption{All unconvolved second-degree monomials of $s$, given four chosen filters. For the 2-tensor outputs (order $k=2$), we traced them back to scalars.}
+\end{figure}
 
@@
 
+\paragraph{Acknowlegements:}
+This project made use of Python, numpy, matplotlib, and cmastro.
+
 \end{document}

### 2022-08-01 notebooks/geometric.py
tweak colors

@@ -289,17 +289,6 @@ class geometric_filter:
 FIGSIZE = (4, 3)
 XOFF, YOFF = 0.15, -0.1
 
-def make_colormap(unph=True):
-    if unph:
-        return cm.get_cmap("cma:unph")
-    else:
-        foo = np.linspace(0., 0.5, 256)
-        cmap = cm.get_cmap("gray_r")
-        colors = [cmap(f) for f in foo]
-        return ListedColormap(colors)
-    return
-cmap = make_colormap()
-
 def setup_plot():
     fig = plt.figure(figsize=FIGSIZE)
     return fig
@@ -328,19 +317,19 @@ def plot_boxes(ax, xs, ys):
                  [y-0.5, y+0.5, y+0.5, y-0.5, y-0.5], "k-", lw=0.5, zorder=10)
     return
 
-def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False):
-    cmx = cm.ScalarMappable(cmap=cmap)
+def fill_boxes(ax, xs, ys, ws, vmin, vmax, cmap, zorder=-100, colorbar=False, alpha=1.):
+    cmx = cm.ScalarMappable(cmap=cm.get_cmap(cmap))
     cmx.set_clim(vmin, vmax)
     cs = cmx.to_rgba(ws)
     if colorbar:
         plt.colorbar(cmx, ax=ax)
     for x, y, c in zip(xs, ys, cs):
         ax.fill_between([x - 0.5, x + 0.5], [y - 0.5, y - 0.5], [y + 0.5, y + 0.5],
-                             color=c, zorder=zorder)
+                             color=c, zorder=zorder, alpha=alpha)
     return
 
 def plot_scalars(ax, M, xs, ys, ws, boxes=True, fill=True, symbols=True,
-                 vmin=-2., vmax=2., cmap=cmap, colorbar=False):
+                 vmin=-2., vmax=2., cmap="cma:unph", colorbar=False):
     if boxes:
         plot_boxes(ax, xs, ys)
     if fill:
@@ -378,11 +367,11 @@ def plot_scalar_filter(filter, title, ax=None):
     return ax
 
 def plot_vectors(ax, xs, ys, ws, boxes=True, fill=True,
-                 vmin=-2., vmax=2., cmap=cmap, scaling=0.33):
[…]

### 2022-08-01 notebooks/geometric.py
tensors; colors

@@ -404,6 +404,47 @@ def plot_vector_filter(filter, title, ax=None):
     finish_plot(ax, title, filter.pixels(), filter.D)
     return ax
 
+Rx = np.array([[-1.0, -1.0, 1.0, 1.0, -1.0, 0.0,  1.0],
+               [-1.5,  1.5, 1.5, 0.2, -0.2, 0.0, -1.5]])
+
+def plot_one_tensor(ax, x, y, T, color="k", zorder=0, scaling=0.2):
+    dx, dy = scaling * T @ Rx
+    ax.plot(x + dx, y + dy, ls="-", color=color)
+    return
+
+def plot_tensors(ax, xs, ys, ws, boxes=True, fill=True,
+                 vmin=0., vmax=2., cmap="cma:hesperia_r", scaling=0.33):
+    if boxes:
+        plot_boxes(ax, xs, ys)
+    if fill:
+        fill_boxes(ax, xs, ys, np.linalg.norm(ws, axis=-1), vmin, vmax, cmap, alpha=0.25)
+    for x, y, w in zip(xs, ys, ws):
+        normw = np.linalg.norm(w)
+        if normw > TINY:
+            plot_one_tensor(ax, x, y, w, color="k", zorder=100)
+    return
+
+def plot_tensor_filter(filter, title, ax=None):
+    assert filter.k == 2, "plot_tensor_filter(): Only 2-tensors (for now)."
+    if filter.D not in [2, ]:
+        print("plot_vector_filter(): Only works for D in [2, ].")
+        return
+    if ax is None:
+        fig = setup_plot()
+        ax = fig.gca()
+    MtotheD = filter.M ** filter.D
+    xs, ys = np.zeros(MtotheD), np.zeros(MtotheD)
+    ws = np.zeros((MtotheD, filter.D, filter.D))
+    for i, (kk, pp) in enumerate(zip(filter.keys(), filter.pixels())):
+        ws[i] = filter[kk].data
+        if filter.D == 2:
+            xs[i], ys[i] = pp
+        elif filter.D == 3:
+            xs[i], ys[i] = pp[0] + XOFF * pp[2], pp[1] + YOFF * pp[2]
+    plot_tensors(ax, xs, ys, ws, vmin=0., vmax=3.)
+    finish_plot(ax, title, filter.pixels(), filter.D)
+    return ax
+
 def plot_nothing(ax):
     ax.set_title(" ")
     nobox(ax)
[…]

### 2022-08-01 notebooks/geometric.py
changed norm, made everything; other tweaks

@@ -11,7 +11,7 @@ TBD
 - Fix sizing of multi-filter plots.
 - Need to implement index permutation operation.
 - Need to implement Levi-Civita contraction.
-- Make all monomial images.
+- Need to implement bin-down and bin-up operators.
 """
 
 import numpy as np
@@ -134,14 +134,17 @@ class ktensor:
     def norm(self):
         if self.k == 0:
             return np.abs(self.data)
-        return np.linalg.norm(self.data, ord=2)
+        return np.linalg.norm(self.data)
 
     def times_scalar(self, scalar):
         return ktensor(scalar * self.data, self.parity, self.D)
 
     def times_group_element(self, gg):
-        # BUG: THIS IS UNTESTED.
-        # BUG: This is incomprehensible.
+        """
+        # Notes / Bugs:
+        - THIS IS UNTESTED.
+        - This is incomprehensible.
+        """
         assert self.k < 14
         assert gg.shape == (self.D, self.D)
         sign, logdet = np.linalg.slogdet(gg)
@@ -701,7 +704,7 @@ def plot_images(images):
       a LaTeX expression.
     """
     nim = len(images)
-    n = np.ceil(np.sqrt(nim)).astype(int)
+    n = np.floor(np.sqrt(nim)).astype(int)
     m = np.ceil(nim / n).astype(int)
     print(len(images), n, m)
     bar = 10. # inches?

### 2022-08-02 notebooks/geometric.py
tweaks, unit tests of group actions

@@ -2,12 +2,15 @@
 # Core code for GeometricConvolutions
 
 ## License:
-TBD
+Copyright 2022 David W. Hogg and contributors.
+The code in GeometricConvolutions is licensed under the open-source MIT License.
+See the file `LICENSE` for more details.
 
 ## Authors:
 - David W. Hogg (NYU)
 
 ## To-do items:
+- Create tests for group operations on k-tensors.
 - Fix sizing of multi-filter plots.
 - Need to implement index permutation operation.
 - Need to implement Levi-Civita contraction.
@@ -171,6 +174,53 @@ class ktensor:
         letters  = "bcdefghijklmnopqrstuvwxyz"
         einstr = letters[:i] + "a" + letters[i:j-1] + "a" + letters[j-1:self.k-2]
         return ktensor(np.einsum(einstr, self.data), self.parity, self.D)
+
+# Now test group actions on k-tensors:
+def test_group_actions(operators):
+    """
+    # Notes:
+    - This only does minimal tests!
+    """
+    D = len(operators[0])
+
+    for parity in [-1, 1]:
+
+        # vector dot vector
+        v1 = ktensor(np.random.normal(size=D), parity, D)
+        v2 = ktensor(np.random.normal(size=D), parity, D)
+        dots = [(v1.times_group_element(gg)
+                 * v2.times_group_element(gg)).contract(0, 1).data
+                for gg in operators]
+        dots = np.array(dots)
+        if not np.allclose(dots, np.mean(dots)):
+            print("failed (parity = {}) vector dot test.".format(parity))
+            return False
+        print("passed (parity = {}) vector dot test.".format(parity))
+
+        # tensor times tensor
+        T3 = ktensor(np.random.normal(size=(D, D)), parity, D)
+        T4 = ktensor(np.random.normal(size=(D, D)), parity, D)
+        dots = [(T3.times_group_element(gg)
+                 * T4.times_group_element(gg)).contract(1, 2).contract(0, 1).data
+                for gg in operators]
+        dots = np.array(dots)
+        if not np.allclose(dots, np.mean(dots)):
+            print("failed (parity = {}) tensor times tensor test".format(parity))
+            return False
+        print("passed (parity = {}) tensor times tensor test".format(parity))
+
+        # vectors dotted through tensor
+        v5 = ktensor(np.random.normal(size=D), 1, D)
[…]

### 2022-08-02 notebooks/geometric.py
tweak plot layout

@@ -506,13 +506,11 @@ def plot_nothing(ax):
 def plot_filters(filters, names, n):
     assert len(filters) <= n
     bar = 10. # figure width in inches?
-    fig, axes = plt.subplots(1, n, figsize = (bar, 0.1 + bar / n)) # magic
+    fig, axes = plt.subplots(1, n, figsize = (bar, 0.03 * bar + bar / n)) # magic
     if n == 1:
         axes = [axes, ]
-    bar = 0.2
-    foo = bar / n # fractional?
-    plt.subplots_adjust(left=foo/2, right=1-foo/2, wspace=foo,
-                        bottom=foo)
+    plt.subplots_adjust(left=0.001, right=0.999, wspace=0.2/n,
+                        bottom=0.001, top=0.999-0.1, hspace=0)
     for i, (ff, name) in enumerate(zip(filters, names)):
         if ff.k == 0:
             plot_scalar_filter(ff, name, ax=axes[i])
@@ -758,12 +756,10 @@ def plot_images(images):
     m = np.ceil(nim / n).astype(int)
     print(len(images), n, m)
     bar = 10. # inches?
-    fig, axes = plt.subplots(m, n, figsize = (bar, 0.5 + 0.5 * m + bar * m / n)) # magic
+    fig, axes = plt.subplots(m, n, figsize = (bar, 0.2 * m + bar * m / n)) # magic
     axes = np.atleast_1d(axes).flatten()
-    bar = 0.2 / m
-    foo = 0.2 / n # fractional?
-    plt.subplots_adjust(left=foo/2, right=1-foo/2, wspace=foo,
-                        bottom=bar, hspace=bar)
+    plt.subplots_adjust(left=0.001, right=0.999, wspace=0.2/n,
+                        bottom=0.001, top=0.999-0.07/m, hspace=0.2/m)
     for ax, (image, latex) in zip(axes, images):
         plot_image(image, ax=ax)
         finish_plot(ax, "$" + latex + "$", image.pixels(), image.D)

### 2022-08-17 notebooks/geometric.py
tweak

@@ -677,9 +677,6 @@ class geometric_image:
         assert self.N == other.N
         newk, newparity = self.k + other.k, self.parity * other.parity
         newimage = geometric_image.zeros(self.N, newk, newparity, self.D)
-        assert newimage.D == self.D
-        assert newimage.N == self.N
-        assert newimage.k == newk
         for kk in self.keys():
             newimage[kk] = self[kk] * other[kk] # handled by ktensor
         return newimage

### 2022-08-19 paper/main.tex

@@
 
+% write the levi civita contraction
+% write the procedure of which we make all linear functions from k tensor images to k' tensor images of the same size. 
+% do we include a pooling move?
+
+% how do we enumerate all tensor functions
+
+% 1, 6, 22, 49, 87 d=3, n=1,3,7,9 (11N^2-24N+21)/8
+
 \documentclass{article}
@@
 \usepackage{amsthm}
+
+% stuff for framing figures
+\usepackage[framemethod=tikz]{mdframed}
+\usetikzlibrary{shadows}
+\definecolor{captiongray}{HTML}{555555}
+\mdfsetup{%
+  innertopmargin=2ex,
+  innerbottommargin=1.8ex,
+  linecolor=captiongray,
+  linewidth=0.5pt,
+  roundcorner=5pt,
+  shadow=true,
+  shadowcolor=black!05,
+  shadowsize=4pt
+}
 
@@
 \pagestyle{myheadings}
-\markboth{foo}{Hogg, Blum-Smith, \& Villar /Geometric filters for geometric images}
+\markboth{foo}{Hogg, Blum-Smith, \& Villar / Geometric filters for geometric images}
 \newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
@@
 \title{\bfseries%
-Equivariant convolution-based functions
-on tensor fields}
+Convolution-based functions
+on tensor fields and tensor images}
 \author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
@@
 
-\paragraph{Abstract:} We provide a construction for general equivariant convolution-based functions of images or lattices
-containing scalars, vectors, and tensors.
-The formulation, inspired by Einstein summation notation, does not use point-wise activation functions, but it can universally approximate all analytic equivariant functions with respect to discrete rotations of the image.
-CHECK WITH BEN
+\paragraph{Abstract:}
+Convolutional neural networks are the main tools for machine learning with images at the present day.
+These methods effectively assume that the input images are scalar images (an intensity or brightness in each pixel, possibly in three or more channels), and they enforce a kind of locality and translation-equivariance.
+In natural-science domains, image-like data sets can have a vector (velocity, say), or tensor (polarization, say), or pseudovector (magnetic field, say), or any combination in each pixel.
+We provide a construction for general convolution-based functions of images or lattices
+containing scalars, vectors, and tensors, producing scalar, vector, and tensor outputs.
+Our formulation, which combines convolutions with products, tensor contractions, and tensor index permutations, does not require point-wise activation functions (RELU, say) but can universally approximate any suitably local analytic scalar, vector, or tensor functions on scalar, vector, or tensor images.
+Our method permits, with a very simple adjustment, restriction to function spaces that are exactly equivariant to translations, discrete rotations, and reflections.
 
@@
 
-Contemporary physics teems with data sets that are images or lattices or grids of geometric objects.
+Contemporary natural science teems with data sets that are images or lattices or grids of geometric objects.
 Rectangular images or lattices of density (a scalar field), temperature (also a scalar field), velocity (a vector field), magnetic field (a pseudo-vector field), and polarization (a 2-tensor field) are regularly taken with detectors or made in simulations.
-That is, physics has lots of data that consists of geometric objects on regular grids.
-These objects (scalars, vectors, pseudo-vectors, and tensors, for example) are \emph{geometric} in the sense that they are defined by their transformation properties under rotation and reflection.
+That is, there is plenty of data consisting of geometric objects on regular grids.
+These kinds of images (grids of scalars, vectors, pseudo-vectors, and tensors, for example) are \emph{geometric} in the sense that they contain (at every pixel) objects are defined by their transformation properties under rotation and reflection.
 For example, scalars are defined to be quantities in $\mathbb R$ that are invariant with respect to all rotations and reflections; vectors are defined (in $d$-dimensional space) to be quantities in $\mathbb R^d$ that transform under rotation or reflection by multiplication by the orthogonal $d\times d$ matrix representing the rotation or reflection; and so on.
 
-For one concrete example, cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
+For a concrete example close to one of our hearts (DWH's), contemporary cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
 For this reason, accurate, fast emulators might be of enormous economic and scientific value.
@@
 \begin{figure}[tp]
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
 \includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
@@
 \includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
+\includegraphics[width=\textwidth]{notebooks/filter+2_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-2_2_3.png}
+  \end{center}
 \caption{All the filters for $d=2$, $m=1$. HOGG: notes!}
+  \end{mdframed}
 \end{figure}
@@
 \begin{figure}[tp]
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
 \includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
@@
 \includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
+  \end{center}
 \caption{All the filters for $d=2$, $m=2$. HOGG: notes!}
+  \end{mdframed}
 \end{figure}
@@
 \begin{figure}
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
     \includegraphics[width=\textwidth]{notebooks/monomials_1.png}
+  \end{center}
     \caption{A scalar image $s$ and convolutions with four chosen filters.}
+  \end{mdframed}
 \end{figure}
@@
 \begin{figure}
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
     \includegraphics[width=\textwidth]{notebooks/monomials_2.png}
+  \end{center}
     \caption{All unconvolved second-degree monomials of $s$, given four chosen filters. For the 2-tensor outputs (order $k=2$), we traced them back to scalars.}
+  \end{mdframed}
 \end{figure}

### 2022-08-19 paper/main.tex
how to make all linear functions

@@
 \newtheorem{theorem}{Theorem}
+
+% math macros
+% HOGG SAY: I DON'T KNOW HOW TO USE AMSMATH \choose AND I'M ON A PLANE. FEEL FREE TO FIX THIS.
+\renewcommand{\choose}[2]{\begin{pmatrix}{#1}\\{#2}\end{pmatrix}}
 
@@
 
+SOLE: WHERE SHOULD THIS GO?
+Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY?
+IF NOT, IT SHOULD BE DEFINED.
+IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
+Here are the linear operations (linear in the input image) that are permitted):
+\begin{itemize}
+\item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
+  Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
+  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kroenecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kroenecker contractions.
+\item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
+  Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
+  Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
+  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
+  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kroenecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kroenecker contractions.
+(SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
+\end{itemize}
+
 \section{Nonlinear equivariant local geometric operators}\label{sec:nonlinear}

### 2022-08-23 notebooks/geometric.py
fix parity bit / input

@@ -91,7 +91,7 @@ class ktensor:
             nn = "scalar"
         if k == 1:
             nn = "vector"
-        if parity < 0:
+        if parity % 2 == 1:
             nn = "pseudo" + nn
         if k > 1:
             nn = "${}$-".format(k) + nn
@@ -102,9 +102,7 @@ class ktensor:
         self.D = D
         assert self.D > 1, \
         "ktensor: geometry makes no sense if D<2."
-        self.parity = parity
-        assert np.abs(self.parity) == 1, \
-        "ktensor: parity must be 1 or -1."
+        self.parity = parity % 2
         if len(np.atleast_1d(data)) == 1:
             self.data = data
             self.k = 0
@@ -127,9 +125,9 @@ class ktensor:
     def __mul__(self, other):
         if self.k == 0 or other.k == 0:
             return ktensor(self.data * other.data,
-                           self.parity * other.parity, self.D)
+                           self.parity + other.parity, self.D)
         return ktensor(np.multiply.outer(self.data, other.data),
-                       self.parity * other.parity, self.D)
+                       self.parity + other.parity, self.D)
 
     def __str__(self):
         return "<k-tensor object in D={} with k={} and parity={}>".format(
@@ -161,9 +159,7 @@ class ktensor:
             einstr = "".join([firstletters[i] for i in range(self.k)]) +"," + \
             ",".join([secondletters[i] + firstletters[i] for i in range(self.k)])
             foo = (self.data, ) + self.k * (gg, )
-            newdata = np.einsum(einstr, *foo)
-        if self.parity < 0:
-            newdata *= sign
+            newdata = np.einsum(einstr, *foo) * sign ** self.parity
         return ktensor(newdata, self.parity, self.D)
 
     def contract(self, i, j):
@@ -183,7 +179,7 @@ class ktensor:
             otherdata = np.zeros_like(self.data)
             otherdata[..., 0] =  1. * np.take(self.data, 1, axis=index)
             otherdata[..., 1] = -1. * np.take(self.data, 0, axis=index)
-            return ktensor(otherdata, self.parity * -1, self.D)
+            return ktensor(otherdata, self.parity + 1, self.D)
         if self.D == 3:
             assert len(index) == 2
             i, j = index
@@ -195,7 +191,7 @@ class ktensor:
                               - np.take(np.take(self.data, 2, axis=j), 0, axis=i)
[…]

### 2022-08-23 notebooks/geometric.py
bugfix on pseudoscalars

@@ -152,7 +152,7 @@ class ktensor:
         sign, logdet = np.linalg.slogdet(gg)
         assert logdet == 0.
         if self.k == 0:
-            newdata = 1. * self.data
+            newdata = 1. * self.data * sign ** self.parity
         else:
             firstletters  = "abcdefghijklm"
             secondletters = "nopqrstuvwxyz"

### 2022-08-23 paper/main.tex
Updates from Overleaf

@@
 
-\section{Geometric convolutions}\label{sec:convolution}
+\paragraph{Related work:}
 
-We consider the orthogonal group $O(d)$ and $g\in O(d)$. For our purposes $d$ will be 2 or 3, but everything generalizes.  
-We consider an image $A$ in a $d$-torus with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
-Each pixel can contain tensors and pseudotensors (of which scalars, pseudoscalars, vectors, and pseudovectors are particular cases) that we define below.
-Similarly, each linear geometric convolution operator $C$ will be a pixelized cubical image patch with $M$ pixels (with $M$ odd) in each dimension ($M^d$ pixels total) with a tensor or pseudotensor in each pixel, which acts as a linear pixel weight.
+\paragraph{Our contribution:}
+
+\section{Geometric objects and geometric images}\label{sec:geometric}
+
+In this \sectionname{} we define the geometric objects and geometric images that we use to generalize classical images in scientific contexts.
+The main point is that the channels of geometric images---which will be like the components of vectors and tensors---are not independent.
+There is an underlying action by the orthogonal group that rules what are the allowed objects and transformations in this space.
+
+We first fix $d$, the dimension of the space (typically 2 or 3 but it could be arbitrary). The geometric objects are vectors and tensors. 
+The orthogonal group $O(d)$ can act on $v \in \mathbb R^d$ in the following ways:
+\begin{align} \label{eq.action}
+    g\cdot v = \det(M(g))^p M(g) v 
+\end{align}
+where $g\in O(d)$, $M(g)\in \mathbb R^{d\times d}$ is the standard matrix representation of $g$, and $p\in\{0,1\}$ is the parity of the action.
+If $p=0$ we obtain the standard $O(d)$ action on $\mathbb R^d$ \emph{vectors}.
+If $p=1$ we obtain what in physics is commonly known as the action of $O(d)$ on \emph{pseudovectors}.
+In this context the objects are defined by the actions that they carry.
+
+\begin{definition}[$k$-$p$-tensors] \label{def.tensors}
+We say that $v\in \mathbb R^d$ is a 1-$p$-tensor if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
+If $v_i$ is a 1-$p_i$-tensor then $T:=v_{1}\otimes\ldots \otimes v_k$ is a rank-1 $k$-$p$-tensor, where $p=\sum_{i=1}^k p_i \mod 2$ and the action of $O(d)$ is defined as
+\begin{align}
+    g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)
+\end{align}
+Higher rank $k$-$p$-tensors are defined as linear combinations of rank-1  $k$-$p$-tensors where the action of $O(d)$ is extended linearly.
+\end{definition}
+
+We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are \emph{pseudovectors}, the 0-0-tensors are \emph{scalars}, 0-1-tensors are \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are known as \emph{tensors}.
+
+We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
+Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel).
+Sometimes we will consider the image to be in a $d$-torus.
+
+\section{Linear functions of geometric images}
+
+%Similarly, each linear geometric convolution operator $C$ will be a pixelized cubical image patch with $M$ pixels (with $M$ odd) in each dimension ($M^d$ pixels total) with a tensor or pseudotensor in each pixel, which acts as a linear pixel weight.
 
@@
 
-\begin{definition}[vectors, scalars, and $k$-tensors] \label{def.tensors}
-We say that $v\in \mathbb R^d$ is a vector (or 1-tensor) if $O(d)$ acts on $v$ via the canonical action.
-Namely, for any $g\in O(d)$ we have $g\cdot v = M(g)\, v$ where $M(g)\in \mathbb R^{d\times d}$ is the the matrix representation of $g$.
-We say that $a\in \mathbb R$ is a scalar (or 0-tensor) if $O(d)$ acts trivially on $a$.
-Namely $g\cdot a = a$.
-We say that $T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor if $O(d)$ acts on $T$ as follows:
-$O(d)$ acts on a rank-1 $k$-tensor by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and it acts on higher rank tensors by linearly extending the actions to linear combinations of such rank-1 $k$-tensors.
-\end{definition}
-
 One consequence of the above definitions is that 
@@
 
-\begin{definition}[pseudovectors, pseudoscalars, and $k$-pseudotensors]
-We say that $B\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot B := \det(g)\, g \cdot B$ where $g\cdot B$ is the canonical action on $k$-tensors as in Definition \ref{def.tensors}.
-The pseudovectors and pseudoscalars are the 1-pseudotensors and 0-pseudotensors respectively.
+\begin{definition}[outer products of tensors]
+If $a$ is a $k$-$p$-tensor and $b$ is a $k'$-$p'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-$(p+p')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$
 \end{definition}
 
-\begin{definition}[outer products of tensors]
-If $a$ is a $k$-tensor and $b$ is a $k'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$
-Similarly, if $a$ is a $k$-pseudotensor and $b$ is a $k'$-pseudotensor, then $a\otimes b$ is a $(k+k')$-tensor.
-If $a$ is a $k$-tensor and $b$ is a $k'$-pseudotensor, or if $a$ is a $k$-pseudotensor and $b$ is a $k'$-tensor, then $a\otimes b$ is a $(k+k')$-pseudotensor.
+\begin{definition}[Kronecker contractions of tensors]
+Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
+We define the $(\alpha,\beta)$ Kronecker contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$.
+In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$s are placed in the positions $\alpha$ and $\beta$:
+\begin{equation}
+[a^{(\alpha,\beta)}]_{i_1,\ldots, i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, i_{\alpha-1}, x, i_{\alpha+1}, \ldots, i_{\beta-1}, x, i_{\beta+1}.
+\ldots, i_k}
+\end{equation}
+There are $\choose{k}{2}$ different index choices for this contraction, corresponding to all the possible index pairs among $k$ original indices.
 \end{definition}
 
-\begin{definition}[contractions of tensors]
-Let $a$ be a $k$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
-We define the $a^{(\alpha,\beta)}$ contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$.
-In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$'s are placed in the positions $\alpha$ and $\beta$:
-\begin{equation}
-[a^{(\alpha,\beta)}]_{i_1,\ldots i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, x, \ldots , x, \ldots, i_k}
-\end{equation}
+\begin{definition}[Levi-Civita contractions of tensors]
+Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq d-1$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
+We define the Levi-Civita contraction of $a$ to be the $(k-d+2)$-$(p+1)$-tensor resulting from taking the outer product of
+$a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol.
+There are $(d-1)!\,\choose{k}{d-1}$ different index choices for this contraction corresponding to an ordered list of $d-1$ indices chosen from among $k$ choices.
 \end{definition}
@@
 
+\section{Linear functions of geometric images}\label{sec:convolution}
+
+In this \sectionname{} we define the linear operations that can be performed on a geometric image such that the outputs of these operations are also geometric images.
+
 \begin{definition}[geometric convolution]
@@
 
-Geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor and pseudotensor inputs and outputs.
+Geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor inputs and outputs.
 
-The linear geometric convolution operation can be followed by contractions with Kroenecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
-The Kroenecker delta is the object with 2 indices such that if the two indices are identical the value is $1$ and the value is $0$ otherwise.
+The linear geometric convolution operation can be followed by contractions with Kronecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
+The Kronecker delta is the object with 2 indices such that if the two indices are identical the value is $1$ and the value is $0$ otherwise.
 The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
 For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
-Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-tensor.
-Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
+Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-tensor.
+Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
 Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired.
@@
 
+\section{Higher-order functions of geometric images}\label{sec:nonlinear}
+
 With these definitions in hand, we can make geometric filters, take outer products, and contract to make a combinatorically large set of nonlinear geometric convolution operators on images made up of geometric objects.
@@
 
-\section{Equivariant geometric convolutions}\label{sec:equivariant}
+\section{Equivariant functions of geometric images}\label{sec:equivariant}
+
+\section{Expressive power of geometric convolutions}\label{sec:universality}
+
+\section{Implementation and numerical examples}\label{sec:examples}
+
+\section{Discussion}\label{sec:discussion}
+
+Hello world
+
+\appendix
+
+\section{All that follows needs to move up above}
 
@@
   Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
-  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kroenecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kroenecker contractions.
+  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
 \item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
@@
   There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
-  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kroenecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kroenecker contractions.
+  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
 (SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)

### 2022-08-23 paper/main.tex
furniture re-arranged

@@
 
-We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
-Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel).
-Sometimes we will consider the image to be in a $d$-torus.
-
-\section{Linear functions of geometric images}
-
-%Similarly, each linear geometric convolution operator $C$ will be a pixelized cubical image patch with $M$ pixels (with $M$ odd) in each dimension ($M^d$ pixels total) with a tensor or pseudotensor in each pixel, which acts as a linear pixel weight.
-
-For some of the expressions below we will use Ricci and Levi-Civita summation notation (or Einstein summation notation).
+\begin{definition}[summation notation]
 In this notation, tensor products are written in component form, and repeated indices are summed over.
@@
 This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice).
+\end{definition}
 
 One consequence of the above definitions is that 
-$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor iff (in summation notation)
-$[g\cdot T]_{i_1,\ldots, i_k} = [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}$ for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
-For example, a 2-tensor is defined by its transformation property
+$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-$p$-tensor if and only if (in summation notation)
+\begin{align}
+    [g\cdot T]_{i_1,\ldots, i_k} = \det(M(g))^p\, [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}
+\end{align} for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
+For example, a 2-0-tensor is defined by its transformation property
 $[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
@@
 
+We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
+Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel).
+Sometimes we will consider the image to be in a $d$-torus.
+Let $\mathcal T_{d,k,p}$ the set of $k$-$p$-tensors in $\mathbb R^d$. We define the geometric images as follows.
+\begin{definition}[geometric image]
+A geometric image is a function $A:[N]^d \to \mathcal T_{d,k,p}$, where $[N]=\{0,1,\ldots, N-1\}$.
+We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
+\end{definition}
+
+\section{Generalized convolutions on geometric images}\label{sec:convolution}
+
+In this \sectionname{} we generalize the notion of convolution to take geometric images as inputs and return geometric images as outputs.
+The idea is that the (presumably large) geometric image (containing $k$-$p$-tensors) is convolved with a (presumably small) geometric filter (also a geometric image, but containing $k'$-$p'$-tensors) to produce an output image that contains $(k+k')$-$(p+p')$-tensors, where each pixel is a sum of outer products.
+These outer products can be contracted down to lower-$k$ tensors using Kronecker or Levi-Civita contractions.
+
 \begin{definition}[outer products of tensors]
@@
 \end{definition}
+
+If the next definition is going to use ``the torus'' then we should say a few words about that up-front, or above in the image definition.
+
+\begin{definition}[geometric convolution]
+Let $A$ be a $k$-$p$-tensor image on the $d$-torus with side length $N$.
+Let $C$ be a $k'$-$p'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
+The geometric convolution $A\ast C$ is a $(k+k')$-$(p+p')$-tensor image such that
+\begin{equation}
+    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
+\end{equation}
+where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
+\end{definition}
+This definition is on the torus. If you are not in the torus just pad with zero tensors of the corresponding order and parity.
+
+Convolutions can be followed by Kronecker contractions, Levi-Civita contractions and index permutations.
+The Kronecker delta is the object with two indices such that it has the value $+1$ when the two indices have the same value, and $0$ otherwise.
+Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-tensor.
+Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-pseudotensor, because the contraction sums out one pair of indices.
+The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices (where $d$ is the spatial dimension) such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
+Contraction of a $k$-$p$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-$(p+1)$-pseudotensor, because the contraction sums over $d$ pairs of indices, and it changes the parity from even to odd or odd to even.
+For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
+For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
+$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a pseudovector.
 
@@
 \end{equation}
+\end{definition}
 There are $\choose{k}{2}$ different index choices for this contraction, corresponding to all the possible index pairs among $k$ original indices.
-\end{definition}
 
@@
 $a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol.
+\end{definition}
 There are $(d-1)!\,\choose{k}{d-1}$ different index choices for this contraction corresponding to an ordered list of $d-1$ indices chosen from among $k$ choices.
-\end{definition}
 
 \begin{definition}[permutations of tensor indices]
-Given a $k$-tensor or $k$-pseudotensor $T\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $T^\sigma$ the result of permuting the indices of $T$ according to $\sigma$:
+Given a $k$-$p$-tensor $a\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $a^\sigma$ the result of permuting the indices of $a$ according to $\sigma$:
 \begin{equation}
-    [T^\sigma]_{i_1, \ldots, i_k} := [T]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
+    [a^\sigma]_{i_1, \ldots, i_k} := [a]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
 \end{equation}
@@
 
-\begin{definition}[$k$-tensor image]
-We say that $A$ is a $k$-tensor image if $A$ is a function $A: [N]^d \to (\mathbb R^d)^{\otimes k}$ such that $A(\bar \i)$ is a $k$-tensor for all $\bar \i \in [N]^d$, and $[N]=\{0,\ldots, N-1\}$.
-By a similar definition there are also $k$-pseudotensor images.
-We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
+Define POOLING operation?
+
+The following is not correct -- doesn't include permutations or pooling, and doesn't get the combinatorics right.
+
+Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY?
+IF NOT, IT SHOULD BE DEFINED.
+IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
+Here are the linear operations (linear in the input image) that are permitted):
+\begin{itemize}
+\item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
+  Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
+  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
+\item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
+  Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
+  Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
+  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
+  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
+  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
+(SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
+\end{itemize}
+
+other things that may be useful: eigenvalues
+
+is it obvious how this contractions and permutations apply to the entire image? the same permutation, contraction should be applied to all pixels of the image.
+
+\section{Higher-order functions of geometric images}\label{sec:nonlinear}
+
+The convolution and contraction operations above defines a large class of linear functions from geometric images to geometric images. Non-linear functions can be constructed as products.
+
+Algorithm: \newline
+Construct all linear functions using the constructions from previous section. \newline
+Construct all pixel-wise products across them. \newline
+All linear functions of all pixel-wise products. \newline
+That gives everything to second order. \newline
+
+\begin{definition}[Products of geometric images]
+Given an $N^d$-pixel $k_A$-tensor image $A$ and an $N^d$-pixel $k_B$-tensor image $B$, we can define a product $A\otimes B$, which is an $N^d$-pixel $(k_A+k_B)$-tensor image such that $(A\otimes B)(\bar\i)=A(\bar\i)\otimes B(\bar\i)$.
+These products can be contracted by the contraction rules above.
 \end{definition}
 
-\section{Linear functions of geometric images}\label{sec:convolution}
-
-In this \sectionname{} we define the linear operations that can be performed on a geometric image such that the outputs of these operations are also geometric images.
-
-\begin{definition}[geometric convolution]
-Let $A$ be a $k$-tensor image on the $d$-torus with side length $N$.
-Let $C$ be a $k'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
-The geometric convolution $A\ast C$ is a $k+k'$-tensor image such that
+\begin{definition}[Polynomial local geometric operator of degree $\ell$]
+Given a $k$-tensor image $A$ and a set of $\ell$ $k'_j$-tensor geometric filters $C_j$, a geometric image can be formed by the degree-$\ell$ image product
 \begin{equation}
-    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
+    (A\ast C_1) \otimes (A\ast C_2) \otimes \cdots \otimes (A\ast C_\ell) ~. \label{monomial}
 \end{equation}
-where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
-If $A$ is a $k$-pseudotensor image and $C$ is a $k'$-pseudotensor filter, then $A\ast C$ is also a $k+k'$-tensor.
-If $A$ is $k$-tensor and $C$ is $k'$-pseudotensor, or if $A$ is $k$-pseudotensor and $C$ is $k'$-tensor, then $A\ast C$ is a $k+k'$-pseudotensor.
+This product can be subsequently contracted $r$ times to deliver a $K$-tensor geometric image with
+\begin{equation}
+    K = \ell\,k - 2\,r + \sum_{j=1}^\ell k'_j ~.
+\end{equation}
+A polynomial local geometric operator is a function that takes a $k$-tensor image $A$ and outputs the sum of $K$-tensor images constructed as contractions to order $K$ of tensor images from \eqref{monomial} with filters chosen by the user.
 \end{definition}
 
-Geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor inputs and outputs.
+The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
+They are local in the sense that, at every pixel $\bar\i$ of the output image $(A\ast C)$, the convolution only makes use of the pixels of $A$ within a small neighborhood of of $\bar\i$.
+Because tensors can be multiplied together using the outer product, and contracted according to the tensor contractions defined in \secref{sec:convolution}, it is possible, by multiplication, to construct a family of \emph{nonlinear} local geometric operators that take as input geometric images and produce as output geometric images.
+This family of nonlinear local operators will be equivariant to the group $G_d$ when the linear local operators of which they are composed are equivariant.
 
-The linear geometric convolution operation can be followed by contractions with Kronecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
-The Kronecker delta is the object with 2 indices such that if the two indices are identical the value is $1$ and the value is $0$ otherwise.
-The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
-For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
-Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-tensor.
-Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
-Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired.
-SOLE DO WE NEED TO SAY MORE HERE?
-Contraction of a $k$-pseudotensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-tensor.
-
-For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
-For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
-$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.
-
-\section{Higher-order functions of geometric images}\label{sec:nonlinear}
+Assembling the parts into a general, linear function.
 
@@
 
+Assemble all the parts into an algorithm for making all possible monomials.
+
 \section{Equivariant functions of geometric images}\label{sec:equivariant}
 
-\section{Expressive power of geometric convolutions}\label{sec:universality}
+Define equivariance.
 
-\section{Implementation and numerical examples}\label{sec:examples}
+Prove translation invariance.
 
-\section{Discussion}\label{sec:discussion}
-
-Hello world
-
-\appendix
-
-\section{All that follows needs to move up above}
+Prove thing about invariant filters.
 
@@
 
-SOLE: WHERE SHOULD THIS GO?
-Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY?
-IF NOT, IT SHOULD BE DEFINED.
-IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
-Here are the linear operations (linear in the input image) that are permitted):
-\begin{itemize}
-\item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
-  Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
-  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
-\item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
-  Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
-  Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
-  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
-  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
-(SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
-\end{itemize}
-
-\section{Nonlinear equivariant local geometric operators}\label{sec:nonlinear}
-
-The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
-They are local in the sense that, at every pixel $\bar\i$ of the output image $(A\ast C)$, the convolution only makes use of the pixels of $A$ within a small neighborhood of of $\bar\i$.
-Because tensors can be multiplied together using the outer product, and contracted according to the tensor contractions defined in \secref{sec:convolution}, it is possible, by multiplication, to construct a family of \emph{nonlinear} local geometric operators that take as input geometric images and produce as output geometric images.
-This family of nonlinear local operators will be equivariant to the group $G_d$ when the linear local operators of which they are composed are equivariant.
-
-\begin{definition}[Products of geometric images]
-Given an $N^d$-pixel $k_A$-tensor image $A$ and an $N^d$-pixel $k_B$-tensor image $B$, we can define a product $A\otimes B$, which is an $N^d$-pixel $(k_A+k_B)$-tensor image such that $(A\otimes B)(\bar\i)=A(\bar\i)\otimes B(\bar\i)$.
-These products can be contracted by the contraction rules above.
-\end{definition}
-
-\begin{definition}[Polynomial local geometric operator of degree $\ell$]
-Given a $k$-tensor image $A$ and a set of $\ell$ $k'_j$-tensor geometric filters $C_j$, a geometric image can be formed by the degree-$\ell$ image product
-\begin{equation}
-    (A\ast C_1) \otimes (A\ast C_2) \otimes \cdots \otimes (A\ast C_\ell) ~. \label{monomial}
-\end{equation}
-This product can be subsequently contracted $r$ times to deliver a $K$-tensor geometric image with
-\begin{equation}
-    K = \ell\,k - 2\,r + \sum_{j=1}^\ell k'_j ~.
-\end{equation}
-A polynomial local geometric operator is a function that takes a $k$-tensor image $A$ and outputs the sum of $K$-tensor images constructed as contractions to order $K$ of tensor images from \eqref{monomial} with filters chosen by the user.
-\end{definition}
-
-\section{Expressive power of this architecture}
-We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).
-
-We can cite that convolutions with small patches may be seen as low pass filters in some way. Reference?
-
-\section{Construction of equivariant geometric convolution filters}
+\subsection{Construction of equivariant geometric convolution filters}
 In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
@@
 In detail, group averaging happens HOW? That is, what is the exact mathematical operation?
+
+\section{Expressive power of geometric convolutions}\label{sec:universality}
+
+Counting arguments.
+
+Comments on locality.
+
+We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).
+
+We can cite that convolutions with small patches may be seen as low pass filters in some way. Reference?
+
+\section{Implementation and numerical examples}\label{sec:examples}
+
+How do we make the invariant filters?
+
+How do we make the linear functions?
+
+What do they look like?
+
+Etc.
+
+An example?
 
@@
 
-\section{Implementation notes}
 \begin{itemize}
@@
 
-\section{Discussion and comments}
+\section{Discussion}\label{sec:discussion}
+
+Hello world
+
+One discussion point: you might be able to get back to a continuous symmetry group if you consider the image to be a VIEW of a large, continuous tensor field...?
 

### 2022-08-24 notebooks/geometric.py
notes to self

@@ -10,10 +10,13 @@ See the file `LICENSE` for more details.
 - David W. Hogg (NYU)
 
 ## To-do items:
-- Create tests for group operations on k-tensors.
+- Create better tests for group operations on k-tensors.
 - Fix sizing of multi-filter plots.
+- Switch over to jax so this is useful for ML people.
+- Switch the structure of the image and filter so they make better use of jax.numpy array objects.
 - Need to implement index permutation operation.
-- Need to implement Levi-Civita contraction. Uh huh.
+- Need to implement Levi-Civita contraction for general dimensions.
+- Need to build tests for the contractions.
 - Need to implement bin-down and bin-up operators.
 """
 

### 2022-08-24 paper/main.tex

@@
 %  -- Define "local" functions, is there a characterization of functions that can be written as convolutions with small filters?
+% - There is a one-indexing and zero-indexing issue. Should we be consistent?
+% - Some definitions begin with "given" some with "let". Should we try to be more consistent?
 
@@
 \end{definition}
+Typically we would make the parity $p$ be a bit (either 0 for even parity or 1 for odd parity) but in principle it can be any integer.
 
-We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are \emph{pseudovectors}, the 0-0-tensors are \emph{scalars}, 0-1-tensors are \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are known as \emph{tensors}.
+We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are known as \emph{pseudovectors}, the 0-0-tensors are known as \emph{scalars}, 0-1-tensors are known as \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are known as \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
 
@@
 \begin{definition}[outer products of tensors]
-If $a$ is a $k$-$p$-tensor and $b$ is a $k'$-$p'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-$(p+p')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$
+If $a$ is a $k$-$p$-tensor and $b$ is a $k'$-$p'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-$(p+p')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$.
 \end{definition}
 
-If the next definition is going to use ``the torus'' then we should say a few words about that up-front, or above in the image definition.
+SOLE: If the next definition is going to use ``the torus'' then we should say a few words about that up-front, or above in the image definition.
 
@@
 \end{definition}
-This definition is on the torus. If you are not in the torus just pad with zero tensors of the corresponding order and parity.
+This definition is on the torus. If you are not on the torus, just pad the image out with zero tensors of the corresponding order and parity.
 
 Convolutions can be followed by Kronecker contractions, Levi-Civita contractions and index permutations.
-The Kronecker delta is the object with two indices such that it has the value $+1$ when the two indices have the same value, and $0$ otherwise.
-Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-tensor.
-Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-pseudotensor, because the contraction sums out one pair of indices.
+The Kronecker delta ($\delta_{ij}$) is the object with two indices $ij$ such that it has the value $+1$ when the two indices have the same value ($i=j$), and $0$ otherwise.
+Below we define the contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta, which returns a $(k-2)$-$p$-tensor.
 The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices (where $d$ is the spatial dimension) such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
-Contraction of a $k$-$p$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-$(p+1)$-pseudotensor, because the contraction sums over $d$ pairs of indices, and it changes the parity from even to odd or odd to even.
+We define the contraction of a $k$-$p$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol, which returns a $(k-d+2)$-$(p+1)$-tensor, because the contraction sums over $d$ pairs of indices, and it changes the parity from even to odd or odd to even.
 For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
@@
 
-\begin{definition}[Kronecker contractions of tensors]
+\begin{definition}[Kronecker contractions of tensors and geometric images]
 Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
@@
 \begin{equation}
-[a^{(\alpha,\beta)}]_{i_1,\ldots, i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, i_{\alpha-1}, x, i_{\alpha+1}, \ldots, i_{\beta-1}, x, i_{\beta+1}.
-\ldots, i_k}
+[a^{(\alpha,\beta)}]_{i_1, \ldots, i_k \backslash i_\alpha,i_\beta} := \sum_{i_\alpha=1}^d \sum_{i_\beta=1}^d [a]_{i_1, \ldots, i_\alpha,\ldots,i_\beta,\ldots, i_k}\,\delta_{i_\alpha, i_\beta}
 \end{equation}
+When Kronecker contraction is applied to a geometric image, the contraction is applied to every $k$-$p$-tensor in the image identically.
 \end{definition}
@@
 
-\begin{definition}[Levi-Civita contractions of tensors]
+\begin{definition}[Levi-Civita contractions of tensors and geometric images]
 Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq d-1$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
@@
 $a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol.
+When a Levi-Civita contraction is applied to a geometric image, the contraction is applied to every $k$-$p$-tensor in the image identically.
 \end{definition}
-There are $(d-1)!\,\choose{k}{d-1}$ different index choices for this contraction corresponding to an ordered list of $d-1$ indices chosen from among $k$ choices.
+There are $(d-1)!\,\choose{k}{d-1}$ different index choices for the Levi-Civita contraction of a $k$-$p$-tensor.
+These choices correspond to the choices for an ordered list of $d-1$ indices chosen from among $k$ index choices.
 
 \begin{definition}[permutations of tensor indices]
-Given a $k$-$p$-tensor $a\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $a^\sigma$ the result of permuting the indices of $a$ according to $\sigma$:
+Given a $k$-$p$-tensor $a$ and $\sigma\in S_k$ a permutation. We consider $a^\sigma$ the result of permuting the indices of $a$ according to $\sigma$:
 \begin{equation}
@@
 \end{equation}
+When index permutation is applied to a geometric image, the permutation is applied to every $k$-$p$-tensor in the image identically.
 \end{definition}
 
-Define POOLING operation?
+In addition to these contractions and index permutations that act pixel-wise in geometric images, it is possible to change the image size.
+If $N$ is divisible by integer $f>1$, a geometric image of side length $N$ can be reduced in size to a geometric image of side length $N/f$ by the equivalent of convolution with a trivial $f\times f$ scalar filter (geometric filter of 0-0-tensors) operating with a stride of $f$.
+Similarly a geometric image of side length $N$ can be expanded in size to a side length $f\,N$ by application of an interpolation operator with stride $1/f$.
+\begin{definition}[pooling and un-pooling]
+foo and bar.
+\end{definition}
 
-The following is not correct -- doesn't include permutations or pooling, and doesn't get the combinatorics right.
-
-Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY?
-IF NOT, IT SHOULD BE DEFINED.
-IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
-Here are the linear operations (linear in the input image) that are permitted):
+Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-$p$-tensor image of side length $N$ to an output $k'$-$p'$-tensor image of side length $N'$ (where either $N$ is divisible by $N'$, or $N'$ is divisible by $N$).
+Here are the linear operations (linear in the input image) that are permitted:
+\begin{enumerate}
+    \item Perform a convolution and index contraction. For this step there are two qualitatively different options:
         \begin{itemize}
@@
               Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
+              There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
             \item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
@@
               Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
-  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
+              There are $$(d-1)!\,\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
               Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
-  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
-(SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
+              There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
         \end{itemize}
+    \item Perform (generally after the convolution and contraction) any index permutation you like (provided that $k'>1$).
+      There are $k'!$ choices for the index permutation.
+    \item Perform (generally after all the above) a pooling or unpooling step to get the image to side length $N'$.
+\end{enumerate}
 
-other things that may be useful: eigenvalues
-
-is it obvious how this contractions and permutations apply to the entire image? the same permutation, contraction should be applied to all pixels of the image.
+In principle there could also have been an \emph{eigenvalues} operator that is linear in the input and produces $d$ $0$-$p$-tensors.
+But since any such operation can only reasonably be applied to $2$-$p$-tensors, and even for those, only Hermitian ones, we don't include this operation here.
 
@@
 
-The convolution and contraction operations above defines a large class of linear functions from geometric images to geometric images. Non-linear functions can be constructed as products.
+The convolution, contraction, index-permutation, and pooling operators above effectively span a large class of linear functions from geometric images to geometric images.
+Non-linear functions can be constructed as polynomials, or sums of products of linear function outputs.
+Products here will be outer products (possibly followed by further geometric convolutions and contractions).
 
-Algorithm: \newline
-Construct all linear functions using the constructions from previous section. \newline
-Construct all pixel-wise products across them. \newline
-All linear functions of all pixel-wise products. \newline
-That gives everything to second order. \newline
-
-\begin{definition}[Products of geometric images]
-Given an $N^d$-pixel $k_A$-tensor image $A$ and an $N^d$-pixel $k_B$-tensor image $B$, we can define a product $A\otimes B$, which is an $N^d$-pixel $(k_A+k_B)$-tensor image such that $(A\otimes B)(\bar\i)=A(\bar\i)\otimes B(\bar\i)$.
-These products can be contracted by the contraction rules above.
+\begin{definition}[outer products of images]
+Given a $k$-$p$-tensor image $A$ and a $k'$-$p'$-tensor image $B$, both of side length $N$, the outer product $A\otimes B$ of these images is a $(k+k')$-$(p+p')$-tensor image such that
+\begin{equation}
+    (A\otimes B)(\bar\i) = A(\bar\i)\otimes B(\bar\i) ~.
+\end{equation}
+That is, the outer products of geometric images are performed pixelwise.
 \end{definition}
 
-\begin{definition}[Polynomial local geometric operator of degree $\ell$]
-Given a $k$-tensor image $A$ and a set of $\ell$ $k'_j$-tensor geometric filters $C_j$, a geometric image can be formed by the degree-$\ell$ image product
+\begin{definition}[sums of images]
+Given a $k$-$p$-tensor image $A$ and a $k$-$p$-tensor image $B$, both of side length $N$, the sum $A+B$ of these images is a $k$-$p$-tensor image such that
 \begin{equation}
-    (A\ast C_1) \otimes (A\ast C_2) \otimes \cdots \otimes (A\ast C_\ell) ~. \label{monomial}
+    (A+B)(\bar\i) = A(\bar\i)+B(\bar\i) ~.
 \end{equation}
-This product can be subsequently contracted $r$ times to deliver a $K$-tensor geometric image with
-\begin{equation}
-    K = \ell\,k - 2\,r + \sum_{j=1}^\ell k'_j ~.
-\end{equation}
-A polynomial local geometric operator is a function that takes a $k$-tensor image $A$ and outputs the sum of $K$-tensor images constructed as contractions to order $K$ of tensor images from \eqref{monomial} with filters chosen by the user.
+That is, the sums of geometric images are performed pixelwise.
 \end{definition}
 
+Now to construct all degree-2 monomials that produce a $k'$-$p'$-tensor image of side length $N'$ output from a $k$-$p$-tensor image of side length $N$ input, here are the rules:
+\begin{enumerate}
+    \item By the rules in \secref{sec:convolution}, construct any linear function that takes the input geometric image and outputs a geometric image of any order $k''$ and any parity $p''$ and any side length $N''$.
+    \item By the rules in \secref{sec:convolution}, construct any other linear function that takes the input geometric image and outputs a geometric image of any order and any parity.
+    There is no need for this second output image to be the same order or parity as the first, but it does have to have the same side length.
+    \item Outer product these two images.
+    \item By the rules in \secref{sec:convolution}, construct any linear function that takes this image outer product as input and produces a $k'$-$p'$-tensor image of side length $N'$ as output.
+\end{enumerate}
+We conjecture that these rules will generate all possible second-degree monomial functions of the input image (SOLE SUBJECT TO SOME CONDITIONS) that meet the output requirements.
+A polynomial function of degree 2 can be constructed by linear combinations of these second-degree monomial functions along with all the linear functions that also produce $k'$-$p'$-tensor images of side length $N'$ (plus perhaps constant images).
+Higher-degree monomials and polynomials can be made out of products of lower-degree monomials, by generalization of the above.
+
+SOLE: THIS PARAGRAPH OR THESE IDEAS GO ELSEWHERE, I THINK.
 The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
@@
 
-Assembling the parts into a general, linear function.
+\section{Equivariant functions of geometric images}\label{sec:equivariant}
 
-With these definitions in hand, we can make geometric filters, take outer products, and contract to make a combinatorically large set of nonlinear geometric convolution operators on images made up of geometric objects.
-But before we go nonlinear, we will go equivariant.
-
-Assemble all the parts into an algorithm for making all possible monomials.
-
-\section{Equivariant functions of geometric images}\label{sec:equivariant}
+Introduce the discrete group. Why discrete?
 

### 2022-08-26 paper/main.tex
Introduction text

@@
 \title{\bfseries%
-Convolution-based functions
-on tensor fields and tensor images}
+Convolution-based functions on tensor fields and tensor images}
 \author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
@@
 containing scalars, vectors, and tensors, producing scalar, vector, and tensor outputs.
-Our formulation, which combines convolutions with products, tensor contractions, and tensor index permutations, does not require point-wise activation functions (RELU, say) but can universally approximate any suitably local analytic scalar, vector, or tensor functions on scalar, vector, or tensor images.
-Our method permits, with a very simple adjustment, restriction to function spaces that are exactly equivariant to translations, discrete rotations, and reflections.
+Our formulation, which combines a geometric generalization of convolution with products, tensor index contractions, contractions with Levi-Civita symbols, and tensor index permutations, can universally approximate any suitably local analytic scalar, vector, or tensor functions on scalar, vector, or tensor images.
+The framework permits, with a very simple adjustment, restriction to function spaces that are exactly equivariant to translations, discrete rotations, and reflections, but the framework will be useful even in contexts in which the functions are not expected to be equivariant.
 
@@
 Contemporary natural science teems with data sets that are images or lattices or grids of geometric objects.
-Rectangular images or lattices of density (a scalar field), temperature (also a scalar field), velocity (a vector field), magnetic field (a pseudo-vector field), and polarization (a 2-tensor field) are regularly taken with detectors or made in simulations.
-That is, there is plenty of data consisting of geometric objects on regular grids.
-These kinds of images (grids of scalars, vectors, pseudo-vectors, and tensors, for example) are \emph{geometric} in the sense that they contain (at every pixel) objects are defined by their transformation properties under rotation and reflection.
-For example, scalars are defined to be quantities in $\mathbb R$ that are invariant with respect to all rotations and reflections; vectors are defined (in $d$-dimensional space) to be quantities in $\mathbb R^d$ that transform under rotation or reflection by multiplication by the orthogonal $d\times d$ matrix representing the rotation or reflection; and so on.
+These might be observations of intensities (scalars), magnetic fields (pseudovectors), or polarizations (2-tensors) on a surface or in a volume.
+Or they might be the inputs or outputs of a simulation, in which the initial conditions or fields are specified on a regular grid.
+Any lattice or grid of vectors or tensors can be seen as a generalization of the concept of an image, in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor) in each pixel.
+These objects are \emph{geometric} in the sense that they are defined in terms of their transformation properties under geometric operators such as rotation, translation, and reflection.
+Thus there is a need for machine learning methods designed for \emph{geometric images}---lattices or grids of scalars, vectors, and tensors.
+There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines (for example, CITE).
 
-For a concrete example close to one of our hearts (DWH's), contemporary cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
-For this reason, accurate, fast emulators might be of enormous economic and scientific value.
-There are now quite a few successful projects implementing such emulators (CITE THINGS).
-Importantly, the initial conditions for these simulations are tiny displacements (vectors) and velocities (vectors) relative to a cubical lattice on a 3-torus.
-That is, the initial conditions can be described as a 3-dimensional image of vectors on a regular grid.
-For another concrete example, astronomical imaging data taken of the interstellar medium can be represented as rectangular lattices of measurements of temperature (scalar), optical depth (scalar), and induced polarization of starlight (a 2-tensor).
-Here the data can be described as a 2-dimensional image of scalars and tensors on a regular grid.
+%For a concrete example close to one of our hearts (DWH's), contemporary cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
+%For this reason, accurate, fast emulators might be of enormous economic and scientific value.
+%There are now quite a few successful projects implementing such emulators (CITE THINGS).
+%Importantly, the initial conditions for these simulations are tiny displacements (vectors) and velocities (vectors) relative to a cubical lattice on a 3-torus.
+%That is, the initial conditions can be described as a 3-dimensional image of vectors on a regular grid.
+%For another concrete example, astronomical imaging data taken of the interstellar medium can be represented as rectangular lattices of measurements of temperature (scalar), optical depth (scalar), and induced polarization of starlight (a 2-tensor).
+%Here the data can be described as a 2-dimensional image of scalars and tensors on a regular grid.
 
-The application of machine-learning techniques to these problems can be dangerous:
-Machine-learning models have enormous freedom; the laws of physics are constrained to obey a large set of extremely strict symmetries, including rotation, reflection, translation, permutation, and boost equivariances, plus constraints such as units and dimensions symmetry, unitarity, and myriad conservation laws.
-If we want to operate safely on these problems, we want these symmetries built in from the start, and exactly.
+At the present day, the go-to tools for machine learning with images are convolutional neural networks (CNNs; CITE) and their many descendents (for example, CITE).
+These methods are designed to work on one- or few-channel images in which, at the bottom (early) layers of the network, the operations involve image convolutions with learned filters followed by the application of pointwise nonlinearities.
+In typical contexts, the channels of multi-channel input images will be something like the red, green, and blue (or cyan, magenta, yellow, and black) channels of a color image; these can be combined arbitrarily in the layers of the CNN.
+When these CNN-based tools are applied to lattices of vectors (say), typically the components of the vectors are treated just as channels of the input image (for an example, see CITE) and then everything proceeds as with multi-channel color images.
+This is not good!
 
-HOGG: Differences between continuous and discrete symmetries.
+Actually, it \emph{is} good: There are many successful projects that have repurposed CNNs in this way to work on geometric images.
+But there are even better choices.
+Here we propose a set of tools that generalize the concept of convolution to apply to geometric images such that the outputs of the convolutions are also geometric images, obeying the same geometric transformation rules as the inputs.
 
-HOGG: Summarize here how we are going to proceed, in human language.
+The fundamental observation inspiring this work is that when the components of vectors and tensors are acted on with arbitrary functions, the geometric structure of these objects is destroyed.
+There are strict rules, dating back to the invention of differential geometry (CITE RICCI++), about how geometric objects can be combined to produce new geometric objects, consistent with coordinate freedom and transformation rules.
+In previous work (CITE VILLAR++, YAO++) we have capitalized on these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
+Here we use these rules to create a comprehensive set of tools that are universally approximating for functions that take geometric images as input and produce geometric labels (or geometric-image labels) as output.
+
+We are motivated in this work to help solve problems in the natural sciences, where geometric images abound.
+However, we conjecture that these tools are probably very useful even for standard machine-learning image-recognition and image-regression tasks.
+After all, even standard images are measurements of intensity (a scalar) at a regular grid of points on a two-dimensional surface.
+If the labels we seek are also scalars, it will behoove us to obey the rules of geometry.
+
+These rules are roughly as follows:
+A $k$-tensor object (tensor of order $k$) in $d$ dimensions has $k$ indices, each of which can take a value from 1 to $d$; that is, the $k$-tensor is an element of $(\mathbb R^d)^{\otimes k}$.
+It can be contracted to a $(k-2)$-tensor object by identifying a pair of indices and summing over them.
+A $k$-tensor and a $k'$-tensor can be multiplied (outer product) to make a $(k+k')$-tensor object, and then contractions can bring down the order.
+$1$-tensor objects are called vectors and $0$-tensor objects are called scalars.
+There are also odd-parity versions of all these (pseudovectors, pseudoscalars, and pseudotensors), and parity-changing contractions using the Levi-Civita symbol, so in what follows we will define $k$-$p$-tensors that have $k$ indices and a parity $p\in [0, 1]$.
+Two objects can only be added or subtracted if they have the same order $k$ and parity $p$.
+This will all be made precise below.
+But in short, these rules define objects that can be given transformation rules under rotation and reflection such that functions made of these operations are coordinate free, or equivariant to any change of coordinate system (including curvilinear coordinate transformations).
+
+The symmetries that suggest these rules are continuous symmetries.
+But of course images are usually---and for our purposes---discrete grids of values.
+This suggests that in addition to the continuous symmetries respected by the tensor objects in the image pixels there will be discrete symmetries for each geometric image taken as a whole.
+We will define these discrete symmetry groups and use them to define a useful kind of group equivariance for functions of geometric images.
+This equivariance, it turns out, is very easy to enforce, even for nonlinear functions of geometric images, provided that we compose our nonlinear functions from simple geometric operations, including our geometric generalization of convolution.
 
@@
 \end{enumerate}
+The objects obtained by this algorithm may not be linearly independent.
 
@@
 \end{enumerate}
+
 We conjecture that these rules will generate all possible second-degree monomial functions of the input image (SOLE SUBJECT TO SOME CONDITIONS) that meet the output requirements.
@@
 
+There is a continuous set of symmetries that inspire the summation-notation rules. There are discrete symmetries that are exact for these square images. 
 One discussion point: you might be able to get back to a continuous symmetry group if you consider the image to be a VIEW of a large, continuous tensor field...?

### 2022-08-26 notebooks/geometric.py
notes to self

@@ -10,7 +10,9 @@ See the file `LICENSE` for more details.
 - David W. Hogg (NYU)
 
 ## To-do items:
-- Create better tests for group operations on k-tensors.
+- Drop the k-tensor and geometric filter classes; there can be only one class.
+- Move over to jax.
+- Create tests for group operations on k-tensor images.
 - Fix sizing of multi-filter plots.
 - Switch over to jax so this is useful for ML people.
 - Switch the structure of the image and filter so they make better use of jax.numpy array objects.
@@ -20,12 +22,12 @@ See the file `LICENSE` for more details.
 - Need to implement bin-down and bin-up operators.
 """
 
+import itertools as it
 import numpy as np
 import pylab as plt
 import matplotlib.cm as cm
 from matplotlib.colors import ListedColormap
-import cmastro as cma
-import itertools as it
+import cmastro
 
 TINY = 1.e-5
 

### 2022-08-26 notebooks/geometric.py
notes to self

@@ -12,6 +12,8 @@ See the file `LICENSE` for more details.
 ## To-do items:
 - Drop the k-tensor and geometric filter classes; there can be only one class.
 - Move over to jax.
+- Make the geometric_filter inherit geometric_image to reduce repeated code.
+- Is it possible to make the data a big block (not a dictionary) but have the block addressable with keys()? I bet it is...?
 - Create tests for group operations on k-tensor images.
 - Fix sizing of multi-filter plots.
 - Switch over to jax so this is useful for ML people.

### 2022-08-27 paper/main.tex
Updates from Overleaf

@@
 \newcommand{\secref}[1]{\sectionname~\ref{#1}}
+\newcommand{\figref}[1]{\figurename~\ref{#1}}
 
@@
   {1.5ex \@plus .2ex}%
-  {\raggedright\normalfont\large\bfseries}}
+  {\raggedright\normalfont\large\bfseries}}% key thing is: \raggedright!
 \makeatother
@@
 \title{\bfseries%
-Convolution-based functions on tensor fields and tensor images}
+Convolution-based nonlinear functions on tensor fields and tensor images}
 \author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
@@
 This equivariance, it turns out, is very easy to enforce, even for nonlinear functions of geometric images, provided that we compose our nonlinear functions from simple geometric operations, including our geometric generalization of convolution.
+When we enforce this equivariance, the convolution filters that appear look very much like the differential operators that appear in discretizations of vector calculus, which is perhaps not surprising.
 
@@
 For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
-For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
+For another example, the same 2-tensor can (in $d=3$ dimensions) be contracted with the Levi-Civita symbol to give the standard cross product
 $[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a pseudovector.
@@
 
-Introduce the discrete group. Why discrete?
+Introduce the discrete group. Why discrete? SOLE: SEE TEXT BELOW FOR DEFINITIONS.
 
@@
 
-Prove thing about invariant filters.
-
-We will now create linear convolution operators that are group-equivariant, for groups of reflections, rotations, and translations (on the cubical torus).
-We will construct these equivariant convolution operators by an averaging over group operators, followed by a matrix factorization (singular-value decomposition).
+Prove things about invariant filters and equivariant linear functions.
 
@@
 
-\begin{definition}[Action of $B_d$ on $k$-tensors and $k$-pseudotensors]
+\begin{definition}[Action of $B_d$ on $k$-$p$-tensors]
 \begin{align}
-\mbox{$k$-tensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= (g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) \\
-\mbox{$k$-pseudotensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= \det(g)\,(g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) ~,
+g\cdot(v_1 \otimes \ldots \otimes v_k) &= \det(g)^p\,(g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k)
 \end{align}
@@
 The pixel locations in the original image are found by going back, or inverting the transformation.
+
+HOGG: Make a table and a figure showing all the group operators in $B_d$ with $d=2$.
 
@@
 
-\subsection{Construction of equivariant geometric convolution filters}
-In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
-These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
+\paragraph{Numerical construction of invariant geometric convolution filters:}
+In the above we defined equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
+These, in turn, are constructed from products of the outputs of convolutions with invariant geometric filters.
 Here we show how to find all such filters by a group-averaging technique.
@@
 We average the filters over all group operators $g\in B_d$, where $B_d$ is the group of 90-degree rotations and reflections.
-We perform a singular value decomposition of these group-averaged filters to find the $p$ independent orthogonal filters that span the space of equivariant $k'$-tensor (or $k'$-pseudotensor) convolution filters at side length $M$ in $d$ dimensions.
+We perform a singular value decomposition of these group-averaged filters to find the $p$ independent orthogonal filters that span the space of equivariant $k'$-tensor (or $k'$-pseudotensor) convolution filters at side length $M=2\,m+1$ in $d$ dimensions.
+The results of these operations are shown for $d=2$ and $m\in[1, 2]$ in \figref{fig:filters23} and \figref{fig:filters25}.
+HOGG: Give a table saying how many filters there are for each dimension $d$ and each side length $M$ at each order $k$ and each parity $p$.
+\begin{figure}[tp]
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
+\includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter+1_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
+\includegraphics[width=\textwidth]{notebooks/filter+2_2_3.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-2_2_3.png}
+  \end{center}
+\caption{All the filters for $d=2$, $m=1$ ($M=3$). HOGG: notes!\label{fig:filters23}}
+  \end{mdframed}
+\end{figure}
+\begin{figure}[tp]
+  \begin{mdframed}
+  \color{captiongray}
+  \begin{center}
+\includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-0_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter+1_2_5.png}\\
+\includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
+  \end{center}
+\caption{All the filters for $d=2$, $m=2$ ($M=5$). HOGG: notes!\label{fig:filters25}}
+  \end{mdframed}
+\end{figure}
 
-In detail, group averaging happens HOW? That is, what is the exact mathematical operation?
-
-\section{Expressive power of geometric convolutions}\label{sec:universality}
+\section{Expressive power of nonlinear convolution-based functions}\label{sec:universality}
 
@@
 
-How do we make the invariant filters?
+How do we make the invariant filters? See \secref{sec:equivariant}.
 
-How do we make the linear functions?
-
-What do they look like?
-
-Etc.
-
-An example?
-
-\begin{figure}[tp]
-  \begin{mdframed}
-  \color{captiongray}
-  \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter+1_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
-\includegraphics[width=\textwidth]{notebooks/filter+2_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-2_2_3.png}
-  \end{center}
-\caption{All the filters for $d=2$, $m=1$. HOGG: notes!}
-  \end{mdframed}
-\end{figure}
-
-\begin{figure}[tp]
-  \begin{mdframed}
-  \color{captiongray}
-  \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-0_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter+1_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
-  \end{center}
-\caption{All the filters for $d=2$, $m=2$. HOGG: notes!}
-  \end{mdframed}
-\end{figure}
-
-Give a table saying how many filters there are for each $d$ and each $M$ at each $k$ and each parity.
+How do we make the linear functions? See \secref{sec:convolution}.
 
@@
 
+What do they look like? Like calculus operators! See figures.
 \begin{figure}
@@
 
+It is interesting that the convolution operations look (in many cases) very much like discrete versions of vector calculus operations, like div, grad, curl, and so on.
+That's interesting, and potentially useful.
+
 Note that we never take eigenvalues. But these are also real scalars in some cases.
 
-Note that pseudo-$k$-tensor filters only exist for certain triples of $(d, m, k)$.
+Note that $k$-$1$-tensor (pseudo) filters only exist for certain triples of $(d, m, k)$.
 
 \paragraph{Acknowlegements:}
-This project made use of Python, numpy, matplotlib, and cmastro.
+It is a pleasure to thank Teresa Huang (JHU) and Kate Storey-Fisher (NYU) for valuable discussions.
+This project made use of Python~3 \cite{python3}, numpy \cite{numpy}, matplotlib \cite{matplotlib}, and cmastro \cite{cmastro}.
+All code used for making the data and figures in this paper are available at \url{https://github.com/davidwhogg/GeometricConvolutions}.
 
+\bibliographystyle{plain}
+\raggedright
+\bibliography{ref}
 \end{document}

### 2022-08-29 paper/main.tex
How we find the invariant filters; figure captions

@@
 
-\paragraph{Numerical construction of invariant geometric convolution filters:}
-In the above we defined equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
-These, in turn, are constructed from products of the outputs of convolutions with invariant geometric filters.
-Here we show how to find all such filters by a group-averaging technique.
-
-Briefly, the idea is the following:
-We start with a set of independent $k'$-tensor (or $k'$-pseudotensor) convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M$ odd) and dimension $d$.
-There are typically $M^d\,d^{k'}$ independent filters in this set.
-We average the filters over all group operators $g\in B_d$, where $B_d$ is the group of 90-degree rotations and reflections.
-We perform a singular value decomposition of these group-averaged filters to find the $p$ independent orthogonal filters that span the space of equivariant $k'$-tensor (or $k'$-pseudotensor) convolution filters at side length $M=2\,m+1$ in $d$ dimensions.
+In practice, a complete set of invariant $k$-$p$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
+This works by building any set of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
+\begin{equation}
+    \bar{C}(a') = \frac{1}{|B_d|}\,\sum_{g\in B_d} g\cdot C(g^{-1}\cdot a') ~,
+\end{equation}
+where $|B_d|$ is the number of group elements.
+The averaged filters are then reduced to an ``eigen-set'' of orthogonal filters by the singular value decomposition.
+In detail we started this averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
+There are $M^d\,d^{k'}$ independent filters in this set.
 The results of these operations are shown for $d=2$ and $m\in[1, 2]$ in \figref{fig:filters23} and \figref{fig:filters25}.
@@
   \end{center}
-\caption{All the filters for $d=2$, $m=1$ ($M=3$). HOGG: notes!\label{fig:filters23}}
+\caption{All the filters for $d=2$, $m=1$ ($M=3$).
+Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero.
+The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations but they are.
+There are no invariant pseudoscalar filters available at $(d,m) = (2,1)$.
+Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor filters but multiplied by the identity tensor.\label{fig:filters23}}
   \end{mdframed}
@@
   \end{center}
-\caption{All the filters for $d=2$, $m=2$ ($M=5$). HOGG: notes!\label{fig:filters25}}
+\caption{All the filters for $d=2$, $m=2$ ($M=5$).
+Notes: Symbols and colors as in \figref{fig:filters23}.
+We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.\label{fig:filters25}}
   \end{mdframed}
@@
 
-There is a continuous set of symmetries that inspire the summation-notation rules. There are discrete symmetries that are exact for these square images. 
-One discussion point: you might be able to get back to a continuous symmetry group if you consider the image to be a VIEW of a large, continuous tensor field...?
+One of the strange aspects of this work is that there are two groups acting.
+One is a continuous group (the Euclidean group) of translations and rotations (and reflections).
+This group motivates the index summation rules and the other geometric operations in the functions we construct.
+The other group is a discrete group $G_d$ appropriate for $d$-cube lattices of points; it is a sub-group of the continuous group.
+This is the group we use to average the filters (to make invariant filters) and we use to define equivariance for functions of geometric images.
+There are other possible image representations---other than values in pixels---that might create more continuous concepts of images.
+For example, if the data are on the surface of the sphere (the sky, perhaps), it could be represented with tensor spherical harmonics, and be subject to transformations by a continuous rotation group.
 
-It is interesting that the convolution operations look (in many cases) very much like discrete versions of vector calculus operations, like div, grad, curl, and so on.
+But if we think in terms of a physical, generative model for imaging: Usually an image is ``taken'' of a continuous world.
+That is, the kinds of geometric images we consider in this work would mainly be discretized views or samples of a larger, continuous world.
+That world is subject to transformations by continuous symmetries, maybe even gauge symmetries.
+It remains an open question how to think about how to use the operations and symmetries of these large, continuous groups that govern the world being viewed, when the data are only discrete, finite views of that world.
+
+HOGG: It is interesting that the convolution operations look (in many cases) very much like discrete versions of vector calculus operations, like div, grad, curl, and so on.
 That's interesting, and potentially useful.

### 2022-08-29 notebooks/geometric.py
todo items

@@ -10,7 +10,8 @@ See the file `LICENSE` for more details.
 - David W. Hogg (NYU)
 
 ## To-do items:
-- Drop the k-tensor and geometric filter classes; there can be only one class.
+- Fix the norm() operations so they are makeable with index summations! Duh. sqrt(a_hij a_hij / d^(k-2)) maybe??
+- Drop the k-tensor and geometric filter classes; there can be only one class. Or have filter inherit from image with a few extras.
 - Move over to jax.
 - Make the geometric_filter inherit geometric_image to reduce repeated code.
 - Is it possible to make the data a big block (not a dictionary) but have the block addressable with keys()? I bet it is...?

### 2022-08-29 notebooks/geometric.py
figure tweaks; restructuring

@@ -99,10 +99,10 @@ class ktensor:
             nn = "scalar"
         if k == 1:
             nn = "vector"
-        if parity % 2 == 1:
+        if parity % 2 == 1 and k < 2:
             nn = "pseudo" + nn
         if k > 1:
-            nn = "${}$-".format(k) + nn
+            nn = "${}$-${}$-".format(k, parity) + nn
         return nn
 
     def __init__(self, data, parity, D):
@@ -538,13 +538,14 @@ def plot_nothing(ax):
     return
 
 def plot_filters(filters, names, n):
-    assert len(filters) <= n
-    bar = 10. # figure width in inches?
-    fig, axes = plt.subplots(1, n, figsize = (bar, 0.03 * bar + bar / n)) # magic
-    if n == 1:
-        axes = [axes, ]
-    plt.subplots_adjust(left=0.001, right=0.999, wspace=0.2/n,
-                        bottom=0.001, top=0.999-0.1, hspace=0)
+    m = max(1, np.ceil(len(filters) / n).astype(int))
+    assert len(filters) <= n * m
+    bar = 8. # figure width in inches?
+    fig, axes = plt.subplots(m, n, figsize = (bar, 1.15 * bar * m / n), # magic
+                             squeeze=False)
+    axes = axes.flatten()
+    plt.subplots_adjust(left=0.001/n, right=1-0.001/n, wspace=0.2/n,
+                        bottom=0.001/m, top=1-0.001/m-0.1/m, hspace=0.2/m)
     for i, (ff, name) in enumerate(zip(filters, names)):
         if ff.k == 0:
             plot_scalar_filter(ff, name, ax=axes[i])
@@ -552,7 +553,7 @@ def plot_filters(filters, names, n):
             plot_vector_filter(ff, name, ax=axes[i])
         if ff.k == 2:
             plot_tensor_filter(ff, name, ax=axes[i])
-    for i in range(len(filters), n):
+    for i in range(len(filters), n * m):
         plot_nothing(axes[i])
     return fig
 

### 2022-08-29 paper/main.tex

@@
 
-In practice, a complete set of invariant $k$-$p$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
+In practice, a complete set of invariant $k'$-$p'$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
 This works by building any set of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
@@
 \end{equation}
-where $|B_d|$ is the number of group elements.
-The averaged filters are then reduced to an ``eigen-set'' of orthogonal filters by the singular value decomposition.
-In detail we started this averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
-There are $M^d\,d^{k'}$ independent filters in this set.
+where $|B_d|$ is the number of group elements (8 for $d=2$ and 27 for $d=3$).
+In detail we took a systematic (not random) approach that initialized the averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
+There are $M^d\,d^{k'}$ independent filters in this initial set.
+The averaged filters are then reduced to an ``eigen-set'' of orthogonal, non-zero filters by the singular-value decomposition.
+The singular-value decomposition reduces the number of filters dramatically.
+The outputs of the singular-value decomposition can be normalized however seems appropriate.
+In detail we normalized to get unit tensor norms in the non-zero pixels.
 The results of these operations are shown for $d=2$ and $m\in[1, 2]$ in \figref{fig:filters23} and \figref{fig:filters25}.
-HOGG: Give a table saying how many filters there are for each dimension $d$ and each side length $M$ at each order $k$ and each parity $p$.
+HOGG: Give a table saying how many filters there are for each tuple of $(d,m,k',p')$.
+HOGG: Perhaps remind the reader that the code is available.
 \begin{figure}[tp]
@@
   \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter+1_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
-\includegraphics[width=\textwidth]{notebooks/filter+2_2_3.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-2_2_3.png}
+\includegraphics[width=\textwidth]{notebooks/filter_2_3.png}
   \end{center}
@@
 Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero.
-The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations but they are.
-There are no invariant pseudoscalar filters available at $(d,m) = (2,1)$.
-Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor filters but multiplied by the identity tensor.\label{fig:filters23}}
+The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations, but they are.
+There are no invariant pseudoscalar ($0$-$1$-tensor) filters available at $(d,m) = (2,1)$.
+Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor (scalar) filters but multiplied by the identity tensor; there is only one non-trivial invariant $2$-$0$-tensor filter.
+We don't show the $k'$-$p'$-filters at $k'>2$ because we don't know how to visualize them (even the $k'=2$ case is sketchy).\label{fig:filters23}}
   \end{mdframed}
@@
   \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-0_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter+1_2_5.png}\\
-\includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
+\includegraphics[width=\textwidth]{notebooks/filter_2_5.png}
   \end{center}

### 2022-09-21 notebooks/geometric.py
change plot aspect ratios

@@ -541,7 +541,7 @@ def plot_filters(filters, names, n):
     m = max(1, np.ceil(len(filters) / n).astype(int))
     assert len(filters) <= n * m
     bar = 8. # figure width in inches?
-    fig, axes = plt.subplots(m, n, figsize = (bar, 1.15 * bar * m / n), # magic
+    fig, axes = plt.subplots(m, n, figsize = (bar, 1.2 * bar * m / n), # magic
                              squeeze=False)
     axes = axes.flatten()
     plt.subplots_adjust(left=0.001/n, right=1-0.001/n, wspace=0.2/n,
@@ -792,11 +792,11 @@ def plot_images(images):
       a LaTeX expression.
     """
     nim = len(images)
-    n = np.floor(np.sqrt(nim)).astype(int)
+    n = np.ceil(np.sqrt(nim)).astype(int)
     m = np.ceil(nim / n).astype(int)
     print(len(images), n, m)
     bar = 10. # inches?
-    fig, axes = plt.subplots(m, n, figsize = (bar, 0.2 * m + bar * m / n)) # magic
+    fig, axes = plt.subplots(m, n, figsize = (bar, 0.3 * m + bar * m / n)) # magic
     axes = np.atleast_1d(axes).flatten()
     plt.subplots_adjust(left=0.001, right=0.999, wspace=0.2/n,
                         bottom=0.001, top=0.999-0.07/m, hspace=0.2/m)

### 2022-09-21 paper/main.tex

@@
 %   -- another option: can we produce a characterization of all equivariant functions (maybe based on invariants or these filters) that don't require the use of non-linearities.
-%   -- another option: emulate cosmological simulations.
-% - change N to N^3 in the appropriate locations; already done??
 % - Understand what functions can and cannot be expressed on this model 
-% - Extend it to non-linear maps (eg, by multiplications and contractions)
 % - Should we be representing images in fourier space? Ask Leslie!
@@
 % - Some definitions begin with "given" some with "let". Should we try to be more consistent?
-
-% -- if the size of the filter is large enough, are we universal???? or can we express all polynomial functions? is there a version of Stone-Weierstrass for this?
-% --- add the identity and levi civita in the possible output tensors
-% --- define reordering of the indices
+% add zeroth order monomials (constants, very few)
 % --- we don't need to be in the torus and we don't need to be square (or cubical)
-% --- generalize convolution to rectangular non-toroidal images
-
-%-- is it the same to do convolutions + products + convolutions + products
-
-%-- a counting argument may tell us that all the linear functions that are equivariant  (Molien's formula)
 
@@
 
-% write the levi civita contraction
-% write the procedure of which we make all linear functions from k tensor images to k' tensor images of the same size. 
-% do we include a pooling move?
+% SOLEDAD: audit the levi civita contraction
+%GIVE THE METHOD A NAME AND RUN THE NAME THROUGH THE DOCUMENT
+%reformat the monomial figures to look like equations
+%visualize the group $B_d$ at d=2
+%make an introduction figure showing examples of geometric filters
 
-% how do we enumerate all tensor functions
+% - HOGG: Change the aspect ratios of the figures so they can be put into talk slides :)
 
-% 1, 6, 22, 49, 87 d=3, n=1,3,7,9 (11N^2-24N+21)/8
+%pixel positions in images are 1-0-tensors and they transform as 1-0-tensors
+
+%change the notation from k-p to k_p
+%should we use the upper index notation??
+%t
 
@@
 }
+\newenvironment{hoggfigure}{%
+  \begin{figure}[tp]%
+    \begin{mdframed}%
+    \color{captiongray}}{%
+    \end{mdframed}%
+  \end{figure}}
 
@@
 Or they might be the inputs or outputs of a simulation, in which the initial conditions or fields are specified on a regular grid.
+For examples, see \figref{fig:examples} and the caption thereto.
 Any lattice or grid of vectors or tensors can be seen as a generalization of the concept of an image, in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor) in each pixel.
@@
 Thus there is a need for machine learning methods designed for \emph{geometric images}---lattices or grids of scalars, vectors, and tensors.
-There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines (for example, CITE).
+There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines (HOGG CITE THINGS).
+\begin{hoggfigure}
+  \begin{center}
+    \begin{minipage}[b]{2.5in}\noindent%
+      \includegraphics[width=\textwidth]{jpegPIA18916.jpeg}\\
+      \textsl{(a)} temperature and polarization
+    \end{minipage}
+    \begin{minipage}[b]{2in}\noindent%
+      \includegraphics[width=\textwidth]{key_figures_197.pdf}\\
+      \textsl{(b)} salinity and current
+    \end{minipage}\\[2ex]
+    \begin{minipage}[b]{2.5in}\noindent%
+      \includegraphics[width=\textwidth]{applsci-09-02108-g001.png}\\
+      \textsl{(c)} temperature
+    \end{minipage}
+    \begin{minipage}[b]{1.8in}\noindent%
+      \includegraphics[width=\textwidth]{The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature.png}\\
+      \textsl{(d)} vorticity
+    \end{minipage}\\[2ex]
+    \begin{minipage}[b]{1.9in}\noindent%
+      \includegraphics[width=\textwidth]{FbAjqkkWAAANqFc.jpeg}\\
+      \textsl{(e)} intensity and polarization
+    \end{minipage}
+    ~~\begin{minipage}[b]{2.5in}\noindent%
+      \includegraphics[width=\textwidth]{41524.pdf}\\[1ex]
+      \textsl{(f)} stress tensor
+    \end{minipage}
+  \end{center}
+  \caption{Examples of geometric images in the natural sciences.
+  \textsl{(a)}~A visualization of a temperature map and a polarization map from the ESA \textsl{Planck} Mission \cite{planck}.
+  The color map shows a temperature field (a scalar or 0-0-tensor) on the sphere, and the whiskers show the principal eigenvector direction of a 2-0-tensor field in two dimensions.
+  In detail the underlying data are represented on a pixel grid (healpixel \cite{healpix}) on the sphere.
+  \textsl{(b)}~Some other stuff from \url{https://climatedataguide.ucar.edu/climate-data/godas-ncep-global-ocean-data-assimilation-system} HOGG EXPLAIN AND CITE.
+  \textsl{(c)}~Some other stuff from \url{https://www.mdpi.com/2076-3417/9/10/2108/htm} HOGG EXPLAIN AND CITE.
+  \textsl{(d)}~From \url{https://www.researchgate.net/figure/The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature_fig3_26635879}.
+  \textsl{(e)}~Ask Jordy Davelaar (and thank him).
+  \textsl{(f)}~From \url{https://www.nature.com/articles/s41524-019-0234-8}.
+  \label{fig:examples}}
+\end{hoggfigure}
 
@@
 
-\begin{definition}[$k$-$p$-tensors] \label{def.tensors}
+\begin{definition}[$k$-$p$-tensors]\label{def.tensors}
 We say that $v\in \mathbb R^d$ is a 1-$p$-tensor if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
@@
 We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are known as \emph{pseudovectors}, the 0-0-tensors are known as \emph{scalars}, 0-1-tensors are known as \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are known as \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
+In general, any $k$-$p$-tensor $a$ can be written as a sum of outer products of order-$1$ ($k=1$) tensors (vectors and pseudovectors), where each term in the sum is an outer product of $k$ order-$1$ tensors and the parity $p$ is the sum of the parities of the input order-$1$ tensors (mod 2, perhaps).
 
@@
 
-SOLE: If the next definition is going to use ``the torus'' then we should say a few words about that up-front, or above in the image definition.
-
 \begin{definition}[geometric convolution]
@@
 \end{definition}
-This definition is on the torus. If you are not on the torus, just pad the image out with zero tensors of the corresponding order and parity.
+This definition is on the torus. If you the image is not on the torus, we can pad the image out with zero tensors of the corresponding order and parity.
 
@@
 
-We conjecture that these rules will generate all possible second-degree monomial functions of the input image (SOLE SUBJECT TO SOME CONDITIONS) that meet the output requirements.
+We conjecture that these rules will generate all possible second-degree monomial functions of the input image as long as the filter size is sufficiently large (BEN, is this correct?).
 A polynomial function of degree 2 can be constructed by linear combinations of these second-degree monomial functions along with all the linear functions that also produce $k'$-$p'$-tensor images of side length $N'$ (plus perhaps constant images).
@@
 
-SOLE: THIS PARAGRAPH OR THESE IDEAS GO ELSEWHERE, I THINK.
 The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
@@
 
-Introduce the discrete group. Why discrete? SOLE: SEE TEXT BELOW FOR DEFINITIONS.
+Convolutional filters are widely used in machine learning for scalar images.
+The fundamental property of these operators are that they are translation equivariant, and that every translation equivariant linear function can be expressed as a convolution with a fixed filter (as long as the filter can be set to be as large as the image).
+The same property holds for geometry images.
 
-Define equivariance.
+\begin{theorem}
+A function $f:\mathcal A_{N,d,k,p}\to A_{N,d, k'', p''}$ is a linear translation equivariant linear function if and only if it is the convolution with a geometric filter $k'$-$p'$-tensor filter $C$ with possibly sidelength $N$, where $k+k'=k''$ and $p+p'=p''$.
+\end{theorem}
 
-Prove translation invariance.
+\begin{proof}
+TO DO.
+\end{proof}
 
-Prove things about invariant filters and equivariant linear functions.
+In addition to translation symmetries, we want to consider other natural symmetries occurring in the application domains where vectors and tensors arise.
+Ideally we would like to apply continuous rotations to the images, but their discretized nature makes it challenging.
+Recent work has proposed an approach to approximate continuous symmetries \cite{} that require interpolating and cropping the images.
+Here, for simplicity we focus on discrete rotations as in \cite{}, and we extend the group action to the geometric objects in the images.
 
@@
 
-\begin{definition}[Action of $B_d$ on $k$-$p$-tensors]
-\begin{align}
-g\cdot(v_1 \otimes \ldots \otimes v_k) &= \det(g)^p\,(g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k)
-\end{align}
-And extend linearly
-\end{definition}
-
-This may seem familiar: The $k$-tensor and $k$-pseudotensor were defined by how they transform under the orthogonal group $O(d)$, and our group $B_d$ is a subgroup of the orthogonal group.
+Not that the action of $B_d$ is a subgroup of $O(d)$, therefore the action of $B_d$ on $k$-$p$-tensors is the restriction of the action in Definition \ref{def.tensors} to the subgroup.
 
@@
 In practice, a complete set of invariant $k'$-$p'$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
-This works by building any set of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
+This works by, for each setting of order $k'$ and parity $p'$, building a collection of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
 \begin{equation}
@@
 where $|B_d|$ is the number of group elements (8 for $d=2$ and 27 for $d=3$).
-In detail we took a systematic (not random) approach that initialized the averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
+In detail we took a systematic (not random) approach that, for each setting of $(k',p')$, initialized the averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
 There are $M^d\,d^{k'}$ independent filters in this initial set.
@@
 HOGG: Perhaps remind the reader that the code is available.
-\begin{figure}[tp]
-  \begin{mdframed}
-  \color{captiongray}
+\begin{hoggfigure}
   \begin{center}
@@
   \end{center}
-\caption{All the filters for $d=2$, $m=1$ ($M=3$).
+\caption{All the filters for $d=2$, $m=1$ ($M=3$), $k'\in [0,1,2]$.
 Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero.
@@
 Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor (scalar) filters but multiplied by the identity tensor; there is only one non-trivial invariant $2$-$0$-tensor filter.
-We don't show the $k'$-$p'$-filters at $k'>2$ because we don't know how to visualize them (even the $k'=2$ case is sketchy).\label{fig:filters23}}
-  \end{mdframed}
-\end{figure}
-\begin{figure}[tp]
-  \begin{mdframed}
-  \color{captiongray}
+We don't show the $k'$-$p'$-filters at $k'>2$ because we don't know how to visualize them (even the $k'=2$ case is sketchy).
+Note that the vector ($1$-$0$-tensor) filters look like pure divergence and the pseudovector ($1$-$1$-tensor) filters look like pure curl.\label{fig:filters23}}
+\end{hoggfigure}
+\begin{hoggfigure}
   \begin{center}
@@
   \end{center}
-\caption{All the filters for $d=2$, $m=2$ ($M=5$).
+\caption{All the filters for $d=2$, $m=2$ ($M=5$), $k'\in [0,1]$.
 Notes: Symbols and colors as in \figref{fig:filters23}.
-We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.\label{fig:filters25}}
-  \end{mdframed}
-\end{figure}
+We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.
+At $(d,m)=(2,2)$ a pseudoscalar filter appears.
+Again, the vector and pseudovectors look like pure divergence and pure curl, respectively.\label{fig:filters25}}
+\end{hoggfigure}
+
+Define equivariance.
+
+Prove translation invariance.
+
+Prove things about invariant filters and equivariant linear functions.
 
@@
 What do they look like? Like calculus operators! See figures.
-\begin{figure}
-  \begin{mdframed}
-  \color{captiongray}
+\begin{hoggfigure}
   \begin{center}
@@
   \caption{A scalar image $s$ and convolutions with four chosen filters.}
-  \end{mdframed}
-\end{figure}
+\end{hoggfigure}
 
 Now show contractions of products of the above on those images.
-
-\begin{figure}
-  \begin{mdframed}
-  \color{captiongray}
+\begin{hoggfigure}
   \begin{center}
@@
   \caption{All unconvolved second-degree monomials of $s$, given four chosen filters. For the 2-tensor outputs (order $k=2$), we traced them back to scalars.}
-  \end{mdframed}
-\end{figure}
+\end{hoggfigure}
 
@@
 
+We could have added the determinant operation and eigenvalues.....
+
 \paragraph{Acknowlegements:}
-It is a pleasure to thank Teresa Huang (JHU) and Kate Storey-Fisher (NYU) for valuable discussions.
+It is a pleasure to thank
+Leslie Greengard (Flatiron),
+Teresa Huang (JHU), and
+Kate Storey-Fisher (NYU) for valuable discussions.
 This project made use of Python~3 \cite{python3}, numpy \cite{numpy}, matplotlib \cite{matplotlib}, and cmastro \cite{cmastro}.

### 2023-03-20 paper/main.tex
Updates from Overleaf

@@
-% Copyright 2022 the authors
+% Copyright 2022, 2023 the authors
 
@@
 % -----------
-% - What are our GOALS!!?
-%   -- one option: this formulation can express all equivariant differential operators discretized to a cubical d-torus. 
-%   -- another option: can we produce a characterization of all equivariant functions (maybe based on invariants or these filters) that don't require the use of non-linearities.
+
+% - What are our GOALS!!? is it clearly stated in the abstract and introduction?
 % - Understand what functions can and cannot be expressed on this model
-% - Should we be representing images in fourier space? Ask Leslie!
-% - Understand the literature
+
+% MARIA TERESA:
+% - Understand the following very specific things in the literature:
 %   -- Discretized vector calculus and differential geometry.
@@
 %   -- Understand how this formulation compares with steerable CNNs
-%   -- What is known about what functions CNNs can represent? all linear translation equivariant functions can be written as convvolutions if the filters are as large as the image (simple observation)
-% what about non-linear? is there a polynomial basis here? 
+%   -- What is known about what functions CNNs can represent?
+all linear translation equivariant functions can be written as convvolutions if the filters are as large as the image (simple observation) what about non-linear?
+%  -- comment on the fact that Cohen and Kondor showed that in homogeneous spaces all linear equivariant maps are convolutions, say in what sense we generalize the result, by combining it with equivariant functions on tensors and vectors and scalars can be written in terms of scalars.
 %  -- Define "local" functions, is there a characterization of functions that can be written as convolutions with small filters?
-% - There is a one-indexing and zero-indexing issue. Should we be consistent?
+% -- in the introduction we should talk about the geometric principle
+
+% SOLEDAD:
 % - Some definitions begin with "given" some with "let". Should we try to be more consistent?
-% add zeroth order monomials (constants, very few)
-% --- we don't need to be in the torus and we don't need to be square (or cubical)
+% Define the constants like the dirac delta and the levi civita,
+% - COMMENT: In the same way that we can take nonlinear functions of scalars, we can take ODD nonlinear functions of pseudoscalars (with no bias term permitted!).
+% - write our contributions (there is a placeholder in the text).
 
-%-- BEN: compute Molien's formula for this (translations semidirect D_4) action
-% 1/|G|\sum_{g\in G} det(I-\phi(g)t)^{-1}
+% WILSON:
+% - GIVE THE METHOD A NAME AND RUN THE NAME THROUGH THE DOCUMENT
+% - Make a new figure that shows what happends when the filters are applied to a geometric image. Make it look like an equation or a rebus.
+% - visualize the group $B_d$ at d=2
+% - comment that \bar i is a \tensor{1}{+}
+% - look at the different definitions of the action of g and see that they are consistent
+% - add explicit comments after definitions, where appropriate. Use the "remark" environment. Look at Hogg's remarks for examples.
+% - add a description of the numerical example problem
+% - describe the possible architectures and a diagram
 
-% SOLEDAD: audit the levi civita contraction
-%GIVE THE METHOD A NAME AND RUN THE NAME THROUGH THE DOCUMENT
-%reformat the monomial figures to look like equations
-%visualize the group $B_d$ at d=2
-%make an introduction figure showing examples of geometric filters
-
-% - HOGG: Change the aspect ratios of the figures so they can be put into talk slides :)
-
-%pixel positions in images are 1-0-tensors and they transform as 1-0-tensors
-
-%change the notation from k-p to k_p
-%should we use the upper index notation??
-%t
+% - HOGG:
+% - comment on alias alibi in the style of Weyl, both in the Introduction, and at the definition of the group action on a tensor.
+% - Should we make the contravariant / covariant distinction here, or at least comment on it?
+Probably yes.
+should we use the upper index notation??
+We should at least explain why people do that.
+Maybe we should.
+% - Equation (1) is the ugliest thing in the Universe at present.
 
@@
 \usepackage[utf8]{inputenc}
-\usepackage[letterpaper]{geometry} % because Overleaf is weird.
+\usepackage[letterpaper]{geometry} 
 \PassOptionsToPackage{hyphens}{url}
@@
 \usepackage{amsthm}
+\usepackage{verbatim}
+\usepackage{physics}
 
@@
 % theorem, definition and so on
-\theoremstyle{plain}
+\theoremstyle{definition}
 \newtheorem{definition}{Definition}
@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{proposition}{Proposition}
+\newtheorem*{remark}{Remark}
 
 % math macros
+\newcommand{\tensorname}[2]{{#1}_{(#2)}}
+\newcommand{\tensor}[2]{$\tensorname{#1}{#2}$-tensor}
 % HOGG SAY: I DON'T KNOW HOW TO USE AMSMATH \choose AND I'M ON A PLANE. FEEL FREE TO FIX THIS.
@@
 \newcommand{\figref}[1]{\figurename~\ref{#1}}
+\newcommand{\wgg}[1]{\textcolor{purple}{#1}}
 
@@
 \pagestyle{myheadings}
-\markboth{foo}{Hogg, Blum-Smith, \& Villar / Geometric filters for geometric images}
+\markboth{foo}{Gregory, Hogg, Blum-Smith, Arias, \& Villar / Geometric Image CNNs}
 \newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
@@
 Convolution-based nonlinear functions on tensor fields and tensor images}
-\author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
+\author{%
+Wilson~Gregory,
+David~W.~Hogg,
+Ben~Blum-Smith,
+\\
+Maria~Teresa~Arias,
+Soledad~Villar
+}
 \date{}
@@
 In natural-science domains, image-like data sets can have a vector (velocity, say), or tensor (polarization, say), or pseudovector (magnetic field, say), or any combination in each pixel.
-We provide a construction for general convolution-based functions of images or lattices
+Here we provide a construction for general convolution-based functions of images or lattices
 containing scalars, vectors, and tensors, producing scalar, vector, and tensor outputs.
@@
 
-Contemporary natural science teems with data sets that are images or lattices or grids of geometric objects.
+Contemporary natural science is replete with data sets that are images, lattices, or grids of geometric objects.
 These might be observations of intensities (scalars), magnetic fields (pseudovectors), or polarizations (2-tensors) on a surface or in a volume.
-Or they might be the inputs or outputs of a simulation, in which the initial conditions or fields are specified on a regular grid.
-For examples, see \figref{fig:examples} and the caption thereto.
-Any lattice or grid of vectors or tensors can be seen as a generalization of the concept of an image, in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor) in each pixel.
+They might be the inputs or outputs of a simulation where the initial conditions or fields are specified on a regular grid; see \figref{fig:examples} for some examples.
+Any lattice of vectors or tensors can be seen as a generalization of the concept of an image in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor).
 These objects are \emph{geometric} in the sense that they are defined in terms of their transformation properties under geometric operators such as rotation, translation, and reflection.
 Thus there is a need for machine learning methods designed for \emph{geometric images}---lattices or grids of scalars, vectors, and tensors.
-There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines (HOGG CITE THINGS).
+There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines.
 \begin{hoggfigure}
@@
   \caption{Examples of geometric images in the natural sciences.
-  \textsl{(a)}~A visualization of a temperature map and a polarization map from the ESA \textsl{Planck} Mission \cite{planck}.
-  The color map shows a temperature field (a scalar or 0-0-tensor) on the sphere, and the whiskers show the principal eigenvector direction of a 2-0-tensor field in two dimensions.
-  In detail the underlying data are represented on a pixel grid (healpixel \cite{healpix}) on the sphere.
-  \textsl{(b)}~Some other stuff from \url{https://climatedataguide.ucar.edu/climate-data/godas-ncep-global-ocean-data-assimilation-system} HOGG EXPLAIN AND CITE.
-  \textsl{(c)}~Some other stuff from \url{https://www.mdpi.com/2076-3417/9/10/2108/htm} HOGG EXPLAIN AND CITE.
-  \textsl{(d)}~From \url{https://www.researchgate.net/figure/The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature_fig3_26635879}.
-  \textsl{(e)}~Ask Jordy Davelaar (and thank him).
-  \textsl{(f)}~From \url{https://www.nature.com/articles/s41524-019-0234-8}.
+  \textsl{(a)}~A visualization of a temperature map and a polarization map from the ESA \textsl{Planck} Mission \cite{planck2015}.
+  The color map shows a temperature field (a scalar or \tensor{0}{+}) on the sphere, and the whiskers show the principal eigenvector direction of a \tensor{2}{+} field in two dimensions.
+  In detail the underlying data are represented on a pixel grid (healpixel \cite{healpix}) on the sky (a 2-sphere).
+  \textsl{(b)}~Two-dimensional maps of ocean current (shown with arrows; a vector or \tensor{1}{+} field) and ocean salinity (shown with color; a scalar or \tensor{0}{+} field) at a depth of 5\,m \cite{climatedataguide}.
+  \textsl{(c)}~A three-dimensional map of temperature (a scalar or \tensor{0}{+} field) based on sensors distributed throughout the volume of a granary \cite{granary}.
+  \textsl{(d)}~A two-dimensional map of potential vorticity (a pseudoscalar or \tensor{0}{-} field) in the Earth's atmosphere, measured for the purposes of predicting storms \cite{potentialvorticity}.
+  \textsl{(e)}~Two-dimensional maps on the sky of intensity $I$ and the three independent components $Q, U, V$ of the electromagnetic polarization \tensor{2}{+}, from a simulation of a jet outflow from an accreting black hole [Davelaar et al, in preparation].
+  \textsl{(f)}~Components of the three-dimensional stress tensor (a \tensor{2}{+} field) in a diamond anvil cell, which is used to study the behavior of samples at exceedingly high pressures \cite{diamond}.
   \label{fig:examples}}
@@
 
-%For a concrete example close to one of our hearts (DWH's), contemporary cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
-%For this reason, accurate, fast emulators might be of enormous economic and scientific value.
-%There are now quite a few successful projects implementing such emulators (CITE THINGS).
-%Importantly, the initial conditions for these simulations are tiny displacements (vectors) and velocities (vectors) relative to a cubical lattice on a 3-torus.
-%That is, the initial conditions can be described as a 3-dimensional image of vectors on a regular grid.
-%For another concrete example, astronomical imaging data taken of the interstellar medium can be represented as rectangular lattices of measurements of temperature (scalar), optical depth (scalar), and induced polarization of starlight (a 2-tensor).
-%Here the data can be described as a 2-dimensional image of scalars and tensors on a regular grid.
-
-At the present day, the go-to tools for machine learning with images are convolutional neural networks (CNNs; CITE) and their many descendents (for example, CITE).
-These methods are designed to work on one- or few-channel images in which, at the bottom (early) layers of the network, the operations involve image convolutions with learned filters followed by the application of pointwise nonlinearities.
+At the present day, the go-to tools for machine learning with images are convolutional neural networks (CNNs; \cite{lecun1989backpropagation}) and their many descendents (for example, CITE).
+These methods are designed to work on one- or few-channel images in which the early layers of the network involve image convolutions with learned filters followed by the application of pointwise nonlinearities.
 In typical contexts, the channels of multi-channel input images will be something like the red, green, and blue (or cyan, magenta, yellow, and black) channels of a color image; these can be combined arbitrarily in the layers of the CNN.
-When these CNN-based tools are applied to lattices of vectors (say), typically the components of the vectors are treated just as channels of the input image (for an example, see CITE) and then everything proceeds as with multi-channel color images.
-This is not good!
+When these CNN-based tools are applied to lattices of vectors for example, typically the components of the vectors are treated just as channels of the input image (for an example, see CITE) and then everything proceeds as with multi-channel color images.
+This is not good! %WILSON: I don't love this. HOGG says: You are the first author. Be bold; edit.
 
@@
 The fundamental observation inspiring this work is that when the components of vectors and tensors are acted on with arbitrary functions, the geometric structure of these objects is destroyed.
-There are strict rules, dating back to the invention of differential geometry (CITE RICCI++), about how geometric objects can be combined to produce new geometric objects, consistent with coordinate freedom and transformation rules.
-In previous work (CITE VILLAR++, YAO++) we have capitalized on these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
-Here we use these rules to create a comprehensive set of tools that are universally approximating for functions that take geometric images as input and produce geometric labels (or geometric-image labels) as output.
+There are strict rules, dating back to the early days of differential geometry \cite{ricci}, about how geometric objects can be combined to produce new geometric objects, consistent with coordinate freedom and transformation rules (these rules constitute a theme of \cite{mcp}).
+In previous work \cite{villar2021scalars, villar2022dimensionless, yao2021simple} we have capitalized on these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
+Here we use these rules to create a comprehensive set of tools that are universally approximating for functions that take geometric images as input and produce geometric images as output.
 
@@
 After all, even standard images are measurements of intensity (a scalar) at a regular grid of points on a two-dimensional surface.
-If the labels we seek are also scalars, it will behoove us to obey the rules of geometry.
+If the labels we seek are also scalars, it behooves us to obey the rules of geometry.
 
@@
 $1$-tensor objects are called vectors and $0$-tensor objects are called scalars.
-There are also odd-parity versions of all these (pseudovectors, pseudoscalars, and pseudotensors), and parity-changing contractions using the Levi-Civita symbol, so in what follows we will define $k$-$p$-tensors that have $k$ indices and a parity $p\in [0, 1]$.
+There are also odd-parity versions of all these (pseudovectors, pseudoscalars, and pseudotensors), and parity-changing contractions using the Levi-Civita symbol, so in what follows we will define \tensor{k}{p}s that have $k$ indices and a parity $p\in\{-1,+1\}$, sometimes denoted ``$-$'' and ``$+$'' below.
 Two objects can only be added or subtracted if they have the same order $k$ and parity $p$.
-This will all be made precise below.
-But in short, these rules define objects that can be given transformation rules under rotation and reflection such that functions made of these operations are coordinate free, or equivariant to any change of coordinate system (including curvilinear coordinate transformations).
+These rules define objects that can be given transformation rules under rotation and reflection such that functions made of these operations are coordinate free, or equivariant to any change of coordinate system, including curvilinear coordinate transformations.
 
@@
 This equivariance, it turns out, is very easy to enforce, even for nonlinear functions of geometric images, provided that we compose our nonlinear functions from simple geometric operations, including our geometric generalization of convolution.
-When we enforce this equivariance, the convolution filters that appear look very much like the differential operators that appear in discretizations of vector calculus, which is perhaps not surprising.
-
-\paragraph{Related work:}
+When we enforce this equivariance, the convolution filters that appear look very much like the differential operators that appear in discretizations of vector calculus.
 
 \paragraph{Our contribution:}
+SOLEDAD: SUMMARY HERE. 
+To see a comparison of our work with other methods already studied, see \sectionname~\ref{sec:related}
+For a discussion of prior work on group equivariant machine learning, see \sectionname~\ref{sec:related2}.
 
-\section{Geometric objects and geometric images}\label{sec:geometric}
+\section{Non-linear functions of geometric images}
 
-In this \sectionname{} we define the geometric objects and geometric images that we use to generalize classical images in scientific contexts.
+We define the geometric objects and geometric images that we use to generalize classical images in scientific contexts in \sectionname{} \ref{sec:geometric} and \sectionname{} \ref{sec:convolution}, and we define the geometric functions on them in \sectionname{} \ref{sec:nonlinear}.
 The main point is that the channels of geometric images---which will be like the components of vectors and tensors---are not independent.
-There is an underlying action by the orthogonal group that rules what are the allowed objects and transformations in this space.
+There is an underlying action by the orthogonal group that rules what are the allowed objects and transformations in this space that we describe in \sectionname{}~\ref{sec:equivariant}.
 
-We first fix $d$, the dimension of the space (typically 2 or 3 but it could be arbitrary). The geometric objects are vectors and tensors. 
-The orthogonal group $O(d)$ can act on $v \in \mathbb R^d$ in the following ways:
+\subsection{Geometric objects}\label{sec:geometric}
+
+We first fix $d$, the dimension of the space (typically 2 or 3 but it could be arbitrary).
+The geometric objects are vectors and tensors. 
+The orthogonal group $O(d)$ is the space of isometries of $\mathbb R^d$ that fix the origin. 
+%It is represented by matrices $Q\in \mathbb R^{d\times d}$ such that $Q^\top Q = I$. 
+It acts on vectors and pseudovectors $v \in \mathbb R^d$ in the following way:
 \begin{align} \label{eq.action}
-    g\cdot v = \det(M(g))^p M(g) v 
+    g\cdot v = \det(M(g))^{\frac{1-p}{2}}\,M(g)\,v
 \end{align}
-where $g\in O(d)$, $M(g)\in \mathbb R^{d\times d}$ is the standard matrix representation of $g$, and $p\in\{0,1\}$ is the parity of the action.
-If $p=0$ we obtain the standard $O(d)$ action on $\mathbb R^d$ \emph{vectors}.
-If $p=1$ we obtain what in physics is commonly known as the action of $O(d)$ on \emph{pseudovectors}.
+where $g\in O(d)$, $M(g)\in \mathbb R^{d\times d}$ is the standard matrix representation of $g$ (i.e. $M(g)^\top M(g) = I$) and $p\in\{-1,+1\}$ is the parity of $v$.
+If $p=+1$ we obtain the standard $O(d)$ action on $\mathbb R^d$ \emph{vectors}.
+If $p=-1$ we obtain the $O(d)$ action on what in physics are known as \emph{pseudovectors}.
 In this context the objects are defined by the actions that they carry.
 
-\begin{definition}[$k$-$p$-tensors]\label{def.tensors}
-We say that $v\in \mathbb R^d$ is a 1-$p$-tensor if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
-If $v_i$ is a 1-$p_i$-tensor then $T:=v_{1}\otimes\ldots \otimes v_k$ is a rank-1 $k$-$p$-tensor, where $p=\sum_{i=1}^k p_i \mod 2$ and the action of $O(d)$ is defined as
+\begin{definition}[\tensor{k}{p}s]\label{def.tensors}
+We say that $v\in \mathbb R^d$ is a \tensor{1}{p} if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
+If $v_i$ is a \tensor{1}{p_i} then $T:=v_{1}\otimes\ldots \otimes v_k$ is a rank-1 \tensor{k}{p}, where $p=\prod_{i=1}^k p_i$ and the action of $O(d)$ is defined as
 \begin{align}
@@
 \end{align}
-Higher rank $k$-$p$-tensors are defined as linear combinations of rank-1  $k$-$p$-tensors where the action of $O(d)$ is extended linearly.
-\end{definition}
-Typically we would make the parity $p$ be a bit (either 0 for even parity or 1 for odd parity) but in principle it can be any integer.
-
-We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are known as \emph{pseudovectors}, the 0-0-tensors are known as \emph{scalars}, 0-1-tensors are known as \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are known as \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
-In general, any $k$-$p$-tensor $a$ can be written as a sum of outer products of order-$1$ ($k=1$) tensors (vectors and pseudovectors), where each term in the sum is an outer product of $k$ order-$1$ tensors and the parity $p$ is the sum of the parities of the input order-$1$ tensors (mod 2, perhaps).
-
-\begin{definition}[summation notation]
-In this notation, tensor products are written in component form, and repeated indices are summed over.
-In this notation, the product of two 2-tensors (represented as two $d\times d$ matrices $A$ and $B$) is written as
-\begin{equation}
-    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} ~,
-\end{equation}
-where $[A]_{i,k}$ (for example) is the $i,k$ element of matrix $A$, the repeated index $k$ is implicitly summed over the integers from 1 to $d$.
-This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice).
+Higher rank \tensor{k}{p}s are defined as linear combinations of rank-1 \tensor{k}{p}s where the action of $O(d)$ is extended linearly.
+The set of \tensor{k}{p}s in $d$ dimensions is denoted $\mathcal{T}_{d,k,p}$.
 \end{definition}
 
+\begin{remark}[terminology and notation]
+The parity $p$ is a signed bit, either $+1$ for even parity or $-1$ for odd parity.
+Note that we call $k$ the \emph{order} of the \tensor{k}{p}, rather than the ``rank'' of the tensor as some other sources do.
+Using rank for this property is inaccurate because there can be \tensor{2}{+}s of rank 1 (like those we use in Definition~\ref{def.tensors}).
+\end{remark}
+
+\begin{remark}[odd-parity objects]
+It is possible to replace any odd-parity ($p=-1$) object with an even-parity object with the same physical content by multiplying the odd-parity object by the Levi-Civita tensor (see Definition~\ref{def:levicivita}).
+When a \tensor{k}{p} is multiplied (outer product) by a Levi-Civita, the result is a \tensor{(k+d)}{-p}.
+Thus, it is possible to work without odd-parity objects at all, in principle.
+However, since odd-parity objects are of importance in physics and engineering (see \figurename~\ref{fig:examples}), we retain them.
+\end{remark}
+
+In physics the \tensor{1}{+}s (such as velocities) are known as \emph{vectors},
+the \tensor{1}{-}s (such as angular momenta) are known as \emph{pseudovectors},
+the \tensor{0}{+}s (such as rest masses) are known as \emph{scalars},
+the \tensor{0}{-}s (such as surface vorticities) are known as \emph{pseudoscalars},
+the \tensor{k}{-}s with $k\geq 2$ are known as \emph{pseudotensors},
+and finally the \tensor{k}{+}s with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
+In general, any \tensor{k}{p} can be written as a sum of outer products of order-$1$ ($k=1$) tensors (vectors and pseudovectors), where each term in the sum is an outer product of $k$ order-$1$ tensors and the parity $p$ is the product of the parities of the input order-$1$ tensors.
+
+%CONSTANT TENSORS
+We consider two special tensors that will be important for the definition of our models, the Kronecker delta and the Levi-Civita.
+
+\begin{definition}[Kronecker delta] The Kronecker delta ($\delta_{ij}$) is a \tensor{2}{+} represented by the identity matrix, namely the object with two indices $ij$ such that it has the value $+1$ when the two indices have the same value ($i=j$), and $0$ otherwise.
+\end{definition}
+
+\begin{definition}[Levi-Civita symbol]\label{def:levicivita}
+The Levi-Civita symbol in $d\geq 2$ is a \tensor{d}{-} $\epsilon_{i_1,\ldots, i_d}$ such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
+\end{definition}
+
+\begin{definition}[outer products of tensors]
+Given $a \in \mathcal{T}_{d,k,p}$ and $b \in \mathcal{T}_{d,k',p'}$, the outer product, denoted $a\otimes b$, is a tensor in $\mathcal{T}_{d,k+k',p\,p'}$ defined as $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$.
+\end{definition}
+
+\begin{definition}[contractions and Einstein summation notation]\label{def:contraction}
+We use Einstein summation notation where tensor products are written in component form, and repeated indices are summed over.
+For example, in this notation, the product of two \tensor{2}{+}s (represented as two $d\times d$ matrices $A$ and $B$) is written as
+\begin{equation}
+    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} = \sum_{k=1}^d [A]_{i,k}[B]_{k,j}
+\end{equation}
+where $[A]_{i,k}$ is the $i,k$ element of matrix $A$, and the sum from 1 to $d$ on repeated index $k$ is implicit in the middle expression.
+This notation works for tensor expressions of any order, provided that every index appears either exactly once, so it isn't summed over, or exactly twice, so it is summed over.
+\end{definition}
+
+\begin{remark}[lower and upper indices]
+In the real Einstein summation notation \cite{einstein}, or Ricci calculus \cite{ricci}, a distinction is made between lower and upper indices, which correspond to covariant and contravariant components.
+The pairs of indices that are summed always have one member of the pair an upper index and one member a lower index.
+We drop this upper/lower distinction here, because we will work with flat images that implicitly have a trivial metric, such that there is no numerical difference between covariant and contravariant component values for a given object.
+That said, there truly is a distinction (for example, if a spatial displacement is a contravariant vector, the gradient of a scalar function with respect to that spatial displacement is a covariant vector), so there might be advantages to reinstating this distinction.
+It would complexify method architectures (we discuss these in \secref{sec:architectures}) because the code would have to additionally track the covariance or contravariance of each index of each object.
+\end{remark}
+
 One consequence of the above definitions is that 
-$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-$p$-tensor if and only if (in summation notation)
+$T\in (\mathbb R^d)^{\otimes k}$ is a \tensor{k}{p} if and only if, in summation notation,
 \begin{align}
-    [g\cdot T]_{i_1,\ldots, i_k} = \det(M(g))^p\, [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}
+    [g\cdot T]_{i_1,\ldots, i_k} = \det(M(g))^{\frac{1-p}{2}}\,[T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}
 \end{align} for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
-For example, a 2-0-tensor is defined by its transformation property
+For example, a \tensor{2}{+} is defined by its transformation property
 $[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
@@
 
-We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
-Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel).
-Sometimes we will consider the image to be in a $d$-torus.
-Let $\mathcal T_{d,k,p}$ the set of $k$-$p$-tensors in $\mathbb R^d$. We define the geometric images as follows.
-\begin{definition}[geometric image]
-A geometric image is a function $A:[N]^d \to \mathcal T_{d,k,p}$, where $[N]=\{0,1,\ldots, N-1\}$.
-We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
-\end{definition}
-
-\section{Generalized convolutions on geometric images}\label{sec:convolution}
-
-In this \sectionname{} we generalize the notion of convolution to take geometric images as inputs and return geometric images as outputs.
-The idea is that the (presumably large) geometric image (containing $k$-$p$-tensors) is convolved with a (presumably small) geometric filter (also a geometric image, but containing $k'$-$p'$-tensors) to produce an output image that contains $(k+k')$-$(p+p')$-tensors, where each pixel is a sum of outer products.
-These outer products can be contracted down to lower-$k$ tensors using Kronecker or Levi-Civita contractions.
-
-\begin{definition}[outer products of tensors]
-If $a$ is a $k$-$p$-tensor and $b$ is a $k'$-$p'$-tensor, then the outer product $a\otimes b$ is a $(k+k')$-$(p+p')$-tensor with $[a\otimes b]_{i_1,\ldots,i_{k+k'}} = [a]_{i_1,\ldots,i_k}\,[b]_{i_{k+1},\ldots,i_{k+k'}}$.
-\end{definition}
-
-\begin{definition}[geometric convolution]
-Let $A$ be a $k$-$p$-tensor image on the $d$-torus with side length $N$.
-Let $C$ be a $k'$-$p'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
-The geometric convolution $A\ast C$ is a $(k+k')$-$(p+p')$-tensor image such that
-\begin{equation}
-    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
-\end{equation}
-where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
-\end{definition}
-This definition is on the torus. If you the image is not on the torus, we can pad the image out with zero tensors of the corresponding order and parity.
-
-Convolutions can be followed by Kronecker contractions, Levi-Civita contractions and index permutations.
-The Kronecker delta ($\delta_{ij}$) is the object with two indices $ij$ such that it has the value $+1$ when the two indices have the same value ($i=j$), and $0$ otherwise.
-Below we define the contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta, which returns a $(k-2)$-$p$-tensor.
-The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices (where $d$ is the spatial dimension) such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
-We define the contraction of a $k$-$p$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol, which returns a $(k-d+2)$-$(p+1)$-tensor, because the contraction sums over $d$ pairs of indices, and it changes the parity from even to odd or odd to even.
-For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
-For another example, the same 2-tensor can (in $d=3$ dimensions) be contracted with the Levi-Civita symbol to give the standard cross product
-$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a pseudovector.
-
-\begin{definition}[Kronecker contractions of tensors and geometric images]
-Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
-We define the $(\alpha,\beta)$ Kronecker contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$.
-In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$s are placed in the positions $\alpha$ and $\beta$:
-\begin{equation}
-[a^{(\alpha,\beta)}]_{i_1, \ldots, i_k \backslash i_\alpha,i_\beta} := \sum_{i_\alpha=1}^d \sum_{i_\beta=1}^d [a]_{i_1, \ldots, i_\alpha,\ldots,i_\beta,\ldots, i_k}\,\delta_{i_\alpha, i_\beta}
-\end{equation}
-When Kronecker contraction is applied to a geometric image, the contraction is applied to every $k$-$p$-tensor in the image identically.
-\end{definition}
-There are $\choose{k}{2}$ different index choices for this contraction, corresponding to all the possible index pairs among $k$ original indices.
-
-\begin{definition}[Levi-Civita contractions of tensors and geometric images]
-Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq d-1$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$.
-We define the Levi-Civita contraction of $a$ to be the $(k-d+2)$-$(p+1)$-tensor resulting from taking the outer product of
-$a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol.
-When a Levi-Civita contraction is applied to a geometric image, the contraction is applied to every $k$-$p$-tensor in the image identically.
-\end{definition}
-There are $(d-1)!\,\choose{k}{d-1}$ different index choices for the Levi-Civita contraction of a $k$-$p$-tensor.
-These choices correspond to the choices for an ordered list of $d-1$ indices chosen from among $k$ index choices.
+We can combine multiplication with Kronecker and Levi-Civita symbols with contractions to define relevant operations.
+For example the \tensor{2}{+} formed by the outer product of \tensor{1}{+}s $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a \tensor{1}{+} or scalar.
+For another example, the same \tensor{2}{+} can (in $d=3$ dimensions) be contracted with the Levi-Civita symbol to give the standard cross product
+$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a \tensor{1}{-} or pseudovector.
 
 \begin{definition}[permutations of tensor indices]
-Given a $k$-$p$-tensor $a$ and $\sigma\in S_k$ a permutation. We consider $a^\sigma$ the result of permuting the indices of $a$ according to $\sigma$:
+Given $a \in \mathcal{T}_{d,k,p}$ and permutation sequence $\sigma\in S_k$, the permutation of $a$ by $\sigma$, denoted $a^\sigma$, is:
 \begin{equation}
@@
 \end{equation}
-When index permutation is applied to a geometric image, the permutation is applied to every $k$-$p$-tensor in the image identically.
 \end{definition}
 
-In addition to these contractions and index permutations that act pixel-wise in geometric images, it is possible to change the image size.
-If $N$ is divisible by integer $f>1$, a geometric image of side length $N$ can be reduced in size to a geometric image of side length $N/f$ by the equivalent of convolution with a trivial $f\times f$ scalar filter (geometric filter of 0-0-tensors) operating with a stride of $f$.
-Similarly a geometric image of side length $N$ can be expanded in size to a side length $f\,N$ by application of an interpolation operator with stride $1/f$.
-\begin{definition}[pooling and un-pooling]
-foo and bar.
+\begin{comment}
+(WILSON PONDERS: I think I might just cut this below, I understand that physicists think about tensors in this way, but if we aren't using it I don't think it adds much.) There is an alternative definition of \tensor{k}{p}s in terms of geometric functions (see, for example, \cite{mcp} chapter 1):
+A \tensor{k}{+} can be thought of as representing a multilinear function of $k$ vectors (\tensor{1}{+}s) that produces a scalar (\tensor{0}{+}) output.
+For example, if $A$ is a \tensor{4}{+}, and $u,v,w,x$ are vectors (\tensor{1}{+}s) then (in summation notation)
+\begin{equation}
+    \rho = [A]_{ijk\ell}\,[u]_i\,[v]_j\,[w]_k\,[x]_\ell
+\end{equation}
+is a scalar (\tensor{0}{+}).
+\tensor{k}{-}s can be similarly defined in terms of input vectors and an output pseudoscalar.
+\end{comment}
+
+\subsection{Geometric images and convolutions}\label{sec:convolution}
+
+We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total).
+Each pixel contains a \tensor{k}{p} where $k$ and $p$ are the same for each pixel.
+Sometimes we will consider the image to be in a $d$-torus.
+Let $\mathcal T_{d,k,p}$ the set of \tensor{k}{p}s in $\mathbb R^d$. We define the geometric images as follows.
+
+\begin{definition}[geometric image]
+A geometric image is a function $A:[N]^d \to \mathcal T_{d,k,p}$, where $[N]=\{0,1,\ldots, N-1\}$.
+The set of geometric images is denoted $\mathcal{A}_{N,d,k,p}$.
+We will also consider \tensor{k}{p} images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
 \end{definition}
 
-Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-$p$-tensor image of side length $N$ to an output $k'$-$p'$-tensor image of side length $N'$ (where either $N$ is divisible by $N'$, or $N'$ is divisible by $N$).
-Here are the linear operations (linear in the input image) that are permitted:
-\begin{enumerate}
-    \item Perform a convolution and index contraction. For this step there are two qualitatively different options:
-        \begin{itemize}
-            \item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
-              Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
-              Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
-              There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
-            \item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
-              Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
-              Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
-              There are $$(d-1)!\,\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
-              Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
-              There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
-        \end{itemize}
-    \item Perform (generally after the convolution and contraction) any index permutation you like (provided that $k'>1$).
-      There are $k'!$ choices for the index permutation.
-    \item Perform (generally after all the above) a pooling or unpooling step to get the image to side length $N'$.
-\end{enumerate}
-The objects obtained by this algorithm may not be linearly independent.
+In this \sectionname{} we generalize the notion of convolution to take geometric images as inputs and return geometric images as outputs.
+The idea is that a geometric image of \tensor{k}{p}s is convolved with a geometric filter of \tensor{k'}{p}s to produce a geometric image that contains \tensor{(k+k')}{p\,p'}s, where each pixel is a sum of outer products.
+These outer products can be contracted down to lower-order tensors using contractions (Definition~\ref{def:contraction}).
+Note that the sidelength $N$ of the geometric image is typically much larger than the sidelength $M$ of the geometric filter.
 
-In principle there could also have been an \emph{eigenvalues} operator that is linear in the input and produces $d$ $0$-$p$-tensors.
-But since any such operation can only reasonably be applied to $2$-$p$-tensors, and even for those, only Hermitian ones, we don't include this operation here.
+\begin{definition}[geometric convolution]\label{def:convolution}
+Let $A \in \mathcal{A}_{N,d,k,p}$ on the $d$-torus, and let $C \in \mathcal{A}_{M,d,k',p'}$ where $M=2m+1\leq N$.
+The geometric convolution $A\ast C$ is a \tensor{(k+k')}{p\,p'} image such that
+\begin{equation}
+    (A\ast C)(\bar\imath) = \sum_{\bar a\in[-m, m]^d} A(\bar\imath - \bar a)\otimes C(\bar a) ~,
+\end{equation}
+where $\bar\imath - \bar a$ is the translation of $\bar\imath$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
+\end{definition}
+This definition is on the torus. To define the convolution in $[N]^d$ instead of the torus, we can pad the image out with zero tensors of the corresponding order and parity.
 
-\section{Higher-order functions of geometric images}\label{sec:nonlinear}
+In addition to contractions and index permutations that act pixel-wise in geometric images, it is possible to change the image size.
+If $N$ is divisible by integer $f>1$, a geometric image of side length $N$ can be reduced in size to a geometric image of side length $N/f$ by the equivalent of convolution with a trivial $f\times f$ scalar filter (geometric filter of \tensor{0}{+}s) operating with a stride of $f$.
+Similarly a geometric image of side length $N$ can be expanded in size to a side length $f\,N$ by application of an interpolation operator with stride $1/f$.
+
+\subsection{Higher-order functions of geometric images}\label{sec:nonlinear}
 
@@
 Non-linear functions can be constructed as polynomials, or sums of products of linear function outputs.
-Products here will be outer products (possibly followed by further geometric convolutions and contractions).
+Products here will be outer products, possibly followed by further geometric convolutions and contractions.
 
 \begin{definition}[outer products of images]
-Given a $k$-$p$-tensor image $A$ and a $k'$-$p'$-tensor image $B$, both of side length $N$, the outer product $A\otimes B$ of these images is a $(k+k')$-$(p+p')$-tensor image such that
+Given $A \in \mathcal{A}_{N,d,k,p}$ and $B \in \mathcal{A}_{N,d,k',p'}$, the outer product $A\otimes B \in \mathcal{A}_{N,d,k+k',p\,p'}$ is defined as
 \begin{equation}
-    (A\otimes B)(\bar\i) = A(\bar\i)\otimes B(\bar\i) ~.
+    (A\otimes B)(\bar\imath) = A(\bar\imath)\otimes B(\bar\imath) ~.
 \end{equation}
-That is, the outer products of geometric images are performed pixelwise.
+for each pixel $\,\bar\imath$. That is, the outer products of geometric images are performed pixel-wise.
 \end{definition}
@@
 \begin{definition}[sums of images]
-Given a $k$-$p$-tensor image $A$ and a $k$-$p$-tensor image $B$, both of side length $N$, the sum $A+B$ of these images is a $k$-$p$-tensor image such that
+Given $A,B \in \mathcal{A}_{N,d,k,p}$, the sum $A+B \in \mathcal{A}_{N,d,k,p}$ is defined as
 \begin{equation}
-    (A+B)(\bar\i) = A(\bar\i)+B(\bar\i) ~.
+    (A+B)(\bar\imath) = A(\bar\imath)+B(\bar\imath) ~.
 \end{equation}
-That is, the sums of geometric images are performed pixelwise.
+for pixel $\,\bar\imath$. That is, the sums of geometric images are performed pixel-wise.
 \end{definition}
 
-Now to construct all degree-2 monomials that produce a $k'$-$p'$-tensor image of side length $N'$ output from a $k$-$p$-tensor image of side length $N$ input, here are the rules:
+HOGG IS CONFUSED: HOW ARE WE MAKING $N'\ne N$ IN THE BELOW?
+
+Now to construct all degree-2 monomials that produce a \tensor{k'}{p'} image of side length $N'$ output from a \tensor{k}{p} image of side length $N$ input, here are the rules:
 \begin{enumerate}
@@
     \item Outer product these two images.
-    \item By the rules in \secref{sec:convolution}, construct any linear function that takes this image outer product as input and produces a $k'$-$p'$-tensor image of side length $N'$ as output.
+    \item By the rules in \secref{sec:convolution}, construct any linear function that takes this image outer product as input and produces a \tensor{k'}{p'} image of side length $N'$ as output.
 \end{enumerate}
@@
 We conjecture that these rules will generate all possible second-degree monomial functions of the input image as long as the filter size is sufficiently large (BEN, is this correct?).
-A polynomial function of degree 2 can be constructed by linear combinations of these second-degree monomial functions along with all the linear functions that also produce $k'$-$p'$-tensor images of side length $N'$ (plus perhaps constant images).
+A polynomial function of degree 2 can be constructed by linear combinations of these second-degree monomial functions along with all the linear functions that also produce \tensor{k'}{p'} images of side length $N'$ (plus perhaps constant images).
 Higher-degree monomials and polynomials can be made out of products of lower-degree monomials, by generalization of the above.
@@
 The geometric convolutions of \secref{sec:convolution} are local linear operators that take as input geometric images, and produce as output geometric images.
-They are local in the sense that, at every pixel $\bar\i$ of the output image $(A\ast C)$, the convolution only makes use of the pixels of $A$ within a small neighborhood of of $\bar\i$.
+They are local in the sense that, at every pixel $\bar\imath$ of the output image $(A\ast C)$, the convolution only makes use of the pixels of $A$ within a small neighborhood of of $\bar\imath$.
 Because tensors can be multiplied together using the outer product, and contracted according to the tensor contractions defined in \secref{sec:convolution}, it is possible, by multiplication, to construct a family of \emph{nonlinear} local geometric operators that take as input geometric images and produce as output geometric images.
-This family of nonlinear local operators will be equivariant to the group $G_d$ when the linear local operators of which they are composed are equivariant.
+This family of nonlinear local operators will be equivariant to the group $G_{N,d}$ when the linear local operators of which they are composed are equivariant.
 
@@
 
-\begin{theorem}
-A function $f:\mathcal A_{N,d,k,p}\to A_{N,d, k'', p''}$ is a linear translation equivariant linear function if and only if it is the convolution with a geometric filter $k'$-$p'$-tensor filter $C$ with possibly sidelength $N$, where $k+k'=k''$ and $p+p'=p''$.
-\end{theorem}
+\begin{definition}[Translation of \tensor{k}{p} images and translation equivariance] Given a \tensor{k}{p} image $A$ on the $d$-torus, and a translation $\tau \in (\mathbf{Z}/N \mathbf{Z})^d$, the action $L_\tau A$ produces a \tensor{k}{p} image on the $d$-torus such that
+\begin{equation}
+    (L_\tau A)(\bar \imath) = A(\bar \imath-\tau)
+\end{equation}
+where $\bar\imath - \tau $ is the translation of $\bar\imath$ by $\tau$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
+A function $f:\mathcal A_{N,d,k,p}\to A_{N,d, k'', p''}$ is translation equivariant if $f(L_\tau A) = L_\tau f(A)$ for all $A\in \mathcal A_{N,d,k,p}$ and $ \tau \in (\mathbf{Z}/N \mathbf{Z})^d$.
+\end{definition}
+It follows directly from the Definition \ref{def:convolution} that the convolution with a geometric filter \tensor{k'}{p'} filter is translation equivariant:
+\begin{equation*}
+    (L_\tau A\ast C)(\bar\imath) = \sum_{\bar a\in[-m, m]^d} A((\bar\imath-\tau) - \bar a)\otimes C(\bar a)= L_\tau (A\ast C)(\bar \imath).
+\end{equation*}
 
+\begin{proposition}
+A function $f:\mathcal A_{N,d,k,p}\to A_{N,d, k'', p''}$ is a linear translation equivariant linear function if and only if it is the convolution with a geometric filter \tensor{k'}{p'} filter $C$ with possibly sidelength $N$, where $k+k'=k''$ and $p\,p'=p''$.
+\end{proposition}
 \begin{proof}
 TO DO.
+Also say this is a particular case of bigger theoretical results, e.g.
+cite results by T.
+Cohen and R.
+Kondor.
+\textcolor{blue}{Is this Proposition 1 true?
+Consider the case $N=1$, $A_{1,d,k,p}\cong T_{d,k,p} \cong \mathbf{R}^{d^k}$.
+In this case, all functions $f:\mathcal A_{1,d,k,p}\to A_{1,d, k'', p''}$ are translation equivariant.
+But it is not true that all linear translation equivariant linear functions are the convolution with a geometric filter: e.g.
+$d=2, k=1, k''=2$, then $\mathcal A_{1,2,1,p}\cong \mathbf{R}^2$ and $\mathcal A_{1,2,2,p}\cong \mathbf{R}^4$, the linear function
+$(x_1,x_2)\mapsto (0,0,0, \lambda x_1+\beta x_2)$ is not a convolution with a geometric filter, since a \tensor{1}{p} image is $C=(c_1, c_2)$ and the convolution $(x_1,x_2)\mapsto (x_1,x_2)\otimes(c_1,c_2)=(c_1x_1, c_1x_2, c_2x_1, c_2x_2)$ }
 \end{proof}
@@
 \begin{definition}[Group $B_d$ of symmetries of a $d$-hypercube]
-We denote by $B_d$ the group of rotation and reflection symmetries of the $d$-dimensional hypercube.
+We denote by $B_d$ the group of Euclidean symmetries of the $d$-dimensional hypercube.
 \end{definition}
 
-Notes about the group $B_d$: In $d=2$ there are $8$ group elements, in $d=3$ there are $48$ group elements (and in $d=4$ there are 384).
+The group $B_d$ is often called the {\em hyperoctahedral group} since the $d$-dimensional hyperoctahedron (also known as the cross-polytope) is dual to the hypercube, so they have the same group of symmetries.
+The notation $B_d$ is standard nomenclature coming from the classification theorem for finite irreducible reflection groups \cite{humphreys1990reflection}.
+
+%BBS: I made minor edits.
+In representation theory and algebraic combinatorics it is called the hyperoctahedral group.
+Imo, ``rotation and reflection symmetries" was a misnomer, this just captures the symmetries with fixed point sets of codimension 1 and 2.
+When $d\geq 4$, there are others.
+(I'm being a nomenclature partisan in that some people do use ``rotation" to mean any element of $SO(d)$, whereas I only use it to mean a literal rotation in a plane (fixing the plane's complement).
+But even with the more liberal meaning of rotation, ``rotations and reflections" still does not include the elements with fixed point sets of odd codimension >1, since these are not in $SO(d)$.) I changed it to Euclidean symmetries.
+
+Notes about the group $B_d$: It has $2^d\cdot d!$ elements (so in $d=2$ there are $8$ elements, in $d=3$ there are $48$, and in $d=4$ there are $384$).
 Because the groups $B_d$ are subgroups of $O(d)$, all determinants of the matrix representations of the group elements are either $+1$ or $-1$, and the matrix representation $M(g^{-1})$ of the inverse $g^{-1}$ of group element $g$ is the transpose of the matrix representation $M(g)$ of group element $g$.
 
-Not that the action of $B_d$ is a subgroup of $O(d)$, therefore the action of $B_d$ on $k$-$p$-tensors is the restriction of the action in Definition \ref{def.tensors} to the subgroup.
+Note that because $B_d$ is a subgroup of $O(d)$, the action of $B_d$ on \tensor{k}{p}s is the restriction of the action in Definition \ref{def.tensors} to the subgroup.
 
-\begin{definition}[Action of $B_d$ on $k$-tensor images]
-Given a $k$-tensor image $A$ on the $d$-torus, and a group element $g\in B_d$, the action $g\cdot A$ produces a $k$-tensor image on the $d$-torus such that
+\begin{definition}[Action of $B_d$ on \tensor{k}{p} images]
+Given a \tensor{k}{p} image $A$ on the $d$-torus, and a group element $g\in B_d$, the action $g\cdot A$ produces a \tensor{k}{p} image on the $d$-torus such that
 \begin{equation}
-    (g\cdot A)(\bar\i) = g\cdot A({g^{-1}\cdot \bar\i}) ~.
+    (g\cdot A)(\bar\imath) = g\cdot A({g^{-1}\cdot \bar\imath}) ~.
 \end{equation}
-SOmeThiNg about how $g^{-1}\cdot\bar\i$ works here, possibly preceded or followed by proofs.
+SOmeThiNg about how $g^{-1}\cdot\bar\imath$ works here, possibly preceded or followed by proofs.
 Because of what follows, this has to be defined BOTH for the $N^d$ boxels in the $d$-torus and on the $(2\,m+1)^d$ boxels of the filter (image patch).
@@
 
-\begin{definition}[The group $G_d$, and its action on $k$-tensor images]
-THIS NEEDS TO BE WRITTEN: $G_d$ is the group generated by the elements of $B_d$ and the discrete translations on the $N^d$-pixel lattice on the $d$-torus.
+\begin{definition}[The group $G_{N,d}$, and its action on \tensor{k}{p} images]\label{def:GdN}
+THIS NEEDS TO BE WRITTEN: $G_{N,d}$ is the group generated by the elements of $B_d$ and the discrete translations on the $N^d$-pixel lattice on the $d$-torus.
 \end{definition}
 
-HOGG: Comment on the fact that $G_d$ is a subgroup of the Euclidean group, but that we care about the Euclidean group because we are imagining that the discrete images are images of a continuous world!!
+BBS: Somewhere here we should mention the canonical map $G_{N,d} \rightarrow B_d$ obtained by quotienting out by the translations.
+(This is used in the Hilbert series computation below.) Should I write this part?
 
-\begin{definition}[Equivariant convolution]
-We say that a convolution filter $C$ produces convolutions that are equivariant with respect to the group $G_d$ if for any $k$-tensor image $A$ on the $d$-torus, and any group element $g\in G_d$, we have
+HOGG: Comment on the fact that $G_{N,d}$ is a subgroup of the Euclidean group, but that we care about the Euclidean group because we are imagining that the discrete images are images of a continuous world!!
+
+\begin{definition}[Equivariant convolution]\label{def:GNd equivariant}
+We say that a convolution filter $C$ produces convolutions that are equivariant with respect to the group $G_{N,d}$ if for any \tensor{k}{p} image $A$ on the $d$-torus, and any group element $g\in G_{N,d}$, we have
 \begin{equation}
@@
 \begin{theorem}
-A $k'$-tensor (or $k'$-pseudotensor) convolution filter $C$ produces convolutions that are equivariant with respect to the big group $G_d$ if $C$ is invariant under the small group $B_d$.
+A \tensor{k'}{p'} convolution filter $C$ produces convolutions that are equivariant with respect to the big group $G_{N,d}$ if $C$ is invariant under the small group $B_d$.
 \end{theorem}
@@
 \begin{eqnarray}
-(g\cdot (A* C) )(\bar \i) &=& g\cdot( (A*C)(g^{-1} \cdot \bar \i)) \\
-&=& g \cdot \left (\sum_{a\in [-m,m]^d} A(g^{-1}\cdot \bar \i - a) \otimes C(a)\right)\\
-&=& \sum_{a\in [-m,m]^d} g \cdot A(g^{-1} \cdot \bar \i -a ) \otimes g \cdot C(a).
+(g\cdot (A* C) )(\bar \imath) &=& g\cdot( (A*C)(g^{-1} \cdot \bar \imath)) \\
+&=& g \cdot \left (\sum_{a\in [-m,m]^d} A(g^{-1}\cdot \bar \imath - a) \otimes C(a)\right)\\
+&=& \sum_{a\in [-m,m]^d} g \cdot A(g^{-1} \cdot \bar \imath -a ) \otimes g \cdot C(a).
 \end{eqnarray}
@@
 \begin{eqnarray}
-((g\cdot A)* C) (\bar \i) &=& \sum_{a\in [-m,m]^d} g\cdot A(g^{-1} \cdot (\bar \i - a) ) C(a) \\
-&=& \sum_{a'\in [-m,m]^d} g \cdot A(g^{-1}\cdot \bar \i -a') \otimes C(g\cdot a'),
+((g\cdot A)* C) (\bar \imath) &=& \sum_{a\in [-m,m]^d} g\cdot A(g^{-1} \cdot (\bar \imath - a) ) C(a) \\
+&=& \sum_{a'\in [-m,m]^d} g \cdot A(g^{-1}\cdot \bar \imath -a') \otimes C(g\cdot a'),
 \end{eqnarray}
@@
 
-In practice, a complete set of invariant $k'$-$p'$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
-This works by, for each setting of order $k'$ and parity $p'$, building a collection of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
+In practice, a complete set of invariant \tensor{k'}{p'} filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
+This works by, for each setting of order $k'$ and parity $p'$, building a collection of linearly independent \tensor{k'}{p'} filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
 \begin{equation}
@@
 where $|B_d|$ is the number of group elements (8 for $d=2$ and 27 for $d=3$).
-In detail we took a systematic (not random) approach that, for each setting of $(k',p')$, initialized the averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
+In detail we took a systematic (not random) approach that, for each setting of $(k',p')$, initialized the averaging with a set of independent \tensor{k'}{p'} convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
 There are $M^d\,d^{k'}$ independent filters in this initial set.
@@
   \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter_2_3.png}
+\includegraphics[width=\textwidth]{pics/invariant_filters.png}
   \end{center}
@@
 Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero.
-The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations, but they are.
-There are no invariant pseudoscalar ($0$-$1$-tensor) filters available at $(d,m) = (2,1)$.
-Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor (scalar) filters but multiplied by the identity tensor; there is only one non-trivial invariant $2$-$0$-tensor filter.
-We don't show the $k'$-$p'$-filters at $k'>2$ because we don't know how to visualize them (even the $k'=2$ case is sketchy).
-Note that the vector ($1$-$0$-tensor) filters look like pure divergence and the pseudovector ($1$-$1$-tensor) filters look like pure curl.\label{fig:filters23}}
+The \tensor{2}{p'} filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the \tensor{2}{p'} filters are such that these filters may not look obviously invariant to rotations, but they are, gosh darn it.
+There are no invariant pseudoscalar (\tensor{0}{-}) filters available at $(d,m) = (2,1)$.
+Note that the \tensor{2}{+} 0 and \tensor{2}{+} 3 filters are duplicates of the \tensor{0}{+} 0 and \tensor{0}{+} 2 filters respectively after they are multiplied by the identity tensor.
+Additionally, if we add \tensor{2}{+} 1 and \tensor{2}{+} 2 it is the same as \tensor{0}{+} 1 multiplied by the identity tensor.
+We don't show the \tensor{k'}{p'} filters at $k'>2$ because visualizing them is difficult, even the $k'=2$ case is potentially misleading.
+Note that the vector (\tensor{1}{+}) filters look like pure divergence and the pseudovector (\tensor{1}{-}) filters look like pure curl.\label{fig:filters23}}
 \end{hoggfigure}
@@
   \begin{center}
-\includegraphics[width=\textwidth]{notebooks/filter_2_5.png}
+    \includegraphics[width=\textwidth]{pics/filters_m5.png}
   \end{center}
 \caption{All the filters for $d=2$, $m=2$ ($M=5$), $k'\in [0,1]$.
-Notes: Symbols and colors as in \figref{fig:filters23}.
-We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.
+The symbols and colors are as in \figref{fig:filters23}.
+We don't show the \tensor{2}{p'} filters at $m=2$ because there are 26 of them.
 At $(d,m)=(2,2)$ a pseudoscalar filter appears.
@@
 
-Define equivariance.
+\section{Counting equivariant maps}
 
-Prove translation invariance.
+With an eye to understanding the expressive power of convolution-based functions, we show how to compute the dimension of the vector space of equivariant polynomial maps of given degree.
 
-Prove things about invariant filters and equivariant linear functions.
+Suppose a finite group $G$ acts on a pair of real vector spaces $V$ and $W$.
+The collection of all equivariant polynomial maps $V\rightarrow W$ is also known as the {\em module of covariants},\footnote{In this context the word {\em module} refers to the fact that the set of equivariant polynomial maps is closed under multiplication by arbitrary $G$-invariant (polynomial) functions on $V$ (as well as closed under addition).
+{\em Covariant} is another word for equivariant map.} and written $(\mathbb{R}[V]\otimes W)^G$ (or $\operatorname{Mor}_G(V,W)$, or $\operatorname{Mor}(V,W)^G$).
+The homogeneous covariants of a fixed degree $\ell$ form a finite-dimensional real vector space $(\mathbb{R}[V]\otimes W)^G_\ell$.
+As $\ell$ varies, the dimension of this space does too, thus it forms a nonnegative integer sequence indexed by $\ell = 0,1,2,\dots$.
+The generating function of this sequence,
+\begin{equation}
+H((\mathbb{R}[V]\otimes W)^G,t) := \sum_{\ell \geq 0} \left(\dim_\mathbb{R} (\mathbb{R}[V]\otimes W)^G_\ell\right) t^\ell,
+\end{equation}
+is called the {\em Hilbert series} of the module of covariants.
+A variant \cite[Remark~3.4.3]{derksen2015computational} on a classical result known as {\em Molien's theorem} expresses this generating function as a finite sum of explicit rational functions:
+\begin{equation}\label{eq:molien}
+H((\mathbb{R}[V]\otimes W)^G,t) = \frac{1}{|G|}\sum_{g\in G} \frac{\operatorname{tr}\rho_W(g^{-1})}{\det(I-\rho_V(g)t)},
+\end{equation}
+where $\rho_V(g), \rho_W(g^{-1})$ are matrices describing the actions of $g, g^{-1}$ on $V, W$ respectively.
+(The trace in the numerator is also known as the {\em character} of $W$ evaluated at $g^{-1}$.)
 
-\section{Expressive power of nonlinear convolution-based functions}\label{sec:universality}
+The right side of \eqref{eq:molien} is reasonable to compute in practice.
+To illustrate, we compute it for the group $G_{N,2}$ of Definition~\ref{def:GdN}, with $V=W=\mathcal{A}_{N,2,1,0}$, the space of $2$-dimensional geometric images whose pixels consist of vectors.
+We assume $N$ is odd.
 
-Counting arguments.
+We first compute the character $\operatorname{tr}\rho_W(g)$ for $g\in G_{N,2}$.
+This can be done explicitly by writing down a basis for $W=\mathcal{A}_{N,2,1,0}$ and expressing the action of each element of $G_{N,2}$ in terms of that basis.
+The computation is expedited by the choice of a basis in which the action of $G_{N,2}$ is {\em monomial}, that is, such that each group element sends each basis vector to a scalar multiple of some other (or the same) basis vector: when this holds, only the basis {\em eigenvectors} contribute to the trace.
+The group $B_2$ acts monomially on the standard basis vectors for $\mathcal{T}_{2,1,0}\cong \mathbb{R}^2$, and it follows that $G_{N,2}$ acts monomially on a basis for $A_{N,2,1,0}$  consisting of maps $[N]^d \rightarrow\mathcal{T}_{2,1,0}$ mapping one pixel to one standard basis vector and all other pixels to zero.
+This situation generalizes straightforwardly to higher dimensions $d$ and higher order tensors.
 
-Comments on locality.
+Carrying out the computation of $\operatorname{tr}\rho_W(g)$ in the present case, we denote by $e^0,e^1$ the standard basis for $\mathbb{R}^2$, and then by $\mathbf{e}_{i,j}^p$, with $i,j\in [N]$ and $p\in \{0,1\}$, the map that sends $(i,j)\in [N]^2$ to $e^p$ and the other pixels to 0; thus the $\mathbf{e}_{i,j}^p$ form the desired monomial basis for $\mathcal{A}_{N,2,1,0}$.
+If $g\in G_{N,2}$, then $\mathbf{e}_{i,j}^p$ is not an eigenvector for $g$ unless $g$ fixes the pixel $(i,j)$, and even then, there is no contribution to the trace from pixel $(i,j)$ unless $g$ acts with nonzero trace on the space $\langle\mathbf{e}_{i,j}^0,\mathbf{e}_{i,j}^1 \rangle$ of geometric images supported only at that pixel.
+In turn, if $g$ does fix $(i,j)$, then its trace on $\langle\mathbf{e}_{i,j}^0,\mathbf{e}_{i,j}^1 \rangle$ is equal to the trace of the corresponding element $\overline g$ of $B_2$ (under the canonical map $G_{N,2}\rightarrow B_2$) on $\mathbb{R}^2$.
+This is zero unless $\overline g = \pm I$ (since in all other cases, $\overline g$ is either a $\pi/2$-rotation or a reflection).
+It follows that the only elements of $G_{N,2}$ with nonzero trace on $\mathcal{A}_{N,2,1,0}$ are the identity (with trace $2N^2 = \dim \mathcal{A}_{N,2,1,0}$) and the $\pi$-rotations centered at each of the $N^2$ pixels (each with trace $-2$, coming from the fixed pixel $(i,j)$, where $\mathbf{e}_{i,j}^0$ and $\mathbf{e}_{i,j}^1$ are both negated).
 
-We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).
+Thus the only nonzero terms in the sum \eqref{eq:molien} are those with $g^{-1}$ as just described.
+(Conveniently, $g=g^{-1}$ for all such.) We need to compute $\det (I - \rho_V(g)t)$ for such $g$, where $\rho_V(g)$ can be taken to be the matrix representing the action of $g$ on $V=\mathcal{A}_{N,2,1,0}$ with respect to the basis above.
+For $g=I$ we evidently have
+\begin{equation}
+    \det (I - \rho_V(g)t) = (1-t)^{2N^2}.
+\end{equation} 
+If $g$ is a $\pi$-rotation about the pixel $(i,j)$, then $\mathbf{e}_{i,j}^0,\mathbf{e}_{i,j}^1$ have their signs reversed, while all other pixels are transposed in pairs (say $(i',j')\leftrightarrow (i'',j'')$), with the corresponding $\mathbf{e}_{i',j'}^p$ sent to $-\mathbf{e}_{i'',j''}^p$ and vice versa.
+Thus, for such $g$, the matrix $I - \rho_V(g)t$ can be written block-diagonally with two $1\times 1$ blocks of the form $(1+t)$, and $N^2-1$ blocks of the form \begin{equation}
+\begin{pmatrix}1 & -t \\ -t & 1\end{pmatrix}.
+\end{equation}
+So we have
+\begin{equation}
+\det(I - \rho_V(g)t) = (1+t)^2(1-t^2)^{N^2-1}.
+\end{equation}
+Putting all of this together, \eqref{eq:molien} becomes
+\begin{align}
+  H((\mathbb{R}[V]\otimes W)^G,t) &=  \frac{1}{8N^2}\left(\frac{2N^2}{(1-t)^{2N^2}}+\frac{N^2(-2)}{(1+t)^2(1-t^2)^{N^2-1}}\right)\\
+  &= \frac{1}{4}\left(\frac{1}{(1-t)^{2N^2}} - \frac{1}{(1+t)^2(1-t^2)^{N^2-1}}\right).\label{eq:molien-answer}
+\end{align}
+Expanding \eqref{eq:molien-answer} in a power series and extracting the coefficient of $t^\ell$, we find that the dimension of the space of $G_{N,2}$-equivariant maps $\mathcal{A}_{N,2,1,0}\rightarrow\mathcal{A}_{N,2,1,0}$ is
+\begin{equation}
+    \frac{1}{4}\left(\binom{2N^2+\ell - 1}{\ell} +(-1)^{\ell+1}\sum_{j=0}^{\lfloor \ell/2\rfloor} (\ell - 2j + 1)\binom{N^2+j-2}{j}\right).
+\end{equation}
+Evaluating this expression for low values of $\ell$, we find that the dimension of the space of homogoeneous linear equivariant maps $\mathcal{A}_{N,2,1,0}\rightarrow \mathcal{A}_{N,2,1,0}$ is $(N^2+1)/2$, the dimension of the space of homogeneous quadratic equivariant maps is $(N^4-1)/2$, and the dimension of the space of homogeneous cubic equivariant maps is $(2N^6 + 3N^4 + 4N^2 + 3)/6$.
 
-We can cite that convolutions with small patches may be seen as low pass filters in some way. Reference?
+\section{Equivariant Tensor CNN Architectures}\label{sec:architectures}
 
-\section{Implementation and numerical examples}\label{sec:examples}
+Before building our equivariant CNNs, we must pick the tensor order and parity of our invariant filters and construct them using group-averaging.
+The dimension $d$ is determined by the application, and defines our group $B_d$.
+We first construct all group operators for the $B_d$ group by iterating the generators until the group was complete (closed under all products of operators).
+Then for each order $k$ and parity $p$, the set of possible geometric filters is a vector space of dimension $M^d d^k$.
+Thus we can pick a basis of that many elements where each basis element $e_i$ has exactly one component of the tensor in a single pixel set to 1, and all other values are 0.
+Each of these basis elements is then group-averaged by applying all group operators and averaging the results:
+\begin{equation}
+   \tilde{e}_i = \frac{1}{\qty|B_d|}\sum_{g \in B_d} g \cdot e_i
+\end{equation}
+The results of those group averages are unpacked into a matrix and the singular value decomposition is then run to find the principal components of the non-zero singular values.
+One small detail is that, after the SVD, the filters were renormalized such that the magnitudes of the non-zero filter values were as close to unity as possible, the $k=1$ filters were renormalized but also reoriented such that non-zero divergences were set to be positive, and non-zero curls were set to be counter-clockwise.
+See \figurename~\ref{fig:filters23} and \figurename~\ref{fig:filters25} for the invariant convolutional filters
+in $d=2$ dimensions for filters of side length $M=3$ and $M=5$ respectively.
 
-How do we make the invariant filters? See \secref{sec:equivariant}.
+Now that we have our chosen set of invariant filters, we may build our equivariant neural networks using convolution layers, contraction layers, polynomial layers, and nonlinear activation layers.
 
-How do we make the linear functions? See \secref{sec:convolution}.
+\subsection{Architecture Components}
 
-Now show some fake-data geometric images and show the action of the $d=2$, $M=3$ filters on those images.
+\begin{comment}
+[HOGG NOTES: WHAT'S WRITTEN HERE DEPENDS A LOT ON WHAT IS WRITTEN ABOVE ABOUT WHETHER THE GEOMETRIC IMAGES CAN HAVE MULTIPLE CHANNELS PER PIXEL.]
 
+If the \tensor{k'}{p'} convolutions presented above are going to be actually implemented in a machine-learning method, what would that look like?
+There are combinatoric choices.
+Fundamentally, if the incoming image has $(2\,m+1)^d$ pixels, each of which has multiple channels, in each of which there are \tensor{k}{p}s, then the convolutions also have $(2\,m+1)^d$ pixels, but with (in general) many channels per pixel, containing tensors of different orders and parities.
+Thus, in each layer of any kind of ``network'', the input and output will both contain multiple channels per pixel; each channel will in general have tensors of different order $k$ and parity $p$.
+In going from input to output, every channel of the input is convolved with every available convolution filter to make every channel of the output.
+In addition to applying filters, there are also contractions and Levi-Civita contractions possible.
+
+At each layer, nonlinearity can be applied by taking products of channels, or by applying nonlinear functions (such as RELU CITE) to scalar channels.
+A maximal network architecture would consider every possible convolution operator (up to some order $k'$ and some side length $2\,m+1$), possible products with the Kroenecker and Levi-Civita tensors (the constants), every possible index contraction, and every possible permutation of the remaining open indices.
+The results in the channels available at each layer would be nonlinear functions of weighted sums of the multiple geometric images with common order and parity created by the convolutions and contractions of the images at the previous layer.
+\end{comment}
+
+\begin{comment}
+[HOGG: Now show some fake-data geometric images and show the action of the $d=2$, $M=3$ filters on those images.
 What do they look like? Like calculus operators! See figures.
@@
 \end{hoggfigure}
-
-Now show contractions of products of the above on those images.
+Now show contractions of products of those images.
 \begin{hoggfigure}
@@
   \end{center}
-  \caption{All unconvolved second-degree monomials of $s$, given four chosen filters. For the 2-tensor outputs (order $k=2$), we traced them back to scalars.}
+  \caption{All unconvolved second-degree monomials of $s$, given four chosen filters. For the \tensor{2}{p} outputs (order $k=2$), we traced them back to scalars. HOGG ASKS: WHAT DOES THAT MEAN?}
 \end{hoggfigure}
+\end{comment}
 
-\begin{itemize}
-    \item find the invariant geometric filters
-    \item geometric convolution operator
-    \item pixel-wise outer product
-    \item contraction operator, enumeration of all contractions
-    \item output index permutation, enumeration of all permutations
-    \item enumeration of all monomials up to a fix order and degree 
-    \item remove repeated objects (SVD or just lstsq)?
-    \item linear combinations 
-    \item loss function
-    \item optimization
-\end{itemize}
+We think of each building block of our architecture as a layer whose input and output is a set of images grouped by tensor order and parity.
+This is due to the key constraint that we can only add geometric images that share tensor order and parity.
+This partitioning also allows us to efficiently batch operations.
+The operation of each layer is either convolution, contraction, polynomial, or applying a nonlinear activation function.
+
+A convolution layer takes a set of images and a set of convolution filters.
+For each filter, we take a weighted sum of the images of a particular tensor order and parity and apply the convolution\footnote{The geometric convolution package is implemented in JAX, which in turn uses TensorFlow XLA under the hood.
+This means that convolution is actually cross-correlation, in line with how the term in used in machine learning papers.
+For our purposes this results in at most a sign change, which will come out in the parameterization.} on that sum with that filter.
+Unlike a traditional CNN where the filters are parameterized, our filters are fixed to enforce the equivariance, and the weights of the weighted sums are the parameters.
+A convolution layer can also have an optional dilation where the filters are dilated before convolving.
+Dilations are helpful for expanding the effective size of the filters without having to calculate the invariant filters for larger $M$, which grows quickly; see \cite{convolution_math} for a description of dilated convolution.
+If we use filters with tensor order greater than 0, the tensor order of the images will continue to grow as we apply convolutions.
+Thus we need a way to reduce the tensor order -- we do this with the contraction layer.
+
+Given an input layer and a desired tensor order, the contraction layer performs all unique contractions to get the layer to that tensor order.
+Each unique contraction thus results in a different image.
+We always end the neural network with a contraction layer to return the images to the proper tensor order.
+Since contractions can only reduce the tensor order by multiples of 2, the last convolution layer before the final contraction must result in images of order $k'' + 2N$ for some $N \in \mathbb{N}$ which may be different per image.
+We also may include contraction layers after each convolution to cap the tensor order of each layer to avoid running out of memory as the tensor order grows.
+In practice, $k=5$ seems to be a good max tensor order.
+
+A polynomial layer takes a set of images and a polynomial degree and computes the full polynomial of all the images with each other using the tensor product of geoemtric images, up to the degree of the polynomial.
+Typically, this will result in a combinatorial blowup of images; we can take parameterized sums along the way to reduce the number of images created.
+We could also do a smaller set of products if we have some special domain knowledge.
+However, in practice it is usually better to use nonlinear activations functions, as in typically done in machine learning.
+
+The final type of layer is a nonlinear activation layer.
+In order to maintain equivariance, we can either apply a nonlinearity to a scalar, or scale our tensors by a nonlinearity applied to the norm of the tensor CITE PAPER.
+For this paper, we used the first strategy.
+We apply all possible contractions to all even tensor order images to reduce them to scalars, apply the nonlinearity, then return those images to the layer.
+Any typical nonlinearity works -- ReLU, leaky ReLu, sigmoid, etc.
+
+\subsection{Our Architectures}
+
+\begin{comment}
+There can also be pooling and unpooling layers.
+There can even non-simply connected architectures, in which one path pools and unpools and the other path doesn't (as in map2map; CITE).
+
+That is, there are myriad choices, for the investigator to choose through intuition or validation.
+In addition there are choices about loss function, dropouts, and batch size, as in all machine-learning contexts.
+
+Absent contractions, the tensor order will tend to grow as the number of layers increases.
+We imagine that it will be useful or necessary to limit the maximum order $k$ of tensors in any channel at each layer.
+This will require contractions to keep pace with convolutions and outer products to limit total tensor order.
+\end{comment}
+
+\section{Example Problems}\label{sec:examples}
+
+The problem we are trying to solve is the following: Given a dynamical system such as a gif or a turbulent fluid simulation, can we predict the future steps of the system?
+Let $A_k$ be the geometric image at timepoint $k$ in the simulation.
+Given $A_k$, we want to predict $A_{k+1}$, although we will train our model to predict the difference $A_{k+1} - A_k$.
+Thus we are trying to learn the function $Net(A_k, \theta)$ parameterized by $\theta$ such that the loss between $A_{k+1}$ and $A_k + Net(A_k, \theta)$ is minimized.
+For our loss we will use the root mean squared error (RMSE).
+To improve training we will use the loss over two steps, so if $\widetilde{A}_{k+1} = A_k + Net(A_k, \theta)$ then
+$$\text{Loss} = \text{RMSE}(A_{k+1}, \widetilde{A}_{k+1}) + \text{RMSE}(A_{k+2}, \widetilde{A}_{k+1} + Net(\widetilde{A}_{k+1}, \theta))$$
+
+One key fact relevant to the construction of our neural network is that the input have the same tensor order.
+Thus as the tensor order of our images increases as we apply convolutions, we must perform an appropriate number of contractions at the end to bring them back down to the original size.
+We start by finding the invariant filters of even and odd parity of tensor order 1 and 2.
+We may omit the scalar filters because later we will perform all possible contractions, including the one on the two axes of of 2-0 tensors 0, 1, and 2 that revert them to the scalar filters.
+As we see in fig.
+\ref{fig:filters23}, this leaves us with 9 filters.
+
+Next we have a sequence of convolutional layers, much as in a typical CNN.
+The first layer applies the 9 convolution filters to the input image.
+We take linear combinations of each output image and feed those sums each into the same 9 filters for a second layer.
+These linear combinations make the layers fully connected, with the caveat that we can only add geometric images if they have the same parity and tensor order.
+Next we have a quadratic layer where we take the outer product of every image output from the second layer.
+
+Then we apply one final convolution to linear combinations of the quadratic layer.
+In this final layer, we only apply the convolution if the parity would match the input images parity, and the tensor order could be contracted down the the original images tensor order.
+For example, if the original image had order 1 and parity 0, then an image with tensor order 5 and parity 0 would only be convolved with 2-0 tensor filters.
+Finally, we apply all possible unique contractions to return the images to the original images tensor order, and we take the linear combination of all those resulting images.
+Parameters are used whenever we are taking linear combinations, so in the convolution layers and after doing the contractions.
+
+\section{Related work}\label{sec:related}
+
+The most common method to design equivariant maps is via group convolution, on the group or on the homogeneous space where the features lie.
+Regular convolution of a vector field $f: (\mathbb{Z} /N \mathbb{Z})^d \to \mathbb{R}^c$ and a filter $\phi: (\mathbb{Z} /N \mathbb{Z})^d \to \mathbb{R}^c$ is defined as
+\begin{equation}\label{eq:1}
+    f*\phi(x)= \sum_{y\in(\mathbb{Z} /N \mathbb{Z})^d} \underbrace{\langle f(y), \phi(y-x)\rangle}_{\text{scalar product of vectors}}=\sum_{y\in(\mathbb{Z} /N \mathbb{Z})^d} \sum_{j=1}^c \underbrace{f^j(y) \phi^j(y-x)}_{\in \mathbb{R}}
+\end{equation}
+Our generalization of convolution replaces this scalar product of vectors by the outer product of tensors. Probably the most related
+work is by Brandstetter et al. \cite{ref1} who replace it by the geometric product of multivector inputs and multivector filters of a Clifford Algebra.
+
+\subsection{Clifford Convolution}
+This paper deals with multivector fields, i.e.: vector fields $f:\mathbb{Z}^2 \to (Cl_{p,q}(\mathbb{R}))^c$.
+The real Clifford Algebra $Cl_{p,q}(\mathbb{R})$ is an associative algebra generated by $p+q=d$ orthonormal basis elements: $e_1, \ldots, e_{p+q}\in \mathbb{R}^d$ with the relations:
+\begin{eqnarray}
+     e_i\otimes e_i = +1& (i\leq p),  \\
+     e_j \otimes e_j = -1& (p<j\leq n),\\
+     e_i \otimes e_j = -e_j \otimes e_i & (i\neq j).
+\end{eqnarray}
+For instance, $Cl_{2,0}(\mathbb{R})$ has the basis $\{1, e_1, e_2, e_1\otimes e_2\}$ and is isomorphic to the quaternions $\mathbb{H}$.
+
+The Clifford convolution replaces the elementwise product of scalars of the usual convolution of \eqref{eq:1} by the geometric product of multivectors in the Clifford Algebra:
+\begin{equation}
+    f*\phi(x)= \sum_{y\in(\mathbb{Z} /N \mathbb{Z})^d} \sum_{j=1}^c \underbrace{f^j(y) \otimes \phi^j(y-x)}_{\in Cl_{p,q}(\mathbb{R})},
+\end{equation}
+where $f:\mathbb{Z}^2 \to (Cl_{p,q}(\mathbb{R}))^c$ and $\phi:\mathbb{Z}^2 \to (Cl_{p,q}(\mathbb{R}))^c$
+
+The Clifford Algebra $Cl_{p,q}(\mathbb{R})$  is a quotient of the tensor algebra \begin{equation}
+    T(\mathbb{R}^d)= \bigoplus_{k\geq 0} \underbrace{\mathbb{R}^d\otimes \ldots \otimes {\mathbb{R}^d}}_{k \text{ times}}= \bigoplus_{k\geq 0} (\mathbb{R}^d)^{\otimes k},
+\end{equation} 
+by the two-side ideal $\langle\{v\otimes v - Q(v): v \in \mathbb{R}^d\}\rangle$, where the quadratic form $Q$ is defined by $Q(e_i)=+1$,if  $i\leq p$, and $Q(e_j)=-1$, else $p<j\leq n$.
+Our geometric images are functions $A:(\mathbb{Z}/N\mathbb{Z})^d \to \mathcal T_{d,k,p}$, where $\mathcal{T}_{d,k,p}= (\mathbb{R}^d)^{\otimes k}\subset T(\mathbb{R}^d)$.
+They can be related with the Clifford framework by seeing them as $N$-periodic functions from $\mathbb{Z}^d$ whose image is projected via the quotient map on the Clifford Algebra.
+This projection can be seen as a contraction of tensors.
+
+The Clifford convolution is not equivariant under multivector rotations or reflections.
+But the authors derive a constraint on the filters for $d=2$ which allows to build generalized Clifford convolutions which are equivariant with respect to rotations or reflections of the multivectors.
+That is, they prove equivariance of a Clifford layer under orthogonal transformations if the filters satisfies the constraint: $\phi^i (Rx) = R\phi^i(x)$.
+
+Part of our work can be studied under the unified framework for group equivariant networks on homogeneous spaces derived from a Fourier perspective proposed in \cite{ref2}. 
+\subsection{Unified Fourier Framework}
+In \cite{ref2}, they consider general tensor-valued feature fields, before and after a convolution.
+Their fields are functions $f : G/H \to V$ over the homogeneous space $G/H$ taking values in the vector space $V$ and their filters are kernels $\kappa: G \to Hom (V,V')$.
+Essentially, their convolution replaces the scalar product of vectors of traditional convolution by appliying an homomorphism.
+In particular, if $G$ is a finite group and $H=\{0\}$, they define convolution as
+\begin{equation}
+    \kappa* f(x)= \frac{1}{|G|}\sum_{y\in G}\underbrace{\kappa(x^{-1}y)f(y)}_{\in V'}.  
+\end{equation}
+They give a complete characterization of the space of kernels for equivariant convolutions.
+In our framework, the group is $\mathbb{Z}/N\mathbb{Z}$ and the kernel is an outer product by a filter $C$: $\kappa(g)A(g)=A(g)\otimes C(g)$.
+Note that $\mathbb{Z}/N\mathbb{Z}$ is neither an homogeneous space of $O(d)$ nor of $B^d$
+
+We can analize our problem from an spectral perspective, in particular we can describe all linear equivariant using representation theory, using similar tools as in the proof of the ``only if'' part of Theorem 1 in \cite{ref3}.
+This theorem states that convolutional structure is a sufficient and a necessary condition for equivariance to the action of a compact group.
+Some useful references about group representation theory are \cite{ref5}, a classical book about the theory of abstract harmonic analysis and \cite{ref6}, about the particular applications of it.
+
+\subsection{Linear equivariant maps}
+In this work we define an action  over tensor images of $O(d)$, by rotation of tensors in each pixel; of $B^d$ by rotating the grid of pixels and each tensor in the pixel; and of $(\mathbb{Z}/N \mathbb{Z})^d$ by translation of the grid of pixels.
+The action of each one of these groups $G$ over $\mathcal{T}_{d,k,p}$
+\begin{equation}
+    \Phi_{d,k,p}: G \to GL_{con} (\mathcal{T}_{d,k,p}),
+\end{equation}
+can be decomposed into irreducible representations of $G$:
+\begin{equation}
+    \Phi_{d,k,p} \equiv \bigoplus_{\pi \in \hat{G}} m_{d,k,p}(\pi) \pi.
+\end{equation}
+That is, there is a basis of the Hilbert space $\mathcal{T}_{d,k,p}$ in which the action of $G$ is defined via a linear sparse map.
+In the case of $G$ finite, for all $g\in G$ there is a matrix $P$ splitting the representation in the Hilbert space into its irreducible components
+\begin{equation}
+    P^{-1} \Phi_{d,k,p}(g)P=\bigoplus_{\pi \in \hat{G}} m_{d,k,p}(\pi) \pi(g)
+\end{equation}
+
+Consider now linear maps between Tensor images:
+\begin{equation}
+    \mathcal{C}:\mathcal{T}_{d,k,p} \to \mathcal{T}_{d',k',p'}
+\end{equation}
+Linear equivariant maps satisfy that  $\mathcal{C}\circ \Phi_{d,k,p} = \Phi_{d',k',p'} \circ \mathcal{C}$. That is, if $\tilde{\mathcal{C}}$ is the representation of $\mathcal{C}$ in the above basis,
+\begin{equation}
+    \tilde{\mathcal{C}} \circ \bigoplus_{\pi \in G} m_{d,k,p}(\pi) \pi = \bigoplus_{\pi \in G} m_{d',k',p'}(\pi) \pi \circ \tilde{\mathcal{C}}.
+\end{equation}
+By Schur's Lemma, this implies that $\mathcal{C}\equiv \bigoplus_{\pi \in G} m_{d,k,p}(\pi) Id_{d_\pi}$.
+
+The power of representation theory is not limited to compact groups.
+Mackey machinery allow us to study for instance semidirect products of compact groups and other groups, and in general to relate the representations of a normal subgroup with the ones of the hole group.
+This is the spirit of \cite{ref4}, which makes extensive use of the induced representation theory.
+An introduction to this topic can be found in Chapter 7 in \cite{ref5}.
+
+\subsection{Steerable CNNs}
+This paper deals exclusively with signals $f: \mathbb{Z}^2 \to \mathbb{R}^k$.
+They consider the action of \textit{p4m} on $\mathbb{Z}^2$ by translations, rotations by 90 degrees around any point, and reflections.
+This group is a semidirect product of $\mathbb{Z}^2$ and $H= D^4$ (also called $B^2$).
+Every $x\in \text{\textit{p4m}}$ can be written as $x=tr$, for $t\in \mathbb{Z}^2$ and $r\in D^4$.
+They show that equivariant maps with respect to representations $\rho$ and $\rho'$ of rotations and reflections $H$ lead to equivariant maps with respect to certain representations of $G$ $\pi$ and $\pi'$.
+This means that if we find a linear map $\phi: f \mapsto \phi f$ such that $\phi \rho(h)f=\rho'(h)\phi f$ for all $h\in H$, then for the representation of $G$ $\pi'$ defined by
+\begin{equation}\label{eq:2}
+    \pi'(tr)f(y) = \rho(r) [f ((tr)^{-1}y)], \quad tr\in G, y \in \mathbb{Z}^2,
+\end{equation}
+we automatically have that $\phi \pi(g)f=\pi'(g)\phi f$ for all $g\in G$. This is the representation of $G$ induced by the representation $\rho$ of $H$
+
+Note that, the similarity between the definition of the action of $B^d$ on tensor images and equation \eqref{eq:2}.
+The convolution with a symmetric filter produces easily an equivariant map with respect to  the action of the semidirect product of $\mathbb{Z}^d$ and $B^d$ on the tensor images.
+
+\section{Overview of the study of equivariant maps for machine learning}
+Restricting the learning to a hypothesis class where all the functions are equivariant with respect to some group action is a is a powerful tool widely used in machine learning.
+
+\subsection{Motivation}\label{sec:related2}
+It is believed that engineering a model to be equivariant improves its generalisation and accuracy.
+The generalisation benefit of equivariant linear models has been quantified in \cite{elesedy2021provably}, and the  benefit of incorporating invariance in kernel ridge regression when the target is invariant to the action of a compact group has been quantified in \cite{elesedy2021kernel}.
+This benefit has been observed in many applications, for instance \cite{wang2021incorporating}.
+Also, it is believed to be a powerful remedy against the lack of data and an alternative to data augmentation, a technique to train models that preserve symmetries by adding samples to the training set which are transformed versions of other samples.
+For instance, \cite{wang2022data} derives generalization bounds for data augmentation and for equivariant networks, in the setting of dynamics forecasting of non-stationary and
+non-mixing time series.
+
+Other reason is that constraining linear layers in neural networks to be equivariant allows, in certain cases, to approximate any continuous invariant function.
+For example, in \cite{yarotsky2018universal} they show that invariant networks with respect to a subgroup of the symmetric group can universally approximate any real-valued continue function from $\mathbb{R}^n$; and in \cite{dym2020universality}, a similar result is proved for functions from point clouds to some representation of $SO(3)$.
+\cite{bokman2022zz} deals too with point clouds.
+Kumagai et al.
+proposed in \cite{kumagai2020universal} a unified method to obtain universal approximation theorems for equivariant maps by CNNs.
+Note that our convolution is a particular case of the general group convolution that they study.
+They consider a group $G$ that acts on sets $\mathcal{S}$ and $\mathcal{T}$, a $G$-invariant measure $\nu$ on $\mathcal{S}$, a $G$-invariant functions $v: \mathcal{S} \times \mathcal{T} \rightarrow \mathbb{R}$ and $b \in \mathcal{C}(\mathcal{T})$.
+They define the biased $G$-convolution $C_{\nu, v, b}: \mathcal{C}(\mathcal{S}) \rightarrow \mathcal{C}(\mathcal{T})$ as
+\begin{equation}
+C_{\nu, v, b}[x](t):=\int_{\mathcal{S}} v(t, s) x(s) d \nu(s)+b(t)
+\end{equation}
+Thus our maps can be used to approximate continuous non-linear equivariant maps.
+
+The interest in equivariant maps for machine learning has lead to a huge garden of methods for various kinds of data and to the reinvention of the same algorithms in different application domains.
+Also there are some works about the unifying principles.
+
+Irreducible representations or invariant theory are typically
+used to parameterize the space of equivariant functions.
+
+\subsection{Representation theory}
+The first work using representation theory for invariant and equivariant neural networks was the paper of  Wood et al.
+\cite{WOOD199633} from 1996.
+More recent incarnations of these ideas include the works of \cite{makadia}, \cite{ref4}, \cite{Cohen16}, \cite{cohen2019gauge}, \cite{de2020gauge} and \cite{esteves}.
+
+\subsection{Invariant theory}
+One can use classical invariant theory to compute the generators of the algebra of invariant polynomials.
+For instance, in \cite{2022blum-smith-villar-equivariant} we show how to use the generators of the algebra of invariant polynomials to produce a parameterization of equivariant functions for certain groups and actions.
+A classical introduction to the topic can be found in \cite{weyl}.
+This approach is inspired by the he physical sciences, where the data is subject to rules coming from coordinate freedoms and conservation laws.
+In previous work \cite{villar2021scalars, villar2022dimensionless, yao2021simple} we used these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
+The same idea has explored in \cite{haddadin2021invariant}, \cite{Gripaios_2021}, \cite{yao2021simple}.
+
+\subsection{Graph Neural Networks}
+Graph neural networks (GNNs) can also be seen as equivariant functions that take a graph (or an hypergraph) $A\in (\mathbb{R}^n)^{\otimes k}$ and output an embedding $f(A)\in (\mathbb{R}^n)^{\otimes k'}$ so that $f(\Pi *A)=\Pi*f(A)$, for any relabeling of the vertexes induced by a permutation $\Pi$.
+In \cite{maron2018invariant}, they give a characterization of all permutation invariant and equivariant linear layers.
+Graph methods are versatile, for instance, a common approach to define convolutions on meshes is to interpret them as a graph and apply graph convolutional networks \cite{de2020gauge}.
+There are many works about the universal approximation properties of permutation-invariant functions by certain classes of GNNs (see, for instance \cite{kumagai2020universal}), but also, about the limitation of the expresive powers of GNNs via graph isomorphism tests (see \cite{power}).
+These two perspectives are not only connected but also, they are equivalent, as shown in \cite{graphisotest}.
 
 \section{Discussion}\label{sec:discussion}
-
-Hello world
 
@@
 This group motivates the index summation rules and the other geometric operations in the functions we construct.
-The other group is a discrete group $G_d$ appropriate for $d$-cube lattices of points; it is a sub-group of the continuous group.
+The other group is a discrete group $G_{N,d}$ appropriate for $d$-cube lattices of points; it is a sub-group of the continuous group.
 This is the group we use to average the filters (to make invariant filters) and we use to define equivariance for functions of geometric images.
@@
 HOGG: It is interesting that the convolution operations look (in many cases) very much like discrete versions of vector calculus operations, like div, grad, curl, and so on.
+(Refer back to particular panels of particular figures.)
 That's interesting, and potentially useful.
@@
 
-Note that $k$-$1$-tensor (pseudo) filters only exist for certain triples of $(d, m, k)$.
+Note that \tensor{k}{-} (pseudo) filters only exist for certain triples of $(d, m, k)$.
 
@@
 Leslie Greengard (Flatiron),
-Teresa Huang (JHU), and
-Kate Storey-Fisher (NYU) for valuable discussions.
+Teresa Huang (JHU),
+Kate Storey-Fisher (NYU),
+Kaze Wong (Flatiron),
+and the Astronomical Data Group at the Flatiron Institute for valuable discussions and input.
 This project made use of Python~3 \cite{python3}, numpy \cite{numpy}, matplotlib \cite{matplotlib}, and cmastro \cite{cmastro}.
@@
 \end{document}
+
+%SOLEDAD:
+Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-$p$-tensor image of side length $N$ to an output $k'$-$p'$-tensor image of side length $N'$ (where either $N$ is divisible by $N'$, or $N'$ is divisible by $N$).
+Here are the linear operations (linear in the input image) that are permitted:
+\begin{enumerate}
+    \item Perform a convolution and index contraction. For this step there are two qualitatively different options:
+        \begin{itemize}
+            \item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
+              Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
+              Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
+              %There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
+            \item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
+              Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
+              Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
+              %There are $$(d-1)!\,\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
+              Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
+              %There are $$\frac{1}{\ell!}\,\prod_{i=1}^{\ell}\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
+        \end{itemize}
+    \item Perform (generally after the convolution and contraction) any index permutation you like (provided that $k'>1$).
+      There are $k'!$ choices for the index permutation.
+    \item Perform (generally after all the above) a pooling or unpooling step to get the image to side length $N'$.
+\end{enumerate}
+The objects obtained by this algorithm may not be linearly independent.
+
+In principle there could also have been an \emph{eigenvalues} operator that is linear in the input and produces $d$ $0$-$p$-tensors.
+But since any such operation can only reasonably be applied to $2$-$p$-tensors, and even for those, only Hermitian ones, we don't include this operation here.
